### 2023-11-09 Tex/proposal.tex
Initial Commit

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage[version=4]{mhchem}
\usepackage{stmaryrd}
\usepackage{bbold}
\usepackage{enumitem}
\usepackage{hanging} %reference indent
\usepackage[a4paper,left=25mm,right=25mm,top=40mm,bottom=40mm]{geometry}
\usepackage{sectsty}
\sectionfont{\fontsize{10}{15}\selectfont}
\subsectionfont{\fontsize{10}{15}}
\usepackage[mathlines,switch]{lineno}
\usepackage{amsthm}
\usepackage{kotex}
\usepackage{xcolor}
\usepackage{graphicx} % insert image
\usepackage{multicol} % multi column
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{tabularx}
\usepackage{multirow} % multirow in table
\usepackage{tikz}
\usepackage{subcaption}
\modulolinenumbers[5]
% \linenumbers

\usepackage{titlesec}
\titleformat{\section}
{\normalfont\fontsize{12}{12}\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}
{\normalfont\fontsize{12}{12}\bfseries}{\thesubsection}{1em}{}
\titlespacing{\paragraph}{0pt}{\parskip}{5pt}

\usepackage{indentfirst} % first line indent
\setlength{\parindent}{10pt}

\renewcommand{\arraystretch}{1.1}

\title{\textbf{Latent space of VAE for DAG learning}}

\author{김당찬, 강병국, 김재석, 김민찬}

\begin{document}
\maketitle

\section{Research Objective}

본 연구의 목적은 VAE(Variational AutoEncoder)를 기반으로 한 GNN(Graph neural network)의 잠재공간(latent space)에 대한 통계적 접근 및 해석 방법을 찾는 것이다. 일반적으로 linear SEM 가정에서 VAE를 이용하면 노이즈 행렬을 잠재변수(latent variable)로 간주할 수 있다. 이때 노이즈 확률변수가 잠재변수에 대응되기 때문에, Gaussian 노이즈의 경우 정규사전분포를 사용할 수 있으나 non-Gaussian 분포를 따르는 경우 사전분포 설정이 어렵다는 문제가 존재한다. \\

이에, 본 연구에서는 잠재공간의 구조와 분포에 대한 통계적 해석을 진행한다. 구체적으로 VAE에서의 reparametrization trick을 이용하기 위해 non-Gaussian prior에 대한 surrogate distribution 혹은을 이용한다. 이를 통해 적절한 approximated prior distribution를 도출하여 non-Gaussian linear SEM을 학습하는 방법을 제안하고자 한다. 또한, 이를 바탕으로 랜덤 생성된 non-Gaussian DAG에 대한 시뮬레이션과 document 데이터 등의 실제 데이터에 대한 실험을 진행한다.

\section{Preliminary Results}

\subsection{DAG and Linear SEM}

DAG $\mathcal{G}=(V,E)$가 $|V|=m$ 개의 노드를 갖고 가중인접행렬(weighted adjacency matrix) $A\in\mathbb R^{m\times m}$ 을 갖는다고 가정하자. $m$개의 각 노드는 random variable에 대응되는데, 이때 faithful joint distribution으로부터 얻은 표본을 $X\in\mathbb R^{m\times d}$ 라고 하자. Linear SEM에서는 그래프 구조에 대해 다음을 가정한다.
\begin{align}
    X &= A^TX+Z \\
    &= (I-A^T)^{-1}Z. \label{linearSEM}
\end{align}
행렬 $Z\in\mathbb R^{m\times d}$는 노이즈 행렬을 나타낸다. DAG에서는 topological sorting으로 인접행렬이 상삼각행렬이 되므로, 두번째 식이 성립한다. \\

DAG 학습은 DAG의 구조를 주어진 데이터로부터 재구성하는 것이다. 그러나 이는 NP-hard problem에 해당한다. \cite{ChickeringNPhard} 이로 인해 다양한 가정을 기반으로 한 DAG 학습 알고리즘들이 존재한다. 학습 알고리즘에는 크게 Constraint-based methods \cite{Spirtes2000} 와 Score-based methods \cite{ChickeringGES} 가 있다. 또한 보통 이런 방법들은 변수 사이의 functional form을 가정하고 학습하는데, 위의 경우처럼 (식 \ref*{linearSEM}) linear SEM을 가정할 수 있다 \cite{GhoshalSEM}. Score-based methods는 DAG의 search space가 너무 큰 combinatorial problem, 즉 노드의 개수 증가에 따라 score를 고려해야 하는 DAG의 갯수가 superexponential하게 증가하는 문제가 있었는데, 근래 이를 continuous optimization 문제로 해결한 방법도 제안되었다 \cite{zheng2018dags}. 이를 토대로 DAG learning을 Deep learning Framework에서 진행할 수 있게 되었다.

\subsection{Variational Autoencoder}

VAE\cite{kingma2013auto}에서는 데이터 $x$가 잠재변수 $z$로부터 생성된다고 가정한다. 입력변수 $x$의 분포 $p_{\theta}(x)$나, 잠재변수 $z$의 사후분포 $p_{\theta}(z|x)$는 직접 계산하기 어렵기에(intractable), true posterior $p_{\theta}(z|x)$를 근사하는 다른 분포족 $q_{\phi}(z|x)$를 도입한다. 이로부터 $x$의 가능도함수는 다음과 같이 표현된다. \\

\begin{equation}
    \log p_{\theta}(x) = D_{KL}(q_{\phi}(z|x)||p_{\theta}(z|x)) -D_{KL}(q_{\phi}(z|x)||p_{\theta}(z)) + \mathbb{E}_{q_{\phi}(z|x)}[\log{p_{\theta}(x|z)}]
\end{equation}

이로부터 VAE의 목적함수인 ELBO(evidence lower bound)는 다음과 같이 정의된다. VAE에서는 ELBO를 최대화하여 로그가능도를 최대화하고, 사후분포와 그 근사분포의 KL 발산을 최소화한다.
\begin{eqnarray*}
\mathcal{L}(\theta,\phi, x) := -D_{KL}(q_{\phi}(z|x)||p_{\theta}(z)) + \mathbb{E}_{q_{\phi}(z|x)}[\log{p_{\theta}(x|z)}]
  \label{ELBO}
\end{eqnarray*}

\paragraph*{reparametrization trick}

VAE와 같은 stochastic variational inference에서는 reparametrization trick으로 \cite{reparametrization} $\phi$에 대한 gradient 계산을 간접적으로 수행한다. 조건부분포 $q_{\phi}(z|x)$ 로부터 표본을 추출하는 대신, 확률변수 $\epsilon \sim p(\epsilon)$을 도입해 다음과 같이 $z$를 deterministic, differentiable $z=g_{\phi}(\epsilon, x)$ 로 reparametrization하여 나타낸다.

\begin{equation}
\mathbb{E}_{q_{\phi}(z|x)}[- \log{q_{\phi}(z|x)} + \log{p_{\theta}(x,z)}] = \mathbb{E}_{p(\epsilon)}[- \log{q_{\phi}(g_{\phi}(\epsilon, x)|x)} + \log{p_{\theta}(x,g_{\phi}(\epsilon, x))}]
\label{reparametrization}
\end{equation}

\paragraph*{Latent space of VAE}

VAE에서 정의되는 잠재공간는 데이터의 representation을 저차원의 잠재변수로 표현한 것이다. 인코더에 대한 특정 approximation과 디코더에 대한 orthogonality 제약조건을 하에서 PCA와 VAE가 유사한 latent spcae를 갖는다는 것이 알려져 있다. \cite{rolinek2019variational}

\subsection{Graph Neural Network}

Linear SEM 식 (\ref*{linearSEM})를 $X=f_A(Z)$ 라고 표기하면, 이는 일반적인 GNN의 기본 구조가 되며 나아가 다음과 같은 generalized linear SEM도 고려할 수 있다. \cite{yu2019daggnn} 
\begin{equation}
    X = f_2((I-A^T)^{-1}f_1(Z))
    \label{gSEM}
\end{equation}

$f_1,f_2$에 신경망 구조를 가정하면, 이는 autoencoder 형태의 모형이 된다.

\subsubsection*{Latent variable for Graph data}

Graph structured data를 VAE와 같이 잠재공간 기반의 모델로 다루기 위해서는 잠재변수에 대한 설정이 필요하다. Table 1은 graph structured data를 다룬 여러 VAE 선행연구들에서 잠재공간의 구조가 어떤 형태로 가정되었는지 정리한 표이다.

% Table 1

\begin{table}[h!]
    \centering
    \begin{tabular}{|l|c|c|}
        \hline
        \textbf{Model} & \textbf{Latent variable} & \textbf{Prior} \\
        \hline
        DAG-GNN\cite{yu2019daggnn} & Noise matrix from Linear SEM, $Z\in\mathbb{R}^{m\times d}$& $\mathcal{MN}_{m,d}(0,I,I)$ \\
        \hline
        VGAE\cite{kipf2016variational} & stochastic latent variable, $Z\in\mathbb{R}^{n\times d}$ & $\mathcal{N}(0,I)$ \\
        \hline
        \multirow{2}{4em}{CausalVAE\cite{Yang_2021_CVPR}} & Endogeneous latent variable $\bf z\in\mathbb{R}^n$ & $\bf z = \mathbf{A}^T \bf z + \bf u$ \\
        & Exogeneous latent variable $\bf u\in\mathbb{R}^n$ & $\bf u\sim\mathcal{N}(0,I)$ \\
        \hline
        Dirichlet Graph VAE\cite{li2020dirichlet} & cluster membership of each node, $Z_i$ & $Z_i\sim\mathcal{D}(\alpha)$ \\
        \hline
        \end{tabular}
        \caption{Different latent structure of VAEs for graph-structured data}
\end{table}

\section{Experiment}

실험 과정은 simulation study와 application study를 진행한다. Simulation에서는, Erdös-Rényi model \cite{erdds1959random}로 random DAG에 대한 데이터를 생성하여, 제안된 모델을 바탕으로 학습 및 평가를 진행한다. 문서 데이터에는 GNN 기반 학습 방법이 다수 제안된 바 있다. \cite{peng2017crosssentence,marcheggiani2017encoding,schlichtkrull2017modeling,zhang2020document} 다만, citation network \cite{CLOUGH2016235} 와 같이 text(document) 사이의 관계를 나타내는 그래프 데이터의 경우, DAG 기반의 분석이 가능하다. \cite{wuetal2011} 이러한 데이터에 대해 제안된 모델을 적용해 보고자 한다. 모델의 성능 평가의 경우 SHD(Structural Hamming Distance)와 FDR(False Discovery Rate)를 사용한다. \cite{yu2019daggnn}

% Biblography
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}

### 2023-11-09 References/DAG-GNN/models.py

@@ -1,0 +1,144 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import math
+
+from torch.autograd import Variable
+from utils import my_softmax, get_offdiag_indices, gumbel_softmax, preprocess_adj, preprocess_adj_new, preprocess_adj_new1, gauss_sample_z, my_normalize
+
+_EPS = 1e-10
+
+
+class MLPEncoder(nn.Module):
+    """MLP encoder module."""
+    def __init__(self, n_in, n_xdims, n_hid, n_out, adj_A, batch_size, do_prob=0., factor=True, tol = 0.1):
+        super(MLPEncoder, self).__init__()
+
+        self.adj_A = nn.Parameter(Variable(torch.from_numpy(adj_A).double(), requires_grad=True))
+        self.factor = factor
+
+        self.Wa = nn.Parameter(torch.zeros(n_out), requires_grad=True)
+        self.fc1 = nn.Linear(n_xdims, n_hid, bias = True)
+        self.fc2 = nn.Linear(n_hid, n_out, bias = True)
+        self.dropout_prob = do_prob
+        self.batch_size = batch_size
+        self.z = nn.Parameter(torch.tensor(tol))
+        self.z_positive = nn.Parameter(torch.ones_like(torch.from_numpy(adj_A)).double())
+        self.init_weights()
+
+    def init_weights(self):
+        for m in self.modules():
+            if isinstance(m, nn.Linear):
+                nn.init.xavier_normal_(m.weight.data)
+            elif isinstance(m, nn.BatchNorm1d):
+                m.weight.data.fill_(1)
+                m.bias.data.zero_()
+
+
+    def forward(self, inputs, rel_rec, rel_send):
+
+        if torch.sum(self.adj_A != self.adj_A):
+            print('nan error \n')
+
+        # to amplify the value of A and accelerate convergence.
+        adj_A1 = torch.sinh(3.*self.adj_A)
+
+        # adj_Aforz = I-A^T
+        adj_Aforz = preprocess_adj_new(adj_A1)
+
+        adj_A = torch.eye(adj_A1.size()[0]).double()
+        H1 = F.relu((self.fc1(inputs)))
+        x = (self.fc2(H1))
+        logits = torch.matmul(adj_Aforz, x+self.Wa) -self.Wa
+
[…]

### 2023-11-09 References/DAG-GNN/train.py

@@ -1,0 +1,561 @@
+
+''''
+Main function for traininng DAG-GNN
+
+'''
+
+from __future__ import division
+from __future__ import print_function
+
+import time
+import argparse
+import pickle
+import os
+import datetime
+
+# import torch
+import torch.optim as optim
+from torch.optim import lr_scheduler
+import math
+
+# import numpy as np
+from utils import *
+from models import *
+
+parser = argparse.ArgumentParser()
+
+# -----------data parameters ------
+# configurations
+parser.add_argument('--data_type', type=str, default= 'synthetic',
+                    choices=['synthetic', 'discrete', 'real'],
+                    help='choosing which experiment to do.')
+parser.add_argument('--data_filename', type=str, default= 'alarm',
+                    help='data file name containing the discrete files.')
+parser.add_argument('--data_dir', type=str, default= 'data/',
+                    help='data file name containing the discrete files.')
+parser.add_argument('--data_sample_size', type=int, default=5000,
+                    help='the number of samples of data')
+parser.add_argument('--data_variable_size', type=int, default=10,
+                    help='the number of variables in synthetic generated data')
+parser.add_argument('--graph_type', type=str, default='erdos-renyi',
+                    help='the type of DAG graph by generation method')
+parser.add_argument('--graph_degree', type=int, default=2,
+                    help='the number of degree in generated DAG graph')
+parser.add_argument('--graph_dist', type=str, default='normal',
+                    help='the distribution of noise matrix type: normal, uniform, exponential, gamma, beta')
+parser.add_argument('--graph_dist_params', type=float, default=[1.0],
+                    help='the parameters of noise distribution')
+parser.add_argument('--graph_linear_type', type=str, default='nonlinear_2',
+                    help='the synthetic data type: linear -> linear SEM, nonlinear_1 -> x=Acos(x+1)+z, nonlinear_2 -> x=2sin(A(x+0.5))+A(x+0.5)+z')
+parser.add_argument('--edge-types', type=int, default=2,
+                    help='The number of edge types to infer.')
+parser.add_argument('--x_dims', type=int, default=1, #changed here
+                    help='The number of input dimensions: default 1.')
+parser.add_argument('--z_dims', type=int, default=1,
[…]

### 2023-11-09 References/DAG-GNN/utils.py

@@ -1,0 +1,762 @@
+import numpy as np
+import torch
+import torch.nn as nn
+import math
+import pickle
+import re
+import glob
+import os
+import time
+import pandas as pd
+import scipy.linalg as la
+import scipy.sparse as sp
+import networkx as nx
+# torch
+from torch.utils.data import Dataset, DataLoader
+from torch.utils.data.dataset import TensorDataset
+from torch.nn import Parameter
+from torch.optim import Adam, SGD
+from torch.nn import functional as F
+from torch.nn import init
+from torch.autograd import Variable
+
+
+# DAG Generator
+
+def simulate_random_dag(d: int,
+                        degree: float,
+                        graph_type: str,
+                        w_range: tuple = (0.5, 2.0)) -> nx.DiGraph:
+    """Simulate random DAG with some expected degree.
+
+    Args:
+        d: number of nodes
+        degree: expected node degree, in + out
+        graph_type: {erdos-renyi, barabasi-albert, full}
+        w_range: weight range +/- (low, high)
+
+    Returns:
+        G: weighted DAG
+    """
+    if graph_type == 'erdos-renyi':
+        prob = float(degree) / (d - 1)
+        B = np.tril((np.random.rand(d, d) < prob).astype(float), k=-1)
+    elif graph_type == 'barabasi-albert':
+        m = int(round(degree / 2))
+        B = np.zeros([d, d])
+        bag = [0]
+        for ii in range(1, d):
+            dest = np.random.choice(bag, size=m)
+            for jj in dest:
+                B[ii, jj] = 1
+            bag.append(ii)
+            bag.extend(dest)
+    elif graph_type == 'full':  # ignore degree, only for experimental use
+        B = np.tril(np.ones([d, d]), k=-1)
+    else:
+        raise ValueError('unknown graph type')
+    # random permutation
+    P = np.random.permutation(np.eye(d, d))  # permutes first axis only
+    B_perm = P.T.dot(B).dot(P)
+    U = np.random.uniform(low=w_range[0], high=w_range[1], size=[d, d])
+    U[np.random.rand(d, d) < 0.5] *= -1
+    W = (B_perm != 0).astype(float) * U
+    G = nx.DiGraph(W)
+    return G
+
[…]

### 2023-11-09 References/NormalizingFlow/flow.py

@@ -1,0 +1,155 @@
+import torch
+import torch.nn as nn
+from tqdm import tqdm
+import numpy as np
+import matplotlib.pyplot as plt
+import math
+
+
+torch.manual_seed(0)
+
+def gaussian_log_pdf(z):
+    """
+    Arguments:
+    ----------
+    - z: a batch of m data points (size: m x data_dim)
+    """
+    return -.5 * (torch.log(torch.tensor([math.pi * 2], device=z.device)) + z ** 2).sum(1)
+
+class PlanarFlow(nn.Module):
+
+    def __init__(self, data_dim):
+        super().__init__()
+
+        self.u = nn.Parameter(torch.rand(data_dim))
+        self.w = nn.Parameter(torch.rand(data_dim))
+        self.b = nn.Parameter(torch.rand(1))
+        self.h = nn.Tanh()
+        self.h_prime = lambda z: (1 - self.h(z) ** 2)
+
+    def constrained_u(self):
+        """
+        Constrain the parameters u to ensure invertibility
+        """
+        wu = torch.matmul(self.w.T, self.u)
+        m = lambda x: -1 + torch.log(1 + torch.exp(x))
+        return self.u + (m(wu) - wu) * (self.w / (torch.norm(self.w) ** 2 + 1e-15))
+
+    def forward(self, z):
+        u = self.constrained_u()
+        hidden_units = torch.matmul(self.w.T, z.T) + self.b
+        x = z + u.unsqueeze(0) * self.h(hidden_units).unsqueeze(-1)
+        psi = self.h_prime(hidden_units).unsqueeze(0) * self.w.unsqueeze(-1)
+        log_det = torch.log((1 + torch.matmul(u.T, psi)).abs() + 1e-15)
+        return x, log_det
+
+class NormalizingFlow(nn.Module):
+
+    def __init__(self, flow_length, data_dim):
+        super().__init__()
+
+        self.layers = nn.Sequential(
+            *(PlanarFlow(data_dim) for _ in range(flow_length)))
+
+    def forward(self, z):
+        log_jacobians = 0
+        for layer in self.layers:
+            z, log_jacobian = layer(z)
+            log_jacobians += log_jacobian
+        return z, log_jacobians
+
+def train(flow, optimizer, nb_epochs, log_density, batch_size, data_dim):
+    training_loss = []
[…]

### 2023-11-09 References/IAF-VAE/layers.py
commit

@@ -1,0 +1,146 @@
+import torch
+import numpy as np
+import torch.nn as nn
+import torch.nn.functional as F
+import torch.nn.utils.weight_norm as wn
+import torch.distributions as D
+
+# Basic Layers 
+# -------------------------------------------------------------------------------------------------------
+
+# taken from https://github.com/jzbontar/pixelcnn-pytorch
+class MaskedConv2d(nn.Conv2d):
+    def __init__(self, mask_type, *args, **kwargs):
+        super(MaskedConv2d, self).__init__(*args, **kwargs)
+        assert mask_type in {'A', 'B'}
+        self.register_buffer('mask', self.weight.data.clone())
+        _, _, kH, kW = self.weight.size()
+        self.mask.fill_(1)
+        self.mask[:, :, kH // 2, kW // 2 + (mask_type == 'B'):] = 0
+        self.mask[:, :, kH // 2 + 1:] = 0
+
+    def forward(self, x):
+        self.weight.data *= self.mask
+        return super(MaskedConv2d, self).forward(x)
+
+class ARMultiConv2d(nn.Module):
+    def __init__(self, n_h, n_out, args, nl=F.elu):
+        super(ARMultiConv2d, self).__init__()
+        self.nl = nl
+
+        convs, out_convs = [], []
+
+        for i, size in enumerate(n_h):
+            convs     += [MaskedConv2d('A' if i == 0 else 'B', args.z_size if i == 0 else args.h_size, args.h_size, 3, 1, 1)]
+        for i, size in enumerate(n_out):
+            out_convs += [MaskedConv2d('B', args.h_size, args.z_size, 3, 1, 1)]
+
+        self.convs = nn.ModuleList(convs)
+        self.out_convs = nn.ModuleList(out_convs)
+
+
+    def forward(self, x, context):
+        for i, conv_layer in enumerate(self.convs):
+            x = conv_layer(x)
+            if i == 0: 
+                x += context
+            x = self.nl(x)
+
+        return [conv_layer(x) for conv_layer in self.out_convs]
+
+
+# IAF building block
+# -------------------------------------------------------------------------------------------------------
+
+class IAFLayer(nn.Module):
+    def __init__(self, args, downsample):
+        super(IAFLayer, self).__init__()
+        n_in  = args.h_size
+        n_out = args.h_size * 2 + args.z_size * 2
+        
[…]

### 2023-11-09 References/IAF-VAE/main.py

@@ -1,0 +1,249 @@
+import torch
+import argparse
+import numpy as np
+import torch.nn as nn
+import torch.nn.functional as F
+
+from os.path import join
+from tensorboardX import SummaryWriter
+from torchvision.utils import save_image
+from collections import OrderedDict as OD 
+from torchvision import datasets, transforms, utils
+
+from layers import IAFLayer
+from utils  import * 
+
+# Model definition
+# ----------------------------------------------------------------------------------------------
+class VAE(nn.Module):
+    def __init__(self, args):
+        super(VAE, self).__init__()
+        self.register_parameter('h', torch.nn.Parameter(torch.zeros(args.h_size)))
+        self.register_parameter('dec_log_stdv', torch.nn.Parameter(torch.Tensor([0.])))
+
+        layers = []
+        # build network
+        for i in range(args.depth):
+            layer = []
+
+            for j in range(args.n_blocks):
+                downsample = (i > 0) and (j == 0)
+                layer += [IAFLayer(args, downsample)]
+
+            layers += [nn.ModuleList(layer)]
+
+        self.layers = nn.ModuleList(layers) 
+        
+        self.first_conv = nn.Conv2d(3, args.h_size, 4, 2, 1)
+        self.last_conv = nn.ConvTranspose2d(args.h_size, 3, 4, 2, 1)
+
+    def forward(self, input):
+        # assumes input is \in [-0.5, 0.5] 
+        x = self.first_conv(input)
+        kl, kl_obj = 0., 0.
+
+        h = self.h.view(1, -1, 1, 1)
+
+        for layer in self.layers:
+            for sub_layer in layer:
+                x = sub_layer.up(x)
+
+        h = h.expand_as(x)
+        self.hid_shape = x[0].size()
+
+        for layer in reversed(self.layers):
+            for sub_layer in reversed(layer):
+                h, curr_kl, curr_kl_obj = sub_layer.down(h)
+                kl     += curr_kl
+                kl_obj += curr_kl_obj
+
+        x = F.elu(h)
+        x = self.last_conv(x)
+        
+        x = x.clamp(min=-0.5 + 1. / 512., max=0.5 - 1. / 512.)
+
+        return x, kl, kl_obj
+
+
+    def sample(self, n_samples=64):
+        h = self.h.view(1, -1, 1, 1)
+        h = h.expand((n_samples, *self.hid_shape))
+        
+        for layer in reversed(self.layers):
+            for sub_layer in reversed(layer):
[…]

### 2023-11-09 model.py

@@ -1,0 +1,8 @@
+# What to do?
+
+# 1. VAE model class
+
+
+
+
+# 2. Normalizing flow model class

### 2023-11-09 train.py

@@ -1,0 +1,32 @@
+# What to do?
+
+
+# 1. Arguments(argparse)
+
+
+
+
+
+
+# 2. Data Loading
+
+
+
+
+
+
+# 3. Model Loading
+
+
+
+
+
+
+# 4. Training Loop (epoch, batch, loss, optimizer, etc.)
+
+
+
+
+
+
+# 5. Save Model (checkpoint, etc.)

### 2023-11-09 data.py
Complete Data.py

@@ -1,20 +1,163 @@
+import numpy as np
+import math
+import networkx as nx
+import scipy.linalg as la
+import scipy.sparse as sp
+
+
 # 1. Generating Random Graph
 
+def generate_random_dag(d, degree, w_range = (0.5, 2.0), seed = 0):
+    """
+    Generate a random Erdos-Renyi DAG with some parameters
 
+    Args:
+        d: number of nodes
+        degree: expected node degree, in + out
+        w_range: range of weights
+        seed: random seed
 
+    Returns:
+        G: nx.DiGraph object
+    """
+    np.random.seed(seed)
 
+    prob = float(degree) / (d - 1)
+    B = np.tril((np.random.rand(d, d) < prob).astype(float), k = -1)
 
+    # random permutation
+    P = np.random.permutation(np.eye(d, d))  # permutes first axis only
+    B_perm = P.T.dot(B).dot(P)
+    U = np.random.uniform(low=w_range[0], high=w_range[1], size=[d, d])
+    U[np.random.rand(d, d) < 0.5] *= -1
+    W = (B_perm != 0).astype(float) * U
+    G = nx.DiGraph(W)
 
-
+    return G
 
 # 2. Generating Linear SEM with independent noise
 
+def generate_linear_sem(graph : nx.DiGraph, 
+                        n : int, dist = 'normal', linear_type = 'linear', 
+                        loc = 0.0, scale = 1.0, seed = 0):
 
+    """
+    Generate a linear SEM with some parameters
 
+    Args:
+        graph: nx.DiGraph object
+        n: number of samples
+        dist: distribution of noise
+        linear_type: type of linear function
+        loc: mean of noise
+        scale: scale of noise
+        seed: random seed
+    """
 
+    np.random.seed(seed)
 
+    A = nx.to_numpy_matrix(graph) # adjacency matrix
+    m = A.shape[0] # number of nodes
+    X = np.zeros((n, m)) # data matrix
+    ordered_nodes = list(nx.topological_sort(graph)) # topological order of nodes
 
+    # generate noise
+    if dist == 'normal':
+        noise = np.random.normal(loc = loc, scale = scale, size = (n, m))
+    elif dist == 'uniform':
+        noise = np.random.uniform(low = loc-scale, high = loc+scale, size = (n, m))
+    elif dist == 'exponential':
+        noise = np.random.exponential(scale = scale, size = (n, m))
+    elif dist == 'laplace':
[…]

### 2023-11-10 model.py
commit

@@ -1,8 +1,120 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import torch.distributions as D
+import torch.nn.utils.weight_norm as wn
+import math
+
+
 # What to do?
 
-# 1. VAE model class
+# args : arguments from train.py (parser)
+
+# 1. VAE Encoder class
+
+class Encoder(nn.Module):
+    """
+    Encoder class for VAE
+    """
+    def __init__(self, args):
+        super(Encoder, self).__init__()
+        self.adj_A = nn.Parameter(Variable(torch.from_numpy(args.adj_A).double(), requires_grad=True))
+        self.factor = args.factor
+
+        self.Wa = nn.Parameter(torch.zeros(args.n_out), requires_grad=True)
+        self.fc1 = nn.Linear(args.n_xdims, args.n_hid, bias = True)
+        self.fc2 = nn.Linear(args.n_hid, args.n_out, bias = True)
+        self.dropout_prob = args.do_prob
+        self.batch_size = args.batch_size
+        self.z = nn.Parameter(torch.tensor(args.tol))
+        self.z_positive = nn.Parameter(torch.ones_like(torch.from_numpy(args.adj_A)).double())
+        self.init_weights()
+
+    def init_weights(self):
+        for m in self.modules():
+            if isinstance(m, nn.Linear):
+                nn.init.xavier_normal_(m.weight.data)
+            elif isinstance(m, nn.BatchNorm1d):
+                m.weight.data.fill_(1)
+                m.bias.data.zero_()
+    
+     def forward(self, inputs, rel_rec, rel_send):
+
+        if torch.sum(self.adj_A != self.adj_A):
+            print('nan error \n')
+
+        # to amplify the value of A and accelerate convergence. (why?)
+        adj_A1 = torch.sinh(3.*self.adj_A)
+
+        # adj_Aforz = $I-A^T$
+        adj_Aforz = (torch.eye(adj_A1.shape[0]).double() - (adj_A1.transpose(0,1)))
+
+        adj_A = torch.eye(adj_A1.size()[0]).double()
+        H1 = F.relu((self.fc1(inputs)))
+        x = (self.fc2(H1))
+        logits = torch.matmul(adj_Aforz, x + self.Wa) - self.Wa
+
+        return ...
 
 
+# 2. Normalizing Flow layer class
+
+class FlowLayer(nn.Module):
+    """
+    Normalizing Flow class (Inverse Autoregressive Flow)
+    """
+
+    def __init__(self, args):
[…]

### 2023-11-10 model.py
model.py 수정

@@ -2,6 +2,7 @@ import torch
 import torch.nn as nn
 import torch.nn.functional as F
 import torch.distributions as D
+from torch.autograd import Variable
 import torch.nn.utils.weight_norm as wn
 import math
 
@@ -21,7 +22,7 @@ class Encoder(nn.Module):
         self.adj_A = nn.Parameter(Variable(torch.from_numpy(args.adj_A).double(), requires_grad=True))
         self.factor = args.factor
 
-        self.Wa = nn.Parameter(torch.zeros(args.n_out), requires_grad=True)
+        self.Wa = nn.Parameter(torch.zeros(args.n_out), requires_grad=True) # Learnable parameter
         self.fc1 = nn.Linear(args.n_xdims, args.n_hid, bias = True)
         self.fc2 = nn.Linear(args.n_hid, args.n_out, bias = True)
         self.dropout_prob = args.do_prob
@@ -38,7 +39,7 @@ class Encoder(nn.Module):
                 m.weight.data.fill_(1)
                 m.bias.data.zero_()
     
-     def forward(self, inputs, rel_rec, rel_send):
+    def forward(self, inputs, rel_rec, rel_send):
 
         if torch.sum(self.adj_A != self.adj_A):
             print('nan error \n')
@@ -113,8 +114,18 @@ class Decoder(nn.Module):
                 m.weight.data.fill_(1)
                 m.bias.data.zero_()
 
-    def forward(self, ...):
-            
-        return ...
+    def forward(self, input_z, origin_A, Wa):
+        # Wa : Encoder에서 학습한 parameter
+        
+        # adj_A_new_inv = $(I-A^T)^{-1}$
+        adj_A_new_inv = torch.inverse(torch.eye(origin_A.shape[0]).double() 
+                                      - origin_A.transpose(0,1))
+
+        mat_z = torch.matmul(adj_A_new_inv, input_z + Wa) - Wa
+
+        H3 = F.relu((self.out_fc1(mat_z)))
+        out = self.out_fc2(H3)
+
+        return mat_z, out
 
         

### 2023-11-10 utils.py
utils.py DAG-GNN에서 긁어왔어요
데이터 생성은 뺐어요
그리고 여기엔 아직 당찬님이 data.py랑 model.py에서 수정한 거 반영안되어있어요

@@ -2,3 +2,576 @@
 
 
 # 1. Metric functions
+
+# 2. utils.py from DAG-GNN
+# removed all codes for generating data: it will be included in data.py
+
+import numpy as np
+import torch
+import torch.nn as nn
+import math
+import pickle
+import re
+import glob
+import os
+import time
+import pandas as pd
+import scipy.linalg as la
+import scipy.sparse as sp
+import networkx as nx
+
+# torch
+from torch.utils.data import Dataset, DataLoader
+from torch.utils.data.dataset import TensorDataset
+from torch.nn import Parameter
+from torch.optim import Adam, SGD
+from torch.nn import functional as F
+from torch.nn import init
+from torch.autograd import Variable
+
+# Metric
+
+def count_accuracy(G_true: nx.DiGraph,
+                   G: nx.DiGraph,
+                   G_und: nx.DiGraph = None) -> tuple:
+    """Compute FDR, TPR, and FPR for B, or optionally for CPDAG B + B_und.
+
+    Args:
+        G_true: ground truth graph
+        G: predicted graph
+        G_und: predicted undirected edges in CPDAG, asymmetric
+
+    Returns:
+        fdr: (reverse + false positive) / prediction positive
+        tpr: (true positive) / condition positive
+        fpr: (reverse + false positive) / condition negative
+        shd: undirected extra + undirected missing + reverse
+        nnz: prediction positive
+    """
+    B_true = nx.to_numpy_array(G_true) != 0
+    B = nx.to_numpy_array(G) != 0
+    B_und = None if G_und is None else nx.to_numpy_array(G_und)
+    d = B.shape[0]
+    # linear index of nonzeros
+    if B_und is not None:
+        pred_und = np.flatnonzero(B_und)
+    pred = np.flatnonzero(B)
+    cond = np.flatnonzero(B_true)
+    cond_reversed = np.flatnonzero(B_true.T)
+    cond_skeleton = np.concatenate([cond, cond_reversed])
+    # true pos
+    true_pos = np.intersect1d(pred, cond, assume_unique=True)
+    if B_und is not None:
+        # treat undirected edge favorably
+        true_pos_und = np.intersect1d(pred_und, cond_skeleton, assume_unique=True)
[…]

### 2023-11-20 loss.py
모델 빌드 완료

@@ -15,11 +15,24 @@ def calculate_reconstruction_loss(preds, target, variance):
 
 # 2. Calculate KL Divergence Loss
 
-def calculate_kl_loss(preds, zsize):
-    predsnew = preds.squeeze(1)
-    mu = predsnew[:,0:zsize]
-    log_sigma = predsnew[:,zsize:2*zsize]
-    kl_div = torch.exp(2*log_sigma) - 2*log_sigma + mu * mu
-    kl_sum = kl_div.sum()
+def calculate_kl_loss(z_0, z_1, z_q_mean, z_q_logvar):
+    log_p_z = log_Normal_standard(z_1, dim=1)
+    log_q_z = log_Normal_diag(z_0, z_q_mean, z_q_logvar, dim=1)
+    kl_loss = -(log_p_z - log_q_z)
 
-    return (kl_sum / (preds.size(0)) - zsize)*0.5
+    return kl_loss.sum() / (z_0.size(0))
+
+
+def log_Normal_diag(x, mean, log_var, average=False, dim=None):
+    log_normal = -0.5 * ( log_var + torch.pow( x - mean, 2 ) * torch.pow( torch.exp( log_var ), -1) )
+    if average:
+        return torch.mean( log_normal, dim )
+    else:
+        return torch.sum( log_normal, dim )
+
+def log_Normal_standard(x, average=False, dim=None):
+    log_normal = -0.5 * torch.pow( x , 2 )
+    if average:
+        return torch.mean( log_normal, dim )
+    else:
+        return torch.sum( log_normal, dim )

### 2023-11-20 model.py

@@ -26,7 +26,7 @@ class FlowLayer(nn.Module):
         """
         # transform L to lower triangular matrix
         L_matrix = L.view(-1, self.args.z_size, self.args.z_size) # resize to get B x L x L
-        LTmask = torch.tril(torch.ones(self.args.z_size, self.args.z_size), k=-1) # lower-triangular mask matrix
+        LTmask = torch.tril(torch.ones(self.args.z_size, self.args.z_size), diagonal=-1) # lower-triangular mask matrix
         I = Variable(torch.eye(self.args.z_size, self.args.z_size).expand(L_matrix.size(0), self.args.z_size, self.args.z_size))
         if self.args.cuda:
             LTmask = LTmask.cuda()
@@ -34,8 +34,7 @@ class FlowLayer(nn.Module):
         LTmask = Variable(LTmask)
         LTmask = LTmask.unsqueeze(0).expand(L_matrix.size(0), self.args.z_size, self.args.z_size)
         LT = torch.mul(L_matrix, LTmask) + I # Lower triangular batches
-
-        z_new = torch.bmm(LT, z.unsqueeze(2)).squeeze(2)
+        z_new = torch.bmm(LT, z) # B x L x L * B x L x 1 = B x L x 1
 
         return z_new
 
@@ -43,27 +42,23 @@ class Encoder(nn.Module):
     """
     Encoder class for VAE
     """
-    def __init__(self, args, tol=0.1):
+    def __init__(self, args, adj_A, tol=0.1):
         super(Encoder, self).__init__()
 
         self.args = args
-        self.adj_A = nn.Parameter(torch.from_numpy(args.adj_A).double(), requires_grad=True)
-        self.Wa = nn.Parameter(torch.zeros(args.n_out), requires_grad=True)
-        self.fc1 = nn.Linear(args.n_xdims, args.n_hid, bias=True)
-        self.fc2 = nn.Linear(args.n_hid, args.n_out, bias=True)
+        self.adj_A = nn.Parameter(torch.from_numpy(adj_A).double(), requires_grad=True)
+        self.Wa = nn.Parameter(torch.zeros(args.z_dims), requires_grad=True)
+        self.fc1 = nn.Linear(args.x_dims, args.encoder_hidden, bias=True)
+        self.fc2 = nn.Linear(args.encoder_hidden, args.z_dims, bias=True)
         self.batch_size = args.batch_size
-
[…]

### 2023-11-20 model.py
코드 간결화
view 부분 없앰

@@ -83,7 +83,7 @@ class Encoder(nn.Module):
 
         adj_A = torch.eye(adj_A1.size()[0]).float()
         h0 = F.relu((self.fc1(inputs.float()))) # first hidden layer
-        h1 = (self.fc2(h0.view(self.batch_size, -1, self.args.encoder_hidden))) # second hidden layer
+        h1 = (self.fc2(h0)) # second hidden layer
         logits = torch.matmul(adj_Aforz, h1 + self.Wa) - self.Wa
 
         # For Flow layer

### 2023-11-21 train.py
Update train.py
간단한 수정과 주석 추가

@@ -118,8 +118,6 @@ def train(epoch, model, best_val_loss, ground_truth_G, lambda_A, c_A, optimizer)
     # set model in training mode
     model.train()
 
-    z = {}
-
     # start training
     if args.warmup == 0:
         beta = 1.
@@ -141,9 +139,16 @@ def train(epoch, model, best_val_loss, ground_truth_G, lambda_A, c_A, optimizer)
         optimizer.zero_grad()
 
 <<<<<<< Updated upstream
+<<<<<<< Updated upstream
         # myA = encoder.adj_A, adj_A_tilt is identity matrix -> 왜 필요한가?
         z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z_q = model.forward(data)
         # 만약 마지막에 에러 -> z_q를 z['0'], z['1']로
+=======
+        # forward VAE
+        # myA = encoder.adj_A, adj_A_tilt is identity matrix -> 왜 필요한가?
+        z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z_0, z_T, LT = model.forward(data)
+        
+>>>>>>> Stashed changes
         edges = logits
 
         """in DAG-GNN
@@ -151,24 +156,31 @@ def train(epoch, model, best_val_loss, ground_truth_G, lambda_A, c_A, optimizer)
         edges = logits
         dec_x, output, adj_A_tilt_decoder = decoder(data, edges, args.data_variable_size * args.x_dims, rel_rec, rel_send, origin_A, adj_A_tilt_encoder, Wa)
         """
+<<<<<<< Updated upstream
 =======
         # Forward VAE
         z = {}
         z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z_q = vae(data)
         # 만약 마지막에 에러 -> z_q를 z['0'], z['1']로
+=======
+>>>>>>> Stashed changes
 
         if torch.sum(output != output):
             print('nan error \n')
 
+<<<<<<< Updated upstream
         # ELBO: 어떻게?
         loss_nll = calculate_reconstruction_loss
         loss_kl = calculate_kl_loss
+=======
[…]

### 2023-11-21 References/DAGGNN/src/train.py
commit

@@ -4,6 +4,7 @@ Main function for traininng DAG-GNN
 
 '''
 
+
 from __future__ import division
 from __future__ import print_function
 
@@ -12,6 +13,7 @@ import argparse
 import pickle
 import os
 import datetime
+from torchviz import make_dot
 
 # import torch
 import torch.optim as optim
@@ -20,7 +22,7 @@ import math
 
 # import numpy as np
 from utils import *
-from models import *
+from modules import *
 
 parser = argparse.ArgumentParser()
 
@@ -41,10 +43,8 @@ parser.add_argument('--graph_type', type=str, default='erdos-renyi',
                     help='the type of DAG graph by generation method')
 parser.add_argument('--graph_degree', type=int, default=2,
                     help='the number of degree in generated DAG graph')
-parser.add_argument('--graph_dist', type=str, default='normal',
-                    help='the distribution of noise matrix type: normal, uniform, exponential, gamma, beta')
-parser.add_argument('--graph_dist_params', type=float, default=[1.0],
-                    help='the parameters of noise distribution')
+parser.add_argument('--graph_sem_type', type=str, default='linear-gauss',
+                    help='the structure equation model (SEM) parameter type')
 parser.add_argument('--graph_linear_type', type=str, default='nonlinear_2',
                     help='the synthetic data type: linear -> linear SEM, nonlinear_1 -> x=Acos(x+1)+z, nonlinear_2 -> x=2sin(A(x+0.5))+A(x+0.5)+z')
 parser.add_argument('--edge-types', type=int, default=2,
@@ -140,21 +140,25 @@ if args.cuda:
 if args.dynamic_graph:
     print("Testing with dynamically re-computed graph.")
 
-# Folder to save the results
+# Save model and meta-data. Always saves in a new sub-folder.
+if args.save_folder:
+    exp_counter = 0
+    now = datetime.datetime.now()
+    timestamp = now.isoformat()
+    save_folder = '{}/exp{}/'.format(args.save_folder, timestamp)
+    # safe_name = save_folder.text.replace('/', '_')
+    os.makedirs(save_folder)
+    meta_file = os.path.join(save_folder, 'metadata.pkl')
+    encoder_file = os.path.join(save_folder, 'encoder.pt')
[…]

### 2023-11-21 References/DAGGNN/src/utils.py

@@ -1,27 +1,25 @@
 import numpy as np
 import torch
+from torch.utils.data.dataset import TensorDataset
+from torch.utils.data import DataLoader
+import torch.nn.functional as F
 import torch.nn as nn
-import math
-import pickle
-import re
-import glob
-import os
-import time
-import pandas as pd
-import scipy.linalg as la
+from torch.autograd import Variable
+import numpy as np
+import scipy.linalg as slin
 import scipy.sparse as sp
 import networkx as nx
-# torch
-from torch.utils.data import Dataset, DataLoader
-from torch.utils.data.dataset import TensorDataset
-from torch.nn import Parameter
-from torch.optim import Adam, SGD
-from torch.nn import functional as F
-from torch.nn import init
-from torch.autograd import Variable
+import pandas as pd
+from pandas import ExcelWriter
+from pandas import ExcelFile
+import os
+import glob
+import re
+import pickle
+import math
+from torch.optim.adam import Adam
 
-
-# DAG Generator
+# data generating functions
 
 def simulate_random_dag(d: int,
                         degree: float,
@@ -64,63 +62,58 @@ def simulate_random_dag(d: int,
     G = nx.DiGraph(W)
     return G
 
-def generate_linearsem(graph : nx.DiGraph, n : int, dist : str, linear_type : str, params : list, seed : int = 0):
-    """
-    Generate data from a linear SEM and given graph
-    :param graph: graph
-    :param n: number of samples
-    :param dist: distribution of the noise (e.g. normal)
-    :param linear_type: type of function (linear, trigonometric, quadratic)
-    :param params: parameters of the SEM
-    :param seed: random seed
+
+def simulate_sem(G: nx.DiGraph,
+                 n: int, x_dims: int,
+                 sem_type: str,
+                 linear_type: str,
+                 noise_scale: float = 1.0) -> np.ndarray:
+    """Simulate samples from SEM with specified type of noise.
+
+    Args:
+        G: weigthed DAG
+        n: number of samples
+        sem_type: {linear-gauss,linear-exp,linear-gumbel}
+        noise_scale: scale parameter of noise distribution in linear SEM
 
     Returns:
-        X: np.array of shape (n, m) where m is the number of nodes
[…]

### 2023-11-21 data.py

@@ -39,7 +39,7 @@ def generate_random_dag(d, degree, w_range = (0.5, 2.0), seed = 0):
 
 def generate_linear_sem(graph : nx.DiGraph, 
                         n : int, dist = 'normal', linear_type = 'linear', 
-                        loc = 0.0, scale = 1.0, seed = 0):
+                        loc = 0.0, scale = 1.0, seed = 0, x_dims : int = 1):
 
     """
     Generate a linear SEM with some parameters
@@ -58,44 +58,47 @@ def generate_linear_sem(graph : nx.DiGraph,
 
     A = nx.to_numpy_array(graph) # adjacency matrix
     m = A.shape[0] # number of nodes
-    X = np.zeros((n, m)) # data matrix
+    X = np.zeros([n, m, x_dims]) # data matrix
     ordered_nodes = list(nx.topological_sort(graph)) # topological order of nodes
-
-    # generate noise
-    if dist == 'normal':
-        noise = np.random.normal(loc = loc, scale = scale, size = (n, m))
-    elif dist == 'uniform':
-        noise = np.random.uniform(low = loc-scale, high = loc+scale, size = (n, m))
-    elif dist == 'exponential':
-        noise = np.random.exponential(scale = scale, size = (n, m))
-    elif dist == 'laplace':
-        noise = np.random.laplace(loc = loc, scale = scale, size = (n, m))
-    elif dist == 'gumbel':
-        noise = np.random.gumbel(loc = loc, scale = scale, size = (n, m))
-    else:
-        raise ValueError('Invalid distribution')
 
     # generate data
     for i in ordered_nodes:
         parents = list(graph.predecessors(i))
         if linear_type == 'linear':
-            eta = np.dot(X[:, parents], A[parents, i])
+            eta = np.dot(X[:, parents, 0], A[parents, i])
         elif linear_type == 'trigonometric':
-            eta = np.sin(np.dot(X[:, parents], A[parents, i]))
+            eta = np.sin(np.dot(X[:, parents, 0], A[parents, i]))
         elif linear_type == 'quadratic':
-            eta = np.dot(X[:, parents], A[parents, i]) ** 2
+            eta = np.dot(X[:, parents, 0], A[parents, i]) ** 2
         else:
             raise ValueError('Unknown linear type')
 
-        X[:, i] = eta.flatten() + noise[:, i]
+        # generate noise
+        if dist == 'normal':
+            noise = np.random.normal(loc = loc, scale = scale, size = n)
[…]

### 2023-11-21 loss.py

@@ -11,7 +11,7 @@ def calculate_reconstruction_loss(preds, target, variance):
     mu2 = target
     neg_log_p = variance + torch.div(torch.pow(mu1-mu2, 2), 2.0 * np.exp(2.0 * variance))
 
-    return neg_log_p.sum() / (target.size(0))
+    return neg_log_p.sum() / (target.size(0)) # average over batch size
 
 # 2. Calculate KL Divergence Loss
 
@@ -33,6 +33,12 @@ def log_Normal_diag(x, mean, log_var, average=False, dim=None):
 def log_Normal_standard(x, average=False, dim=None):
     log_normal = -0.5 * torch.pow( x , 2 )
     if average:
-        return torch.mean( log_normal, dim )
+        return torch.mean(log_normal, dim)
     else:
-        return torch.sum( log_normal, dim )
+        return torch.sum(log_normal, dim)
+
+def kl_gaussian_sem(logits):
+    mu = logits
+    kl_div = mu * mu
+    kl_sum = kl_div.sum()
+    return (kl_sum / (logits.size(0)))*0.5

### 2023-11-21 model.py

@@ -8,6 +8,7 @@ import math
 from utils import *
 import numpy as np
 
+
 from torch.autograd import Variable
 
 class FlowLayer(nn.Module):
@@ -71,7 +72,6 @@ class Encoder(nn.Module):
                 m.bias.data.zero_()
 
     def forward(self, inputs):
-        inputs = inputs.unsqueeze(-1)
 
         if torch.sum(self.adj_A != self.adj_A):
             print('nan error \n')
@@ -136,7 +136,7 @@ class Decoder(nn.Module):
         x_mean = self.sigmoid(self.decoder_mean(out))
         x_logvar = 0.
 
-        return mat_z, out.squeeze(-1), x_mean, x_logvar
+        return mat_z, out, x_mean, x_logvar
 
 
 class VAE(nn.Module):

### 2023-11-21 train.py

@@ -8,6 +8,7 @@ import torch.nn.functional as F
 from torch.optim import lr_scheduler
 from torch.utils.data import Dataset, DataLoader
 from torch.utils.data.dataset import TensorDataset
+from torchviz import make_dot
 import math
 import numpy as np
 from utils import *
@@ -71,7 +72,7 @@ parser.add_argument('--decoder-hidden', type=int, default=64,
                     help='Number of hidden units.')
 parser.add_argument('--temp', type=float, default=0.5,
                     help='Temperature for Gumbel softmax.')
-parser.add_argument('--k_max_iter', type = int, default = 1e2,
+parser.add_argument('--k_max_iter', type = int, default = 100,
                     help ='the max iteration number for searching lambda and c')
 parser.add_argument('--encoder-dropout', type=float, default=0.0,
                     help='Dropout rate (1 - keep probability).')
@@ -115,8 +116,7 @@ if not os.path.exists(folder):
 
 # Save the arguments
 meta_file = os.path.join(folder, 'meta.pkl')
-vae_indep_file = os.path.join(folder, 'vae_indep.pt')
-vae_dep_file = os.path.join(folder, 'vae_dep.pt')
+model_file = os.path.join(folder, 'model.pt')
 
 log_file = os.path.join(folder, 'log.txt')
 log = open(log_file, 'w')
@@ -130,9 +130,9 @@ G = generate_random_dag(d = args.node_size, degree=args.graph_degree, seed=args.
 
 # 2.2. Generate Data
 if args.dependence_type == 1:
-    X, cov, cov_prev = generate_linear_sem_correlated(G, args.data_sample_size, args.dependence_prop, args.seed, return_cov=True)
+    X, cov, cov_prev = generate_linear_sem_correlated(G, args.data_sample_size, args.dependence_prop, args.seed, return_cov=True, x_dims=args.x_dims)
 else:
-    X = generate_linear_sem(graph=G, n=args.data_sample_size, dist=args.graph_dist, linear_type=args.graph_linear_type, loc=args.graph_mean, scale=args.graph_scale, seed=args.seed)
+    X = generate_linear_sem(graph=G, n=args.data_sample_size, dist=args.graph_dist, linear_type=args.graph_linear_type, loc=args.graph_mean, scale=args.graph_scale, seed=args.seed, x_dims=args.x_dims)
[…]

### 2023-11-21 model.py
model.py와 train.py 수정
추가한 것들
1) combination_L을 flow layer에 추가
2) 각종 주석들

@@ -15,7 +15,56 @@ from utils import *
 >>>>>>> Stashed changes
 
 
+<<<<<<< Updated upstream
 # What to do?
+=======
+# z로 변수변환 시 volume을 보존한다고 하는 convex combination (refer to Tomczak & Welling (2017))
+
+class combination_L(nn.Module):
+    def __init__(self,args):
+        super(combination_L, self).__init__()
+        self.args = args
+
+    def forward(self, L, y):
+        '''
+        :param L: batch_size (B) x latent_size^2 * number_combination (L^2 * C)
+        :param y: batch_size (B) x number_combination (C)
+        :return: L_combination = y * L
+        '''
+        # calculate combination of Ls
+        L_tensor = L.view( -1, self.args.z1_size**2, self.args.number_combination ) # resize to get B x L^2 x C
+        y = y.unsqueeze(1).expand(y.size(0), self.args.z1_size**2, y.size(1)) # expand to get B x L^2 x C
+        L_combination = torch.sum( L_tensor * y, 2 ).squeeze()
+        return L_combination
+
+# self.args.number_combination 을 설정해야 될지도
+
+class FlowLayer(nn.Module):
+    """
+    Flow layer class
+    """
+    def __init__(self, args):
+        super(FlowLayer, self).__init__()
+        self.args = args
+    
+    def forward(self, L, z):
+        """
+        :param L: batch_size (B) x latent_size^2 (L^2) from encoder output
+        :param z: batch_size (B) x latent_size (L) from encoder output z0
+        :return: z_new = L * z
+        """
+        # transform L to lower triangular matrix
+        L_matrix = L.view(-1, self.args.z_size, self.args.z_size) # resize to get B x L x L
+        LTmask = torch.tril(torch.ones(self.args.z_size, self.args.z_size), diagonal=-1) # lower-triangular mask matrix
+        I = Variable(torch.eye(self.args.z_size, self.args.z_size).expand(L_matrix.size(0), self.args.z_size, self.args.z_size))
+        if self.args.cuda:
+            LTmask = LTmask.cuda()
+            I = I.cuda()
+        LTmask = Variable(LTmask)
+        LTmask = LTmask.unsqueeze(0).expand(L_matrix.size(0), self.args.z_size, self.args.z_size)
[…]

### 2023-11-21 train.py

@@ -116,7 +116,17 @@ def train(epoch, model, best_val_loss, ground_truth_G, lambda_A, c_A, optimizer)
     mse_train = []
     shd_trian = []
     # set model in training mode
+<<<<<<< Updated upstream
     model.train()
+=======
+    vae.train()
+    # scheduler.step() --> 이거 생략해도 되나요??
+
+    # update optimizer
+    optimizer, lr = update_optimizer(optimizer, args.lr, c_A)
+
+    z = {}
+>>>>>>> Stashed changes
 
     # start training
     if args.warmup == 0:
@@ -141,6 +151,7 @@ def train(epoch, model, best_val_loss, ground_truth_G, lambda_A, c_A, optimizer)
 <<<<<<< Updated upstream
 <<<<<<< Updated upstream
         # myA = encoder.adj_A, adj_A_tilt is identity matrix -> 왜 필요한가?
+<<<<<<< Updated upstream
         z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z_q = model.forward(data)
         # 만약 마지막에 에러 -> z_q를 z['0'], z['1']로
 =======
@@ -150,6 +161,8 @@ def train(epoch, model, best_val_loss, ground_truth_G, lambda_A, c_A, optimizer)
         
 >>>>>>> Stashed changes
         edges = logits
+=======
+>>>>>>> Stashed changes
 
         """in DAG-GNN
         enc_x, logits, origin_A, adj_A_tilt_encoder, z_gap, z_positive, myA, Wa = encoder(data, rel_rec, rel_send)  # logits is of size: [num_sims, z_dims]
@@ -159,15 +172,22 @@ def train(epoch, model, best_val_loss, ground_truth_G, lambda_A, c_A, optimizer)
 <<<<<<< Updated upstream
 =======
         # Forward VAE
+<<<<<<< Updated upstream
         z = {}
         z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z_q = vae(data)
         # 만약 마지막에 에러 -> z_q를 z['0'], z['1']로
 =======
 >>>>>>> Stashed changes
+=======
+        z = {} # z['0'] = z_0 & z['1'] = z_T in Tomczak & Welling (2017)
[…]

### 2023-11-23 References/DAGGNN/src/train.py
commit

@@ -259,11 +259,6 @@ rel_rec = Variable(rel_rec)
 rel_send = Variable(rel_send)
 
 
-x = torch.zeros(args.batch_size, args.data_variable_size, args.x_dims)
-make_dot(encoder(x.double(), rel_rec, rel_send), params=dict(encoder.named_parameters())).render("encoder", format="png")
-exit()
-
-
 # compute constraint h(A) value
 def _h_A(A, m):
     expm_A = matrix_poly(A*A, m)

### 2023-11-23 loss.py

@@ -6,36 +6,37 @@ import numpy as np
 
 # 1. Calculate Reconstruction Loss
 
-def calculate_reconstruction_loss(preds, target, variance):
-    mu1 = preds
+def calculate_reconstruction_loss(decoder_mean, target, variance):
+    """
+    :param decoder_mean: batch_size (B) x number_of_features (F) from decoder output
+    :param target: batch_size (B) x number_of_features (F) from data
+    :param variance: log variance of decoder output = 0.
+    :return: reconstruction loss
+    """
+    
+    mu1 = decoder_mean
     mu2 = target
-    neg_log_p = variance + torch.div(torch.pow(mu1-mu2, 2), 2.0 * np.exp(2.0 * variance))
+    neg_log_p = variance + torch.div(torch.pow(mu1 - mu2, 2), 2.0 * np.exp(2.0 * variance))
 
     return neg_log_p.sum() / (target.size(0)) # average over batch size
 
 # 2. Calculate KL Divergence Loss
 
-def calculate_kl_loss(z_0, z_1, z_q_mean, z_q_logvar):
-    log_p_z = log_Normal_standard(z_1, dim=1)
-    log_q_z = log_Normal_diag(z_0, z_q_mean, z_q_logvar, dim=1)
+def calculate_kl_loss(z_0, z_T, z_q_mean, z_q_logvar, z_dims):
+    log_p_z = log_Normal_standard(z_T, dim=z_dims)
+    log_q_z = log_Normal_diag(z_0, z_q_mean, z_q_logvar, dim=z_dims)
     kl_loss = -(log_p_z - log_q_z)
 
     return kl_loss.sum() / (z_0.size(0))
 
 
-def log_Normal_diag(x, mean, log_var, average=False, dim=None):
+def log_Normal_diag(x, mean, log_var, dim=None):
     log_normal = -0.5 * ( log_var + torch.pow( x - mean, 2 ) * torch.pow( torch.exp( log_var ), -1) )
-    if average:
-        return torch.mean( log_normal, dim )
-    else:
-        return torch.sum( log_normal, dim )
+    return torch.sum( log_normal, dim )
 
-def log_Normal_standard(x, average=False, dim=None):
+def log_Normal_standard(x, dim=None):
     log_normal = -0.5 * torch.pow( x , 2 )
-    if average:
-        return torch.mean(log_normal, dim)
-    else:
-        return torch.sum(log_normal, dim)
[…]

### 2023-11-23 model_new.py

@@ -11,34 +11,6 @@ import numpy as np
 
 from torch.autograd import Variable
 
-class FlowLayer(nn.Module):
-    """
-    Flow layer class
-    """
-    def __init__(self, args):
-        super(FlowLayer, self).__init__()
-        self.args = args
-    
-    def forward(self, L, z):
-        """
-        :param L: batch_size (B) x latent_size^2 (L^2) from encoder output
-        :param z: batch_size (B) x latent_size (L) from encoder output z0
-        :return: z_new = L * z
-        """
-        # transform L to lower triangular matrix
-        L_matrix = L.view(-1, self.args.z_size, self.args.z_size) # resize to get B x L x L
-        LTmask = torch.tril(torch.ones(self.args.z_size, self.args.z_size), diagonal=-1) # lower-triangular mask matrix
-        I = Variable(torch.eye(self.args.z_size, self.args.z_size).expand(L_matrix.size(0), self.args.z_size, self.args.z_size))
-        if self.args.cuda:
-            LTmask = LTmask.cuda()
-            I = I.cuda()
-        LTmask = Variable(LTmask)
-        LTmask = LTmask.unsqueeze(0).expand(L_matrix.size(0), self.args.z_size, self.args.z_size)
-        LT = torch.mul(L_matrix, LTmask) + I # Lower triangular batches
-        z_new = torch.bmm(LT, z) # B x L x L * B x L x 1 = B x L x 1
-
-        return z_new
-
 class Encoder(nn.Module):
     """
     Encoder class for VAE
@@ -48,9 +20,9 @@ class Encoder(nn.Module):
 
         self.args = args
         self.adj_A = nn.Parameter(torch.from_numpy(adj_A).double(), requires_grad=True)
-        self.Wa = nn.Parameter(torch.zeros(args.z_dims), requires_grad=True)
-        self.fc1 = nn.Linear(args.x_dims, args.encoder_hidden, bias=True)
-        self.fc2 = nn.Linear(args.encoder_hidden, args.z_dims, bias=True)
+        self.Wa = nn.Parameter(torch.zeros(args.z_size), requires_grad=True)
+        self.fc1 = nn.Linear(args.z_size, args.encoder_hidden, bias=True)
+        self.fc2 = nn.Linear(args.encoder_hidden, args.z_size, bias=True)
         self.batch_size = args.batch_size
 
         # for other loss
[…]

### 2023-11-23 train.py

@@ -88,6 +88,12 @@ parser.add_argument('--x_dims', type=int, default=1, #changed here
                     help='The number of input dimensions: default 1.')
 parser.add_argument('--z_dims', type=int, default=1,
                     help='The number of latent variable dimensions: default the same as variable size.')
+parser.add_argument('--number_of_flows', type=int, default=5,
+                    help='The number of Householder flows: default 5.')
+parser.add_argument('--flow_type', type=str, default='IAF',
+                    help='The type of flows: "noflow"(DAG-GNN), "IAF", "Householder"')
+parser.add_argument('--lagrange', type=int, default=1,
+                    help='Use lagrange multipliers or not.')
 
 args = parser.parse_args()
 args.z_size = args.node_size # the number of latent variables
@@ -105,11 +111,11 @@ torch.manual_seed(args.seed)
 
 # Folder to save the results, models, dataset
 
-now = datetime.datetime.now().strftime('%Y%m%d_%H%M%S')
+now = datetime.datetime.now().strftime('%m%d_%H%M')
 if args.dependence_type == 1:
-    folder = f'results/dependence/{now}'
+    folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}'
 else:
-    folder = f'results/independence/{args.graph_dist}/{now}'
+    folder = f'results/independence/{args.graph_dist}/{now}_{args.flow_type}_node{args.node_size}'
 
 if not os.path.exists(folder):
     os.makedirs(folder)
@@ -175,8 +181,14 @@ rel_send = Variable(rel_send)
 adj_A = np.zeros([args.node_size, args.node_size])
 
 # 3.1. Load VAE
-
-vae = VAE(args=args, adj_A=adj_A)
+if args.flow_type == 'IAF':
+    vae = VAE_IAF(args=args, adj_A=adj_A)
+elif args.flow_type == 'Householder':
+    vae = VAE_HF(args=args, adj_A=adj_A)
+elif args.flow_type == 'Noflow':
+    vae = daggnn(args=args, adj_A=adj_A)
+else:
+    raise ValueError('Invalid flow type.')
 
 # 3.3. Optimizer
 optimizer = torch.optim.Adam(list(vae.parameters()), lr=args.lr)
[…]

### 2023-11-23 model.py
commit

@@ -227,6 +227,7 @@ class HF(nn.Module):
         # v * v_T
         vT = v.transpose(1,2) # v_T : transpose of v : B x 1 x L
         vvT = torch.bmm(v, vT) # vvT : batchdot( B x L x 1 * B x 1 x L ) = B x L x L
+        
         # v * v_T * z
         vvTz = torch.bmm(vvT, z) # A * z : batchdot( B x L x L * B x L x 1 ).squeeze(2) = (B x L x 1).squeeze(2) = B x L
         # calculate norm ||v||^2
@@ -273,11 +274,12 @@ class VAE_HF(nn.Module):
         # Householder Flow:
         if self.args.number_of_flows > 0:
             v['1'] = self.v_layers[0](h_last.squeeze(-1)).unsqueeze(-1)
-            z['1'] = self.HF(v['1'], z['0'].squeeze(-1)).unsqueeze(-1)
+            z['1'] = self.HF(v['1'], z['0'])
             for i in range(1, self.args.number_of_flows):
-                v[str(i + 1)] = self.v_layers[i](v[str(i)])
+                v[str(i + 1)] = self.v_layers[i](v[str(i)].squeeze(-1)).unsqueeze(-1)
                 z[str(i + 1)] = self.HF(v[str(i + 1)], z[str(i)])
-        return 
+        
+        return z
 
     def forward(self, input, rel_rec, rel_send):
         z = {}

### 2023-11-23 train.py

@@ -89,9 +89,9 @@ parser.add_argument('--x_dims', type=int, default=1, #changed here
 parser.add_argument('--z_dims', type=int, default=1,
                     help='The number of latent variable dimensions: default the same as variable size.')
 parser.add_argument('--number_of_flows', type=int, default=5,
-                    help='The number of Householder flows: default 5.')
+                    help='The number of HF flows: default 5.')
 parser.add_argument('--flow_type', type=str, default='IAF',
-                    help='The type of flows: "noflow"(DAG-GNN), "IAF", "Householder"')
+                    help='The type of flows: "noflow"(DAG-GNN), "IAF", "HF"(Householder).')
 parser.add_argument('--lagrange', type=int, default=1,
                     help='Use lagrange multipliers or not.')
 
@@ -183,7 +183,7 @@ adj_A = np.zeros([args.node_size, args.node_size])
 # 3.1. Load VAE
 if args.flow_type == 'IAF':
     vae = VAE_IAF(args=args, adj_A=adj_A)
-elif args.flow_type == 'Householder':
+elif args.flow_type == 'HF':
     vae = VAE_HF(args=args, adj_A=adj_A)
 elif args.flow_type == 'Noflow':
     vae = daggnn(args=args, adj_A=adj_A)

### 2023-11-23 model.py
오타수정

@@ -139,7 +139,7 @@ class Decoder(nn.Module):
         return mat_z, out, x_mean, x_logvar
 
 
-class VAE(nn.Module):
+class VAE_IAF(nn.Module):
     def __init__(self, args, adj_A):
         super(VAE, self).__init__()
 

### 2023-11-23 model.py
오타수정

@@ -141,7 +141,7 @@ class Decoder(nn.Module):
 
 class VAE_IAF(nn.Module):
     def __init__(self, args, adj_A):
-        super(VAE, self).__init__()
+        super(VAE_IAF, self).__init__()
 
         self.args = args
         self.encoder = Encoder(args, adj_A=adj_A)

### 2023-11-23 train.py
총 훈련시간 나오게 수정

@@ -472,13 +472,13 @@ except KeyboardInterrupt:
     print('threshold 0.3, Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
 
 
-f = open(folder + '/trueG', 'w')
+f = open(folder + '/trueG.txt', 'w')
 matG = np.matrix(nx.to_numpy_array(G))
 for line in matG:
     np.savetxt(f, line, fmt='%.5f')
 f.closed
 
-f1 = open(folder + '/predG', 'w')
+f1 = open(folder + '/predG.txt', 'w')
 matG1 = np.matrix(origin_A.data.clone().numpy())
 for line in matG1:
     np.savetxt(f1, line, fmt='%.5f')
@@ -487,6 +487,9 @@ f1.closed
 # LT to pickle
 pickle.dump(LT, open(folder + '/LT.pkl', 'wb'))
 
+# Total training time
+print("Optimization Finished!")
+print("Total time elapsed: {:.4f}s".format(time.time() - t_total), file=log)
 
 if log is not None:
     print(folder)

### 2023-11-23 train.py
진행바 나오게 수정

@@ -15,6 +15,7 @@ from utils import *
 from model import *
 from data import *
 from loss import *
+from tqdm import tqdm
 
 # 1. Arguments(argparse)
 
@@ -233,7 +234,7 @@ def update_optimizer(optimizer, original_lr, c_A):
 # 4. Training Loop (epoch, batch, loss, optimizer, etc.)
 
 
-def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer):
+def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
     # set loss to 0
     train_loss = 0
     train_re = 0
@@ -331,31 +332,57 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer):
         kl_train.append(loss_kl.item())
         shd_train.append(shd)
 
-    if args.lagrange:
-        print(h_A.item()) # print h(A) value
     nll_val = []
     acc_val = []
     kl_val = []
     mse_val = []
-
-    print('Epoch: {:04d}'.format(epoch),
-          'nll_train: {:.10f}'.format(np.mean(nll_train)),
-          'kl_train: {:.10f}'.format(np.mean(kl_train)),
-          'ELBO_loss: {:.10f}'.format(np.mean(kl_train)  + np.mean(nll_train)),
-          'mse_train: {:.10f}'.format(np.mean(mse_train)),
-          'shd_trian: {:.10f}'.format(np.mean(shd_train)),
-          'time: {:.4f}s'.format(time.time() - t))
-    if np.mean(nll_val) < best_val_loss:
-        torch.save(model.state_dict(), model_file)
-        print('Best model so far, saving...')
+    
+    if pbar is None:
         print('Epoch: {:04d}'.format(epoch),
-              'nll_train: {:.10f}'.format(np.mean(nll_train)),
-              'kl_train: {:.10f}'.format(np.mean(kl_train)),
-              'ELBO_loss: {:.10f}'.format(np.mean(kl_train)  + np.mean(nll_train)),
-              'mse_train: {:.10f}'.format(np.mean(mse_train)),
-              'shd_trian: {:.10f}'.format(np.mean(shd_train)),
-              'time: {:.4f}s'.format(time.time() - t), file=log)
-        log.flush()
+            'nll_train: {:.10f}'.format(np.mean(nll_train)),
[…]

### 2023-11-23 train.py
일부 출력수정

@@ -440,8 +440,8 @@ try:
                     best_epoch = epoch
                     best_MSE_graph = graph
 
-            print("Optimization Finished!")
-            print("Best Epoch: {:04d}".format(best_epoch))
+            # print("Optimization Finished!")
+            # print("Best Epoch: {:04d}".format(best_epoch))
             if ELBO_loss > 2 * best_ELBO_loss:
                 break
 
@@ -453,8 +453,8 @@ try:
             else:
                 break
 
-            # update parameters
-            # h_A, adj_A are computed in loss anyway, so no need to store
+        # update parameters
+        # h_A, adj_A are computed in loss anyway, so no need to store
         h_A_old = h_A_new.item()
         lambda_A += c_A * h_A_new.item()
 

### 2023-11-23 train.py
초기반복에서 warning 제거

@@ -370,7 +370,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
         pbar.set_description('Epoch: {:04d}'.format(epoch))
         pbar.set_postfix(to_print)
     
-        if np.mean(nll_val) < best_val_loss:
+        if nll_val and (np.mean(nll_val) < best_val_loss):
             torch.save(model.state_dict(), model_file)
             pbar.write('Best model so far, saving...')
             pbar.write('Epoch: {:04d}'.format(epoch) +

### 2023-11-23 train.py
모든 케이스 출력 다되도록 변경

@@ -500,9 +500,26 @@ except KeyboardInterrupt:
     print('threshold 0.3, Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
 
 # Save the Graph metrics
-fdr, tpr, fpr, shd, nnz = count_accuracy(G, nx.DiGraph(best_ELBO_graph))
-print('Best ELBO Graph Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
+ls = ['ELBO', 'NLL', 'MSE']
+for idx, graph_res in enumerate([best_ELBO_graph, best_NLL_graph, best_MSE_graph]):
+    fdr, tpr, fpr, shd, nnz = count_accuracy(G, nx.DiGraph(graph_res))
+    print('Best {} Graph Accuracy: fdr'.format(ls[idx]), fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
 
+graph = origin_A.data.clone().numpy()
+graph[np.abs(graph) < 0.1] = 0
+# print(graph)
+fdr, tpr, fpr, shd, nnz = count_accuracy(G, nx.DiGraph(graph))
+print('threshold 0.1, Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
+
+graph[np.abs(graph) < 0.2] = 0
+# print(graph)
+fdr, tpr, fpr, shd, nnz = count_accuracy(G, nx.DiGraph(graph))
+print('threshold 0.2, Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
+
+graph[np.abs(graph) < 0.3] = 0
+# print(graph)
+fdr, tpr, fpr, shd, nnz = count_accuracy(G, nx.DiGraph(graph))
+print('threshold 0.3, Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
 
 f = open(folder + '/trueG.txt', 'w')
 matG = np.matrix(nx.to_numpy_array(G))

### 2023-11-23 train.py
폴더명에 종속성비율추가

@@ -114,7 +114,7 @@ torch.manual_seed(args.seed)
 
 now = datetime.datetime.now().strftime('%m%d_%H%M')
 if args.dependence_type == 1:
-    folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}'
+    folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}'
 else:
     folder = f'results/independence/{args.graph_dist}/{now}_{args.flow_type}_node{args.node_size}'
 

### 2023-11-23 train.py
train.py 수정(중요)
학습된 그래프와 실제 그래프 metric 계산 시, mixed graph를 사용하도록 처음 랜덤그래프 생성 후 그래프 변경
(dependency=1인 경우만)

@@ -137,7 +137,7 @@ G = generate_random_dag(d = args.node_size, degree=args.graph_degree, seed=args.
 
 # 2.2. Generate Data
 if args.dependence_type == 1:
-    X, cov, cov_prev = generate_linear_sem_correlated(G, args.data_sample_size, args.dependence_prop, args.seed, return_cov=True, x_dims=args.x_dims)
+    X, cov, cov_prev, G = generate_linear_sem_correlated(G, args.data_sample_size, args.dependence_prop, args.seed, return_cov=True, x_dims=args.x_dims, return_graph=True)
 else:
     X = generate_linear_sem(graph=G, n=args.data_sample_size, dist=args.graph_dist, linear_type=args.graph_linear_type, loc=args.graph_mean, scale=args.graph_scale, seed=args.seed, x_dims=args.x_dims)
 

### 2023-11-23 Tex/paper.tex
최종 페이퍼 가이드라인

\documentclass[10pt]{article}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{mathrsfs}
\usepackage{amssymb}
\usepackage[version=4]{mhchem}
\usepackage{stmaryrd}
\usepackage{bbold}
\usepackage{enumitem}
\usepackage{hanging} %reference indent
\usepackage[a4paper,left=25mm,right=25mm,top=40mm,bottom=40mm]{geometry}
\usepackage{sectsty}
\sectionfont{\fontsize{10}{15}\selectfont}
\subsectionfont{\fontsize{10}{15}}
\usepackage[mathlines,switch]{lineno}
\usepackage{amsthm}
\usepackage{kotex}
\usepackage{xcolor}
\usepackage{graphicx} % insert image
\usepackage{multicol} % multi column
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{float}
\usepackage{tabularx}
\usepackage{multirow} % multirow in table
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{hyperref}
\modulolinenumbers[5]
% \linenumbers

\usepackage{titlesec}
\titleformat{\section}
{\normalfont\fontsize{12}{12}\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}
{\normalfont\fontsize{11}{11}\bfseries}{\thesubsection}{1em}{}
\titlespacing{\paragraph}{0pt}{\parskip}{5pt}
\titleformat{\title}
{\normalfont\fontsize{13}{13}\bfseries}{\thetitle}{1em}{}

\usepackage{indentfirst} % first line indent
\setlength{\parindent}{10pt}

\renewcommand{\arraystretch}{1.1}

\title{\textbf{NFG-VAE: Learning Semi-Markovian DAGs with Flow based VAE}}

\author{Dangchan Kim, Byungguk Kang, Jaeseok Kim, Minchan Kim}

\begin{document}
\maketitle

\begin{abstract}
    We propose a method to learn the structure of a semi-Markovian directed acyclic graph, i.e. mixed graph with directed and undirected edges using a flow-based variational autoencoder. The proposed method is based on the assumption that the noise matrix in the linear structural equation model is a latent variable. In order to learn the structure of the mixed graph, we used an inverse autoregressive flow to approximate the dependency structure of the noise matrix and their prior distribution. We conducted experiments on simulated data adjusting the number of nodes and the proportion of bidirectional edges. The code is available at \url{https://github.io/ddangchani/NFG-VAE}.
\end{abstract}

\section{Introduction}

Here is the introduction.

\section{Directed Acyclic Graphs and Mixed Graphs}
\subsection{Directed Acyclic Graphs}

Here is the introduction of DAGs

\subsection{Mixed Graphs}

Here is the introduction of mixed graphs

\subsection{Graph Learning}

Here is the introduction of graph learning

\section{Variational Autoencoders and Normalizing Flows}

\subsection{Variational Autoencoders}

Here is the introduction of VAEs

\subsection{Normalizing Flows}

Here is the introduction of normalizing flows

\section{Method}

\subsection{Model}

Here is the introduction of our model

\subsection{Learning}

Here is the introduction (acyclicity constraint, etc.)

\section{Experiment}

Here is the introduction of our simulation study

\paragraph*{Datasets}

Here is the introduction of our datasets

\paragraph*{Evaluation}

Here is the introduction of our evaluation(SHD, etc.)

\subsection{Independent Noise Case}

Here is the introduction of our experiment on independent noise case

\subsection{Dependent Noise Case}

Here is the introduction of our experiment on dependent noise case

\section{Conclusion}

Here is the conclusion

% Biblography
\bibliographystyle{ieeetr}
\bibliography{references}

\end{document}

### 2023-11-23 model_new.py
model_new.py 삭제

@@ -1,302 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import torch.distributions as D
-from torch.autograd import Variable
-import torch.nn.utils.weight_norm as wn
-import math
-from utils import *
-import numpy as np
-
-
-from torch.autograd import Variable
-
-class Encoder(nn.Module):
-    """
-    Encoder class for VAE
-    """
-    def __init__(self, args, adj_A, tol=0.1):
-        super(Encoder, self).__init__()
-
-        self.args = args
-        self.adj_A = nn.Parameter(torch.from_numpy(adj_A).double(), requires_grad=True)
-        self.Wa = nn.Parameter(torch.zeros(args.z_size), requires_grad=True)
-        self.fc1 = nn.Linear(args.z_size, args.encoder_hidden, bias=True)
-        self.fc2 = nn.Linear(args.encoder_hidden, args.z_size, bias=True)
-        self.batch_size = args.batch_size
-
-        # for other loss
-        self.z = nn.Parameter(torch.tensor(tol))
-        self.z_positive = nn.Parameter(torch.ones_like(torch.from_numpy(adj_A)).double())
-
-        # For Flow layer
-        self.encoder_mean = nn.Linear(args.z_size, args.z_size)
-        self.encoder_logvar = nn.Linear(args.z_size, args.z_size)
-
-        self.init_weights()
-
-    def init_weights(self):
-        for m in self.modules():
-            if isinstance(m, nn.Linear):
-                nn.init.xavier_normal_(m.weight.data)
-            elif isinstance(m, nn.BatchNorm1d):
-                m.weight.data.fill_(1)
-                m.bias.data.zero_()
-
-    def forward(self, inputs):
-
-        if torch.sum(self.adj_A != self.adj_A):
-            print('nan error \n')
-        inputs = inputs.squeeze(-1)
-
-        adj_A1 = torch.sinh(3.*self.adj_A.float()) # amplify A
-
-        # adj_Aforz = $I-A^T$
-        adj_Aforz = (torch.eye(adj_A1.shape[0]).float() - (adj_A1.transpose(0,1)))
-        adj_Aforz = adj_Aforz.unsqueeze(0).repeat(self.batch_size, 1, 1)
-        adj_A = torch.eye(adj_A1.size()[0]).float()
-        h0 = F.relu((self.fc1(inputs.float()))) # first hidden layer
[…]

### 2023-11-24 NOTEARS_train.py
NOTEARS 알고리즘 파일 추가

@@ -1,0 +1,159 @@
+import time
+import datetime
+import pickle
+import os
+import argparse
+import torch
+import torch.nn.functional as F
+from torch.optim import lr_scheduler
+from torch.utils.data import Dataset, DataLoader
+from torch.utils.data.dataset import TensorDataset
+from torchviz import make_dot
+import math
+import numpy as np
+from utils import *
+from model import *
+from data import *
+from loss import *
+from tqdm import tqdm
+
+from NOTEARS import notears_linear
+
+# 1. Arguments(argparse)
+
+parser = argparse.ArgumentParser(description='NOTEARS Algorithm')
+
+# 1.1. Arguments for Data
+
+parser.add_argument('--data_sample_size', type=int, default=5000, 
+                    help='the number of samples of data')
+parser.add_argument('--node_size', type=int, default=10, 
+                    help='the number of nodes')
+parser.add_argument('--graph_degree', type=int, default=2, 
+                    help='the expected degree of random graph')
+parser.add_argument('--graph_dist', type=str, default='normal', 
+                    help='the distribution of random graph')
+parser.add_argument('--graph_scale', type=float, default=1.0, 
+                    help='the scale parameter of distribution')
+parser.add_argument('--graph_mean', type=float, default=0.0, 
+                    help='the mean parameter of distribution')
+parser.add_argument('--graph_linear_type', type=str, default='linear', 
+                    help='the type of linear graph')
+parser.add_argument('--dependence_type', type=int, default=0, 
+                    help='Dependent noise distribution or not')
+parser.add_argument('--dependence_prop', type=float, default=0.5, 
+                    help='the proportion of dependent noise')
+parser.add_argument('--seed', type=int, default=0,
+                    help='random seed')
+parser.add_argument('--x_dims', type=int, default=1,
+                    help='the dimension of each variable')
+
+
+# 1.2. Arguments for NOTEARS
+parser.add_argument('--lambda1', type=float, default=0.1,
[…]

### 2023-11-24 train.py
loss 계산 다시

@@ -149,8 +149,6 @@ pickle.dump(X, open(data_file, 'wb'))
 if args.dependence_type == 1:
     cov_file = os.path.join(folder, 'cov.pkl')
     pickle.dump(cov, open(cov_file, 'wb'))
-    cov_prev_file = os.path.join(folder, 'cov_prev.pkl')
-    pickle.dump(cov_prev, open(cov_prev_file, 'wb'))
 
 feat_train = torch.FloatTensor(X).to(device)
 feat_valid = torch.FloatTensor(X).to(device)
@@ -284,10 +282,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             print('nan error \n')
 
         # KL Divergence Loss
-        if args.flow_type == 'DAGGNN':
-            loss_kl = kl_gaussian_sem(logits) # DAG-GNN
-        else:
-            loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims) # VAE_VPFLOWS
+        loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
 
         # Reconstruction Loss
         loss_nll = calculate_reconstruction_loss(x_mean, data, x_logvar)

### 2023-11-24 train.py
LOSS 수정

@@ -281,7 +281,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             print('nan error \n')
 
         # KL Divergence Loss
-        loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
+        loss_kl = calculate_kl_prevent(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
 
         # Reconstruction Loss
         loss_nll = calculate_reconstruction_loss(x_mean, data, x_logvar)

### 2023-11-27 train.py
seed 폴더명에 반영

@@ -113,9 +113,9 @@ torch.manual_seed(args.seed)
 
 now = datetime.datetime.now().strftime('%m%d_%H%M')
 if args.dependence_type == 1:
-    folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}'
+    folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
 else:
-    folder = f'results/independence/{args.graph_dist}/{now}_{args.flow_type}_node{args.node_size}'
+    folder = f'results/independence/{args.graph_dist}/{now}_{args.flow_type}_node{args.node_size}_seed{args.seed}'
 
 if not os.path.exists(folder):
     os.makedirs(folder)

### 2023-11-27 Tex/paper.tex

@@
 
-\title{\textbf{NFG-VAE: Learning Semi-Markovian DAGs with Flow based VAE}}
+\title{
+\textbf{Learning Semi-Markovian DAGs with Flow based VAE}}
 
@@
 
-Here is the introduction of normalizing flows
+Normalizing flows are a method of transforming a simple distribution into a complex distribution.
+\cite{rezende15normalizingflows} Let $z$ be a random variable with a simple distribution $p(z)$, and $x$ be a random variable with a complex distribution $p(x)$.
+Then, we can transform $z$ into $x$ by a series of invertible transformations $f_k$ as follows:
+\begin{equation}
+    x = f_K \circ f_{K-1} \circ \cdots \circ f_1(z)
+\end{equation}
+
+The probability density function of $x$ can be obtained by the change of variables formula:
+\begin{equation}
+    p(x) = p(z) \left| \det \left( \frac{\partial f_K \circ f_{K-1} \circ \cdots \circ f_1(z)}{\partial z} \right) \right|
+\end{equation}
+
+There are several types of normalizing flows, such as Householder flow \cite{tomczak2017householder}, planar flow \cite{rezende15normalizingflows}, etc.
+In this paper, we use the inverse autoregressive flow (IAF) \cite{kingma2016iaf}, especially linear IAF.
+The linear IAF is defined as follows:
+\begin{equation}
+    \bf z_t = \bf \mu_t + \bf \sigma_t \odot \bf z_{t-1}.
+\end{equation}
+Linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance.
+The transformation is invertible and the determinant of the Jacobian is easily computable.
+To use linear IAF at VAE, we need to produce an extra output $\mathbf{L(x)}$ from the encoder network.
+The full-covariance Gaussian distribution is obtained by the following transformation:
+\begin{equation}
+    \mathbf{z}_T = \mathbf{L(x)} \cdot \mathbf{z}_0.
+\end{equation}
 
@@
 
+\begin{figure}
+    \centering
+    \includegraphics[width=0.8\textwidth]{fig/model.png}
+    \caption{Architecture of the proposed model.}
+    \label{diagram}
+\end{figure}
+
 \subsection{Learning}
@@
 
-Here is the introduction of our simulation study
+In this section, we conduct experiments on simulated data to evaluate the performance of our proposed method.
+We compare our method with the DAG-GNN \cite{yu2019daggnn} and NOTEARS \cite{zheng2018dags} with random graph datasets.
+We used a thresholding value of extracting graph as 0.3 which is the same as DAG-GNN and NOTEARS.
 
 \paragraph*{Datasets}
-
-Here is the introduction of our datasets
 

### 2023-11-28 train.py
train.py 오류 수정(윈도우 에러?)

@@ -16,6 +16,8 @@ from data import *
 from loss import *
 from tqdm import tqdm
 
+os.environ["KMP_DUPLICATE_LIB_OK"]="TRUE"
+
 # 1. Arguments(argparse)
 
 parser = argparse.ArgumentParser(description='PyTorch Training')

### 2023-11-28 train.py
dependence_type=1, node_size=20, seed=1 to 5, dependence_prop=0.5, flow_type=DAGGNN,IAF

@@ -283,7 +283,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             print('nan error \n')
 
         # KL Divergence Loss
-        loss_kl = calculate_kl_prevent(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
+        loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
 
         # Reconstruction Loss
         loss_nll = calculate_reconstruction_loss(x_mean, data, x_logvar)

### 2023-11-29 train.py
commit

@@ -283,7 +283,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             print('nan error \n')
 
         # KL Divergence Loss
-        loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
+        loss_kl = calculate_kl_prevent(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
 
         # Reconstruction Loss
         loss_nll = calculate_reconstruction_loss(x_mean, data, x_logvar)

### 2023-11-29 train.py
ccLinIAF 추가

@@ -93,9 +93,11 @@ parser.add_argument('--z_dims', type=int, default=1,
 parser.add_argument('--number_of_flows', type=int, default=5,
                     help='The number of HF flows: default 5.')
 parser.add_argument('--flow_type', type=str, default='IAF',
-                    help='The type of flows: "DAGGNN", "IAF", "HF"(Householder).')
+                    help='The type of flows: "DAGGNN", "IAF", "HF"(Householder), "ccIAF"')
 parser.add_argument('--lagrange', type=int, default=1,
                     help='Use lagrange multipliers or not.')
+parser.add_argument('--number_combination', type=int, default=3,
+                    help='The number of convex combinations: default 3.')
 
 args = parser.parse_args()
 args.z_size = args.node_size # the number of latent variables
@@ -187,6 +189,8 @@ elif args.flow_type == 'HF':
     vae = VAE_HF(args=args, adj_A=adj_A)
 elif args.flow_type == 'DAGGNN':
     vae = daggnn(args=args, adj_A=adj_A)
+elif args.flow_type == 'ccIAF':
+    vae = VAE_ccIAF(args=args, adj_A=adj_A)
 else:
     raise ValueError('Invalid flow type.')
 
@@ -268,7 +272,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
 
         # Forward VAE
         z = {}
-        if args.flow_type == 'IAF':
+        if args.flow_type == 'IAF' or args.flow_type == 'ccIAF':
             z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z['0'], z['1'], LT = model(data, rel_rec, rel_send)
         else:
             z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z['0'], z['1'] = model(data, rel_rec, rel_send)

### 2023-11-29 model.py
encoder_L 수정

@@ -52,6 +52,7 @@ class Encoder(nn.Module):
         self.fc1 = nn.Linear(args.x_dims, args.encoder_hidden, bias=True)
         self.fc2 = nn.Linear(args.encoder_hidden, args.z_dims, bias=True)
         self.batch_size = args.batch_size
+        self.dropout = nn.Dropout(p=args.encoder_dropout)
 
         # for other loss
         self.z = nn.Parameter(torch.tensor(tol))
@@ -83,6 +84,7 @@ class Encoder(nn.Module):
 
         adj_A = torch.eye(adj_A1.size()[0]).float()
         h0 = F.relu((self.fc1(inputs.float()))) # first hidden layer
+        h0 = self.dropout(h0)
         h1 = (self.fc2(h0)) # second hidden layer
         logits = torch.matmul(adj_Aforz, h1 + self.Wa) - self.Wa
 
@@ -90,7 +92,7 @@ class Encoder(nn.Module):
         z_q_mean = self.encoder_mean(logits)
         z_q_logvar = self.encoder_logvar(logits)
         
-        return z_q_mean, z_q_logvar, logits, adj_A1, adj_A, self.adj_A, self.z, self.z_positive, self.Wa
+        return z_q_mean, z_q_logvar, h0, adj_A1, adj_A, self.adj_A, self.z, self.z_positive, self.Wa
 
 
 class Decoder(nn.Module):
@@ -109,6 +111,7 @@ class Decoder(nn.Module):
 
         self.decoder_mean = nn.Linear(args.z_dims, args.z_dims)
         self.sigmoid = nn.Sigmoid()
+        self.dropout = nn.Dropout(p=args.decoder_dropout)
 
         self.init_weights()
 
@@ -130,6 +133,7 @@ class Decoder(nn.Module):
         mat_z = torch.matmul(adj_A_new_inv, input_z + Wa) - Wa
         
         H3 = F.relu((self.out_fc1(mat_z)))
+        H3 = self.dropout(H3)
         out = self.out_fc2(H3)
 
         # to Mean
@@ -163,7 +167,7 @@ class VAE_IAF(nn.Module):
         self.encoder = Encoder(args, adj_A=adj_A)
         self.decoder = Decoder(args)
         self.flow = FlowLayer(args)
-        self.encoder_L = nn.Linear(args.z_dims, args.z_size)
+        self.encoder_L = nn.Linear(args.encoder_hidden, args.z_size)
         self.softmax = nn.Softmax()
 
[…]

### 2023-11-29 train.py

@@ -64,9 +64,9 @@ parser.add_argument('--seed', type=int, default=42,
                     help='Random seed.')
 parser.add_argument('--epochs', type=int, default= 300,
                     help='Number of epochs to train.')
-parser.add_argument('--batch-size', type=int, default = 100, # note: should be divisible by sample size, otherwise throw an error
+parser.add_argument('--batch-size', type=int, default = 200, # note: should be divisible by sample size, otherwise throw an error
                     help='Number of samples per batch.')
-parser.add_argument('--lr', type=float, default=3e-3,  # basline rate = 1e-3
+parser.add_argument('--lr', type=float, default=1e-3,  # basline rate = 1e-3
                     help='Initial learning rate.')
 parser.add_argument('--encoder-hidden', type=int, default=64,
                     help='Number of hidden units.')
@@ -76,9 +76,9 @@ parser.add_argument('--temp', type=float, default=0.5,
                     help='Temperature for Gumbel softmax.')
 parser.add_argument('--k_max_iter', type = int, default = 100,
                     help ='the max iteration number for searching lambda and c')
-parser.add_argument('--encoder-dropout', type=float, default=0.0,
+parser.add_argument('--encoder_dropout', type=float, default=0.0,
                     help='Dropout rate (1 - keep probability).')
-parser.add_argument('--decoder-dropout', type=float, default=0.0,
+parser.add_argument('--decoder_dropout', type=float, default=0.0,
                     help='Dropout rate (1 - keep probability).')
 parser.add_argument('--lr-decay', type=int, default=200,
                     help='After how epochs to decay LR by a factor of gamma.')
@@ -287,7 +287,10 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             print('nan error \n')
 
         # KL Divergence Loss
-        loss_kl = calculate_kl_prevent(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
+        if args.flow_type == 'IAF' or args.flow_type == 'ccIAF':
+            loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
+        else:
+            loss_kl = calculate_kl_prevent(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
 
         # Reconstruction Loss
         loss_nll = calculate_reconstruction_loss(x_mean, data, x_logvar)

### 2023-11-29 model.py
commit

@@ -92,7 +92,7 @@ class Encoder(nn.Module):
         z_q_mean = self.encoder_mean(logits)
         z_q_logvar = self.encoder_logvar(logits)
         
-        return z_q_mean, z_q_logvar, h0, adj_A1, adj_A, self.adj_A, self.z, self.z_positive, self.Wa
+        return z_q_mean, z_q_logvar, logits, adj_A1, adj_A, self.adj_A, self.z, self.z_positive, self.Wa
 
 
 class Decoder(nn.Module):
@@ -167,7 +167,7 @@ class VAE_IAF(nn.Module):
         self.encoder = Encoder(args, adj_A=adj_A)
         self.decoder = Decoder(args)
         self.flow = FlowLayer(args)
-        self.encoder_L = nn.Linear(args.encoder_hidden, args.z_size)
+        self.encoder_L = nn.Linear(args.z_dims, args.z_size)
         self.softmax = nn.Softmax()
 
         for m in self.modules():

### 2023-11-29 train.py

@@ -98,6 +98,8 @@ parser.add_argument('--lagrange', type=int, default=1,
                     help='Use lagrange multipliers or not.')
 parser.add_argument('--number_combination', type=int, default=3,
                     help='The number of convex combinations: default 3.')
+parser.add_argument('--loss_prevent', type=int, default=0,
+                    help='Use loss that prevent overparametrization or not.')
 
 args = parser.parse_args()
 args.z_size = args.node_size # the number of latent variables
@@ -287,7 +289,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             print('nan error \n')
 
         # KL Divergence Loss
-        if args.flow_type == 'IAF' or args.flow_type == 'ccIAF':
+        if args.loss_prevent == 0:
             loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
         else:
             loss_kl = calculate_kl_prevent(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)

### 2023-11-29 train.py
commit

@@ -98,7 +98,7 @@ parser.add_argument('--lagrange', type=int, default=1,
                     help='Use lagrange multipliers or not.')
 parser.add_argument('--number_combination', type=int, default=3,
                     help='The number of convex combinations: default 3.')
-parser.add_argument('--loss_prevent', type=int, default=0,
+parser.add_argument('--loss_prevent', type=int, default=1,
                     help='Use loss that prevent overparametrization or not.')
 
 args = parser.parse_args()

### 2023-11-29 Tex/paper.tex
dependece_type=1, node_size=30, seed=1to2, flow_type=IAF,DAGGNN, dependence_prop=0.5, k_max_iter=50, lagrange=0, tau_A=0.1

@@
 
-\author{Dangchan Kim, Byungguk Kang, Jaeseok Kim, Minchan Kim}
+\author{\normalsize\bfseries Dangchan Kim, Byungguk Kang, Jaeseok Kim, Minchan Kim}
 
@@
 
-Here is the introduction.
+Directed acyclic graphs(DAGs) are commonly used to represent causal relationships between variables.
+However in many cases, recovering the causal structure from observational data is NP-hard problem.
+\cite{ChickeringNPhard} To overcome this problem, several methods have been proposed.
+NOTEARS \cite{zheng2018dags} is a method that learns the structure of DAGs by minimizing the continuous approximation of the discrete constraint.
+This method has made it possible to learn DAGs with continuous optimization problem.
+
+Based on the continuous acyclicty constraint, neural network based approach have been proposed.
+DAG-GNN \cite{yu2019daggnn} is a method that learns the structure of DAGs by using a neural network that takes the adjacency matrix of the graph as an additional input.
+Since the noise variable in the linear structural equation model can be interpreted as a latent variable, it uses a variational autoencoder(VAE) to learn DAGs.
 
@@
 \end{equation}
-Linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance.
+The linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance.
 The transformation is invertible and the determinant of the Jacobian is easily computable.
@@
 \end{equation}
+Note if we restrict the $\mathbf{L(x)}$ to be a lower triangular matrix with diagonal elements of ones, then the log-determinant of the Jacobian is zero, which is so-called volume preserving normalizing flow.
+\cite{kingma2016iaf}\\
+
+With the linear IAF, the KL loss term can be written in closed form as follows:
+\begin{equation}
+    \begin{aligned}
+        D_{KL}(q_\phi(\mathbf{z}_0 | \mathbf{x}) \Vert p(\mathbf{z}_T)) &=
+        \log q_\phi(\mathbf{z}_0 | \mathbf{x}) - \log p(\mathbf{z}_T) \\
+        &= -\frac{1}{2}(\mathbf{z}_0 -\mu_\phi)^T\Sigma_\phi^{-1}(\mathbf{z}_0 -\mu_\phi) + \frac{1}{2}\mathbf{z}_T^T\mathbf{z}_T \\
+    \end{aligned}
+\end{equation}
+where $\mu_\phi$ and $\Sigma_\phi$ are the mean vector and covariance matrix from the encoder network.\\
 
@@
 In this section, we conduct experiments on simulated data to evaluate the performance of our proposed method.
-We compare our method with the DAG-GNN \cite{yu2019daggnn} and NOTEARS \cite{zheng2018dags} with random graph datasets.
-We used a thresholding value of extracting graph as 0.3 which is the same as DAG-GNN and NOTEARS.
+We compare our method with the DAG-GNN \cite{yu2019daggnn} with random graph datasets.
+We used a thresholding value of extracting graph as 0.3 which is the same value used at DAG-GNN and NOTEARS.\\
 
-\paragraph*{Datasets}
+\paragraph*{Datasets} Random graph datasets are generated by the following procedure.
+First, we generate a random directed acyclic graph with Erdős–Rényi model.
+In Section 5.1 we conduct experiment where the noise matrix follows multivariate Gaussian with non-diagonal covariance matrix.
+We randomly choose edges of given proportion and make them bidirectional.
+Then, we generate a corresponding random covariance matrix and generate multivariate Gaussian data with the covariance matrix.
+We generate 5000 samples for each graph.
+For the size of the graph, we used 10, 20, 30, 50 nodes.
+For the proportion of bidirectional edges, we used 0, 0.1, 0.3, 0.5, 0.8.
+In Section 5.2, we considered the case where the noise matrix follows not only independent Gaussian but also independent Laplace and exponential.\\
 
-\paragraph*{Evaluation}
-
-Here is the introduction of our evaluation(SHD, etc.)
+\paragraph*{Evaluation} We evaluate the performance of our method with SHD(Structural Hamming Distance) and FDR(False Discovery Rate).
+SHD is the number of edge additions, deletions, and reversals needed to transform the estimated graph into the true graph.
+FDR is the ratio of false positives to the total number of positives.
+The metrics such as SHD and FDR are calculated by comparing the estimated graph with the true graph with bidirectional edges.
+For each combination of the number of nodes and the proportion of bidirectional edges, we generate at least 5 random graphs and calculate the average metrics.\\
 

### 2023-11-29 Tex/paper.tex
dependece_type=1, node_size=30, seed=3to5, flow_type=IAF,DAGGNN, dependence_prop=0.5, k_max_iter=50, lagrange=0, tau_A=0.1

@@
 \title{
-\textbf{Learning Semi-Markovian DAGs with Flow based VAE}}
+\textbf{Learning semi-Markovian DAGs with flow-based VAE}}
 
@@
 This method has made it possible to learn DAGs with continuous optimization problem.
+\\
 
@@
 DAG-GNN \cite{yu2019daggnn} is a method that learns the structure of DAGs by using a neural network that takes the adjacency matrix of the graph as an additional input.
-Since the noise variable in the linear structural equation model can be interpreted as a latent variable, it uses a variational autoencoder(VAE) to learn DAGs.
+Since the noise variable in the linear structural equation model can be interpreted as a latent variable, it uses a variational autoencoder(VAE) to learn DAGs.\\
+
+However, there are some cases where the noise matrix is not independent.
+For example, in the case of a linear structural equation model with correlated hidden variables, the noise matrix might have a dependency structure.
+In this case, the edges having correlation in the noise variable are considered to be bidirectional edges.
+Such models are called semi-Markovian.
+\cite{Shipster2006semimarkovian} Since the prior distribution of the noise variable is standard Gaussian at DAG-GNN, it is difficult to learn the structure of the semi-Markovian DAGs.
+Therefore, we propose a method to learn the full covariance matrix of the noise variable using a flow-based VAE.\\
 
@@
 
+At VAE, typically the prior distribution of the latent variable $\bf z$ is assumed to be a standard Gaussian distribution, or a multivariate Gaussian distribution with diagonal covariance matrix.
+To have a more flexible prior distribution, we can use normalizing flows to transform the standard Gaussian distribution to a more complex distribution.\\
+
 There are several types of normalizing flows, such as Householder flow \cite{tomczak2017householder}, planar flow \cite{rezende15normalizingflows}, etc.
@@
 The linear IAF is defined as follows:
+
 \begin{equation}
@@
 \end{equation}
+
 The linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance.
@@
 The full-covariance Gaussian distribution is obtained by the following transformation:
+
 \begin{equation}
@@
 \end{equation}
+
 Note if we restrict the $\mathbf{L(x)}$ to be a lower triangular matrix with diagonal elements of ones, then the log-determinant of the Jacobian is zero, which is so-called volume preserving normalizing flow.
@@
 \end{equation}
-where $\mu_\phi$ and $\Sigma_\phi$ are the mean vector and covariance matrix from the encoder network.\\
+where $\mu_\phi$ and $\Sigma_\phi$ are the mean vector and covariance matrix obtained from the encoder network.\\
 

### 2023-11-29 train.py

@@ -208,7 +208,7 @@ def _h_A(A, m):
     h_A = torch.trace(expm_A) - m
     return h_A
 
-prox_plus = torch.nn.Threshold(0.,0.)
+prox_plus = torch.nn.Threshold(0.,0.) # ReLU function
 
 def stau(w, tau):
     w1 = prox_plus(torch.abs(w)-tau)
@@ -289,16 +289,16 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             print('nan error \n')
 
         # KL Divergence Loss
-        if args.loss_prevent == 0:
-            loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
-        else:
+        if args.loss_prevent:
             loss_kl = calculate_kl_prevent(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
+        else:
+            loss_kl = calculate_kl_loss(z['0'], z['1'], z_q_mean, z_q_logvar, args.z_dims)
 
         # Reconstruction Loss
         loss_nll = calculate_reconstruction_loss(x_mean, data, x_logvar)
 
-        # add A loss
-        sparse_loss = args.tau_A * torch.sum(torch.abs(origin_A))
+        # Sparsity loss
+        sparse_loss = args.tau_A * torch.sum(torch.abs(origin_A)) # L1 norm
 
         loss = loss_nll + loss_kl + sparse_loss
         
@@ -351,41 +351,18 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
             'shd_trian: {:.10f}'.format(np.mean(shd_train)),
             'time: {:.4f}s'.format(time.time() - t))
     
-        if np.mean(nll_val) < best_val_loss:
-            torch.save(model.state_dict(), model_file)
-            print('Best model so far, saving...')
-            print('Epoch: {:04d}'.format(epoch),
-                'nll_train: {:.10f}'.format(np.mean(nll_train)),
-                'kl_train: {:.10f}'.format(np.mean(kl_train)),
-                'ELBO_loss: {:.10f}'.format(np.mean(kl_train)  + np.mean(nll_train)),
-                'mse_train: {:.10f}'.format(np.mean(mse_train)),
[…]

### 2023-11-29 train.py
dependece_type=1, node_size=30, seed=1to5, flow_type=IAF, dependence_prop=0.5, k_max_iter=50, lagrange=0, tau_A=0.1

@@ -355,8 +355,8 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
         to_print = {'nll_train' : '{:.4f}'.format(np.mean(nll_train)),
                     'kl_train' : '{:.4f}'.format(np.mean(kl_train)),
                     'ELBO_loss' : '{:.4f}'.format(np.mean(kl_train)  + np.mean(nll_train)),
-                    'shd_trian' : '{:0f}'.format(np.mean(shd_train)),
-                    'nnz': '{:0f}'.format(nnz)}
+                    'shd_trian' : '{:.0f}'.format(np.mean(shd_train)),
+                    'nnz': '{:.0f}'.format(nnz)}
 
         if args.lagrange:
             to_print['h(A)'] = '{:.4f}'.format(h_A.item())

### 2023-11-29 train.py
dependece_type=0, node_size=30, seed=1, flow_type=IAF, k_max_iter=50, lagrange=1

@@ -321,7 +321,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
         loss.backward()
         loss = optimizer.step()
 
-        origin_A.data = stau(origin_A.data, args.tau_A*args.lr)
+        origin_A.data = stau(origin_A.data, args.tau_A*args.lr) # update A
 
         if torch.sum(origin_A != origin_A):
             print('nan error\n')
@@ -401,6 +401,7 @@ if args.lagrange:
     k_max_iter = int(args.k_max_iter)
 else:
     k_max_iter = 1
+k_max_iter = int(args.k_max_iter)
 h_A_old = np.inf
 
 pbar = tqdm(range(args.epochs * k_max_iter), desc='Training')

### 2023-11-30 train.py
Lagrange 안쓸때 폴더명 수정

@@ -119,7 +119,10 @@ torch.manual_seed(args.seed)
 
 now = datetime.datetime.now().strftime('%m%d_%H%M')
 if args.dependence_type == 1:
-    folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
+    if args.lagrange:
+        folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
+    else:
+        folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}_noL'
 else:
     folder = f'results/independence/{args.graph_dist}/{now}_{args.flow_type}_node{args.node_size}_seed{args.seed}'
 

### 2023-11-30 train.py
폴더명 구체적으로 수정

@@ -118,27 +118,40 @@ torch.manual_seed(args.seed)
 # Folder to save the results, models, dataset
 
 now = datetime.datetime.now().strftime('%m%d_%H%M')
-if args.dependence_type == 1:
-    if args.lagrange:
-        folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
-    else:
-        folder = f'results/dependence/{now}_{args.flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}_noL'
+
+if args.flow_type in ['IAF', 'DAGGNN']:
+    flow_type = args.flow_type
+elif args.flow_type == 'HF':
+    flow_type = f'{args.number_of_flows}{args.flow_type}'
+elif args.flow_type == 'ccIAF':
+    flow_type = f'{args.number_combination}{args.flow_type}'
 else:
-    folder = f'results/independence/{args.graph_dist}/{now}_{args.flow_type}_node{args.node_size}_seed{args.seed}'
+    raise ValueError(f'Invalid flow type: {args.flow_type}.')
+
+if args.dependence_type == 1:
+    folder = f'results/dependence/{now}_{flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
+else:
+    folder = f'results/independence/{args.graph_dist}/{now}_{flow_type}_node{args.node_size}_seed{args.seed}'
+
+if not args.lagrange:
+    folder += '_noL'
 
 if not os.path.exists(folder):
     os.makedirs(folder)
 
 # Save the arguments
-meta_file = os.path.join(folder, 'meta.pkl')
+meta_file = os.path.join(folder, 'meta.txt')
 model_file = os.path.join(folder, 'model.pt')
 
 log_file = os.path.join(folder, 'log.txt')
 log = open(log_file, 'w')
 
-pickle.dump(args, open(meta_file, 'wb'))
+# pickle.dump(args, open(meta_file, 'wb'))
+with open(meta_file, 'w') as f:
+    for arg in vars(args):
+        print(arg, getattr(args, arg), file=f)
 
-# 2. Load Data
+
 
[…]

### 2023-11-30 train.py
run node_size=50 IAF dependence_type=1 dependence_prop=0.5 k_max_iter=50 lagrange=1 tau_A=0.1

@@ -460,7 +460,7 @@ try:
         h_A_old = h_A_new.item()
         lambda_A += c_A * h_A_new.item()
 
-        if h_A_new.item() <= h_tol:
+        if h_A_new.item() <= h_tol: # when h(A) is close enough to 0
             break
 
     print("Best Epoch: {:04d}".format(best_epoch), file=log)

### 2023-11-30 model.py
optimization test run

@@ -37,7 +37,7 @@ class FlowLayer(nn.Module):
         LT = torch.mul(L_matrix, LTmask) + I # Lower triangular batches
         z_new = torch.bmm(LT, z) # B x L x L * B x L x 1 = B x L x 1
 
-        return z_new
+        return z_new, LT
 
 class Encoder(nn.Module):
     """
@@ -91,6 +91,11 @@ class Encoder(nn.Module):
         # For Flow layer
         z_q_mean = self.encoder_mean(logits)
         z_q_logvar = self.encoder_logvar(logits)
+        
+        if self.args.logits:
+            logits = logits # B * L * 1
+        else:
+            logits = h0 # B * L * 64
         
         return z_q_mean, z_q_logvar, logits, adj_A1, adj_A, self.adj_A, self.z, self.z_positive, self.Wa
 
@@ -167,7 +172,11 @@ class VAE_IAF(nn.Module):
         self.encoder = Encoder(args, adj_A=adj_A)
         self.decoder = Decoder(args)
         self.flow = FlowLayer(args)
-        self.encoder_L = nn.Linear(args.z_dims, args.z_size)
+        if self.args.logits:
+            self.encoder_L = nn.Linear(args.z_dims, args.z_size)
+        else:
+            self.encoder_L = nn.Linear(args.encoder_hidden, args.z_size)
+        
         self.softmax = nn.Softmax()
 
         for m in self.modules():
@@ -197,12 +206,12 @@ class VAE_IAF(nn.Module):
 
         # Flow layer
         L = self.encoder_L(logits)
-        z['1'] = self.flow(L, z['0']) # z1 : after Flow
+        z['1'], LT = self.flow(L, z['0']) # z1 : after Flow
 
         # z ~ p(x|z) : decoder
         mat_z, out, x_mean, x_logvar = self.decoder(z['1'], origin_A, Wa)
 
-        return z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, out, x_mean, x_logvar, z['0'], z['1'], L
+        return z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, out, x_mean, x_logvar, z['0'], z['1'], LT
 
 class daggnn(nn.Module):
     def __init__(self, args, adj_A):

### 2023-11-30 train.py

@@ -100,10 +100,11 @@ parser.add_argument('--number_combination', type=int, default=3,
                     help='The number of convex combinations: default 3.')
 parser.add_argument('--loss_prevent', type=int, default=1,
                     help='Use loss that prevent overparametrization or not.')
+parser.add_argument('--logits', type=int, default=1,
+                    help='Link encoder_L after 2 hidden layers(1) or 1 layer(0).')
 
 args = parser.parse_args()
 args.z_size = args.node_size # the number of latent variables
-print(args)
 
 # Device configuration (GPU or MPS or CPU)
 
@@ -146,12 +147,6 @@ model_file = os.path.join(folder, 'model.pt')
 log_file = os.path.join(folder, 'log.txt')
 log = open(log_file, 'w')
 
-# pickle.dump(args, open(meta_file, 'wb'))
-with open(meta_file, 'w') as f:
-    for arg in vars(args):
-        print(arg, getattr(args, arg), file=f)
-
-
 
 # 2.1. Generate DAG
 G = generate_random_dag(d = args.node_size, degree=args.graph_degree, seed=args.seed)
@@ -168,8 +163,9 @@ pickle.dump(X, open(data_file, 'wb'))
 
 # save covariance matrix to file
 if args.dependence_type == 1:
-    cov_file = os.path.join(folder, 'cov.pkl')
-    pickle.dump(cov, open(cov_file, 'wb'))
+    cov_file = open(folder + '/cov.txt', 'wb')
+    np.savetxt(cov_file, cov, fmt='%.3f')
+    cov_file.closed
 
 feat_train = torch.FloatTensor(X).to(device)
 feat_valid = torch.FloatTensor(X).to(device)
@@ -295,12 +291,6 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
         else:
             z_q_mean, z_q_logvar, logits, origin_A, adj_A_tilt, myA, z_gap, z_positive, Wa, mat_z, output, x_mean, x_logvar, z['0'], z['1'] = model(data, rel_rec, rel_send)
 
-        """
-        in DAG-GNN
-        enc_x, logits, origin_A, adj_A_tilt_encoder, z_gap, z_positive, myA, Wa = encoder(data, rel_rec, rel_send)  # logits is of size: [num_sims, z_dims]
[…]

### 2023-11-30 train.py
optimization test run

@@ -15,6 +15,7 @@ from model import *
 from data import *
 from loss import *
 from tqdm import tqdm
+import pandas as pd
 
 os.environ["KMP_DUPLICATE_LIB_OK"]="TRUE"
 
@@ -409,6 +410,10 @@ h_A_old = np.inf
 
 pbar = tqdm(range(args.epochs * k_max_iter), desc='Training')
 
+shd_curve = []
+nnd_curve = []
+fdr_curve = []
+
 print(args)
 
 with open(meta_file, 'w') as f:
@@ -435,8 +440,11 @@ try:
                     best_epoch = epoch
                     best_MSE_graph = graph
 
-            # print("Optimization Finished!")
-            # print("Best Epoch: {:04d}".format(best_epoch))
+            fdr, tpr, fpr, shd, nnz = count_accuracy(G, nx.DiGraph(graph))
+            shd_curve.append(shd)
+            nnd_curve.append(nnz)
+            fdr_curve.append(fdr)
+            # early stopping
             if ELBO_loss > 2 * best_ELBO_loss:
                 break
 
@@ -547,3 +555,7 @@ print("Total time elapsed: {:.4f}s".format(time.time() - t_total), file=log)
 if log is not None:
     print(folder)
     log.close()
+
+# Save curves
+df = pd.DataFrame({'SHD': shd_curve, 'predicted edges': nnd_curve, 'FDR': fdr_curve})
+df.to_csv(folder + '/metrics.csv', index=True)

### 2023-12-01 train.py
DAG도 저장하도록 수정

@@ -152,6 +152,8 @@ log = open(log_file, 'w')
 # 2.1. Generate DAG
 G = generate_random_dag(d = args.node_size, degree=args.graph_degree, seed=args.seed)
 
+G_DAG = nx.to_numpy_array(G)
+
 # 2.2. Generate Data
 if args.dependence_type == 1:
     X, cov, cov_prev, G = generate_linear_sem_correlated(G, args.data_sample_size, args.dependence_prop, args.seed, return_cov=True, x_dims=args.x_dims, return_graph=True)
@@ -536,6 +538,12 @@ matG2 = np.matrix(best_ELBO_graph)
 for line in matG2:
     np.savetxt(f2, line, fmt='%.5f')
 f2.closed
+
+f3 = open(folder + '/trueG_DAG.txt', 'w')
+matG3 = np.matrix(G_DAG)
+for line in matG3:
+    np.savetxt(f3, line, fmt='%.5f')
+f3.closed
 
 # LT to pickle
 if args.flow_type == 'IAF':

### 2023-12-01 Tex/paper.tex
optimization test run DAGGNN

--- 2023-11-27 Tex/paper.tex
@@
 \title{
-\textbf{Learning Semi-Markovian DAGs with Flow based VAE}}
+\textbf{Learning semi-Markovian DAGs with flow-based VAE}}
 
-\author{Dangchan Kim, Byungguk Kang, Jaeseok Kim, Minchan Kim}
+\author{\normalsize\bfseries Dangchan Kim, Byungguk Kang, Jaeseok Kim, Minchan Kim}
 
@@
 \begin{abstract}
-    We propose a method to learn the structure of a semi-Markovian directed acyclic graph, i.e.
-mixed graph with directed and undirected edges using a flow-based variational autoencoder.
-The proposed method is based on the assumption that the noise matrix in the linear structural equation model is a latent variable.
-In order to learn the structure of the mixed graph, we used an inverse autoregressive flow to approximate the dependency structure of the noise matrix and their prior distribution.
-We conducted experiments on simulated data adjusting the number of nodes and the proportion of bidirectional edges.
+    We propose a method to learn the structure of a semi-Markovian directed acyclic graph, which is a mixed graph containing both directed and bidirectional edges, using a flow-based variational autoencoder.
+The proposed method is based on the assumption that noise variables in the linear structural equation model can be considered as latent variables.
+To learn the structure of the mixed graph, we employ an inverse autoregressive flow to approximate the dependency structure of the noise variables and its prior distribution.
+We conducted experiments on simulated data by adjusting the number of nodes and the proportion of bidirectional edges.
 The code is available at \url{https://github.io/ddangchani/NFG-VAE}.
@@
 
-Here is the introduction.
+Directed acyclic graphs (DAGs) are commonly used to represent causal relationships between variables \cite{pearl_2009}.
+However, recovering the causal structure from observational data is an NP-hard problem \cite{ChickeringNPhard}.
+To overcome this problem, several methods have been proposed.
+NOTEARS \cite{zheng2018dags} is a method that learns the structure of DAGs by minimizing the continuous approximation of the discrete constraint, enabling the learning of DAGs through continuous optimization.\\
+
+Based on the continuous acyclicity constraint neural network-based approaches known as graph neural networks, have been proposed.
+\cite{yu2019daggnn,li2020dirichlet,kipf2016variational,zhang2019dvae} The most representative model would be DAG-GNN \cite{yu2019daggnn}, a method that learns the structure of DAGs by using a neural network that takes the adjacency matrix of the graph as an additional input.
+It utilizes a variational autoencoder (VAE) to learn DAGs by interpreting the noise variable in the linear structural equation model as a latent variable.\\
+
+However, in some cases, the noise variables may not be independent.
+For example, in a linear structural equation model with correlated hidden variables, the noise variables may exhibit a dependency structure.
+In such cases, edges with correlation in the noise variable are considered bidirectional edges, and these models are referred to as semi-Markovian \cite{Shipster2006semimarkovian}.
+Since the prior distribution of the noise variable is assumed to be a standard Gaussian in DAG-GNN or other graph neural networks, it becomes challenging to learn the structure of semi-Markovian DAGs.
+Therefore, we propose a method to learn the full covariance matrix of the noise variable using a flow-based VAE.
 
@@
 
-Normalizing flows are a method of transforming a simple distribution into a complex distribution.
-\cite{rezende15normalizingflows} Let $z$ be a random variable with a simple distribution $p(z)$, and $x$ be a random variable with a complex distribution $p(x)$.
-Then, we can transform $z$ into $x$ by a series of invertible transformations $f_k$ as follows:
+Normalizing flows are a method of transforming a simple distribution into a complex distribution \cite{rezende15normalizingflows}.
+Let $z$ be a random variable with a simple distribution $p(z)$, and $x$ be a random variable with a complex distribution $p(x)$.
+We can transform $z$ into $x$ using a series of invertible transformations $\{f_k\}_{k=1,\ldots,T}$ as follows:
 \begin{equation}
-    x = f_K \circ f_{K-1} \circ \cdots \circ f_1(z)
+    x = f_T \circ f_{T-1} \circ \cdots \circ f_1(z)
 \end{equation}
@@
 \begin{equation}
-    p(x) = p(z) \left| \det \left( \frac{\partial f_K \circ f_{K-1} \circ \cdots \circ f_1(z)}{\partial z} \right) \right|
+    p(x) = p(z) \left| \det \left( \frac{\partial f_T \circ f_{T-1} \circ \cdots \circ f_1(z)}{\partial z} \right) \right|
 \end{equation}
+
+In VAE, the prior distribution of the latent variable $\mathbf{z}$ is typically assumed to be a standard Gaussian distribution or a multivariate Gaussian distribution with a diagonal covariance matrix.
+\cite{kingma2013auto} However, to employ a more flexible prior distribution, we can utilize normalizing flows to transform the standard Gaussian distribution into a more complex distribution.
+This allows us to capture richer and more intricate patterns in the latent space, such as correlation between latent variables.\\
 
@@
 \end{equation}
-Linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance.
+
+The linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance.
 The transformation is invertible and the determinant of the Jacobian is easily computable.
@@
 \end{equation}
+
+Note if we restrict the $\mathbf{L(x)}$ to be a lower triangular matrix with diagonal elements of ones, then the log-determinant of the Jacobian is zero, which is so-called volume preserving normalizing flow.
+\cite{kingma2016iaf}\\
+
+With the linear IAF, the KL loss term can be written in closed form as follows:
+\begin{equation}
+    \begin{aligned}
+        D_{KL}(q_\phi(\mathbf{z}_0 | \mathbf{x}) \Vert p(\mathbf{z}_T)) &=
+        \log q_\phi(\mathbf{z}_0 | \mathbf{x}) - \log p(\mathbf{z}_T) \\
+        &= -\frac{1}{2}(\mathbf{z}_0 -\mu_\phi)^T\Sigma_\phi^{-1}(\mathbf{z}_0 -\mu_\phi) + \frac{1}{2}\mathbf{z}_T^T\mathbf{z}_T \\
+    \end{aligned}
+\end{equation}
+
+where $\mu_\phi$ and $\Sigma_\phi$ are the mean vector and covariance matrix obtained from the encoder network.\\
 
@@
 
-Here is the introduction of our model
+The overall architecture of our model is shown in Figure \ref*{diagram}.
+The encoder network takes the input data matrix $\mathbf{X}\in \mathbb{R}^{B \times m \times 1}$ and adjacency matrix $\mathbf{A}\in \mathbb{R}^{m \times m}$ as inputs and outputs the mean vector $\mu_\phi$ and (log) variance $\sigma_\phi$ of the latent variable $\mathbf{z}_0$.
+The decoder network takes the latent variable $\mathbf{z}_T$ and the same adjacency matrix $\mathbf{A}$ used at encoder as inputs and outputs the mean vector $\mu_\theta$ and covariance matrix $\Sigma_\theta$ of the reconstructed data matrix $\mathbf{X}$.
+Also, during the training process the decoder variance $\sigma_\theta$ is fixed to 1.
+For the size of the hidden layer, we use 64.
 
@@
 
-Here is the introduction (acyclicity constraint, etc.)
+Our model aims to learn the adjacency matrix of the directed acyclic graph.
+Instead of regarding the mixed graph as a cyclic graph, we consider them as DAG structure with correlated noise variables.
+Thus we can use the same optimization procedure as DAG-GNN, which uses the acyclicity constraint as a continuous approximation of the discrete constraint.
+The optimization procedure is minimizing the following loss function:
+
+\begin{equation}
+    \mathcal{L}(A, W, \lambda) =-\mathcal{L}_{\mathrm{ELBO}} + \tau_A \Vert A \Vert_1 + \lambda h(A) + \frac{c}{2} |h(A)|^2.
+\end{equation}
+
+The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
+The third and fourth terms are the augmented Lagrangian terms.
+Since larger value of $c$ makes the acyclicity constraint to zero, the training process is done by gradually increasing the value of $c$.
+\cite{yu2019daggnn}\\
 
@@
 In this section, we conduct experiments on simulated data to evaluate the performance of our proposed method.
-We compare our method with the DAG-GNN \cite{yu2019daggnn} and NOTEARS \cite{zheng2018dags} with random graph datasets.
-We used a thresholding value of extracting graph as 0.3 which is the same as DAG-GNN and NOTEARS.
+We compare our method with the DAG-GNN \cite{yu2019daggnn} with random graph datasets.
+We used a thresholding value of extracting graph as 0.3 which is the same value used at DAG-GNN and NOTEARS.\\
 
-\paragraph*{Datasets}
+\paragraph*{Datasets} Random graph datasets are generated using the following procedure.
+First, we generate a random directed acyclic graph using the Erdős–Rényi model.
+In Section 5.1, we conduct experiments where the noise variables follows a multivariate Gaussian distribution with a non-diagonal covariance matrix.
+We randomly select a given proportion of edges and make them bidirectional.
+Then, we generate a corresponding random covariance matrix and generate multivariate Gaussian data using that covariance matrix.
+We generate 5000 samples for each graph.
+For the size of the graph, we use 10, 20, 30, and 50 nodes.
+For the proportion of bidirectional edges, we use 0, 0.1, 0.3, 0.5, and 0.8.
+In Section 5.2, we consider the case where the noise variables follow not only an independent Gaussian distribution but also independent Laplace and exponential distributions.\\
 
-\paragraph*{Evaluation}
-
-Here is the introduction of our evaluation(SHD, etc.)
+\paragraph*{Evaluation} We evaluate the performance of our method using two metrics: Structural Hamming Distance (SHD) and False Discovery Rate (FDR).
+SHD measures the number of edge additions, deletions, and reversals required to transform the estimated graph into the true graph.
+FDR represents the ratio of false positives to the total number of predicted edges.
+These metrics are calculated by comparing the estimated graph with the true graph, considering bidirectional edges.
+For each combination of the number of nodes and the proportion of bidirectional edges, we generate at least 5 random graphs and calculate the average metrics.
 

### 2023-12-02 train.py
l1 대신 edge 출력하도록 변경

@@ -366,7 +366,7 @@ def train(epoch, model, best_val_loss, G, lambda_A, c_A, optimizer, pbar=None):
                     'kl_train' : '{:.4f}'.format(np.mean(kl_train)),
                     'ELBO_loss' : '{:.4f}'.format(np.mean(kl_train)  + np.mean(nll_train)),
                     'shd_trian' : '{:.0f}'.format(np.mean(shd_train)),
-                    'l1': '{:.1f}'.format(l1_norm.item())}
+                    'edge': '{:.0f}'.format(nnz)}
 
         if args.lagrange:
             to_print['h(A)'] = '{:.4f}'.format(h_A.item())

### 2023-12-02 Tex/paper.tex
run DAGGNN vs IAF 20 nodes

@@
 Let $z$ be a random variable with a simple distribution $p(z)$, and $x$ be a random variable with a complex distribution $p(x)$.
-We can transform $z$ into $x$ using a series of invertible transformations $\{f_k\}_{k=1,\ldots,T}$ as follows:
+We can transform $z_0$ into $z_T$ using a series of invertible transformations $\{f_k\}_{k=1,\ldots,T}$ as follows:
 \begin{equation}
-    x = f_T \circ f_{T-1} \circ \cdots \circ f_1(z)
+    z_T = f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)
 \end{equation}
@@
 \begin{equation}
-    p(x) = p(z) \left| \det \left( \frac{\partial f_T \circ f_{T-1} \circ \cdots \circ f_1(z)}{\partial z} \right) \right|
+    p(z_T) = p(z_0) \left| \det \left( \frac{\partial f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)}{\partial z_0} \right) \right|
 \end{equation}
@@
 
-The linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance.
+The linear IAF is the simplest case of IAF, which transforms a multivariate Gaussian with diagonal covariance to a multivariate Gaussian with full covariance using a single flow layer.
 The transformation is invertible and the determinant of the Jacobian is easily computable.

### 2023-12-02 Tex/paper.tex
tex

@@
 
-\begin{figure}
+\begin{figure}[H]
     \centering
@@
 First, we generate a random directed acyclic graph using the Erdős–Rényi model.
-In Section 5.1, we conduct experiments where the noise variables follows a multivariate Gaussian distribution with a non-diagonal covariance matrix.
+In Section 5.1, we consider the case where the noise variables follow not only an independent Gaussian distribution but also independent Laplace and exponential distributions.
+In Section 5.2, we conduct experiments where the noise variables follows a multivariate Gaussian distribution with a non-diagonal covariance matrix.
 We randomly select a given proportion of edges and make them bidirectional.
 Then, we generate a corresponding random covariance matrix and generate multivariate Gaussian data using that covariance matrix.
-We generate 5000 samples for each graph.
+In both sections, we generate 5000 samples for each graph.
 For the size of the graph, we use 10, 20, 30, and 50 nodes.
-For the proportion of bidirectional edges, we use 0, 0.1, 0.3, 0.5, and 0.8.
-In Section 5.2, we consider the case where the noise variables follow not only an independent Gaussian distribution but also independent Laplace and exponential distributions.\\
+For the proportion of bidirectional edges, we use 0.1, 0.3, 0.5, and 0.8.
+\\
 
@@
 
-Here is the introduction of our experiment on dependent noise case
+First, when the proportion of bidirectional edges is fixed to 0.5, we compare the performance of our method with DAG-GNN.
+The results are shown in Figure \ref*{SHD_FDR_prop50}.
+Our method outperforms DAG-GNN in terms of both SHD and FDR.
+Also, at the node size of 20, our method has shown stable training, which results a huge difference in SHD and FDR.
+The example of estimated graphs of each method is shown in Figure \ref*{example_graphs}.
+The estimated graph from DAG-GNN(third column) has much more edges than the true graph(first column), while our method(second column) has much similar number of edges with the true graph.\\
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_dependence_50.pdf}
+        \caption{SHD with dependence proportion 0.5}
+        \label{SHD_prop50}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_dependence_50.pdf}
+        \caption{FDR with dependence proportion 0.5}
+        \label{FDR_prop50}
+    \end{subfigure}
+    \caption{SHD and FDR with dependence proportion 0.5}
+    \label{SHD_FDR_prop50}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{fig/comparison_dep_20_50_seed41.pdf}
+    \caption{Example of estimated graphs with node size 20 and dependence proportion 0.5}
+    \label{example_graphs}
+\end{figure}
+
+Next, we compare the performance of two methods with different proportions of bidirectional edges.
+The results are shown in Figure \ref*{SHD_FDR_node10_20}.
+When the node size is 10, DAG-GNN outperforms our method in terms of SHD.
+However, when the node size is 20, our method outperforms DAG-GNN in terms of both SHD and FDR.\\
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_dependence_10_DAG_threshold0.3.pdf}
+        \caption{SHD with node size 10}
+        \label{SHD_node10}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_dependence_10_DAG_threshold0.3.pdf}
+        \caption{FDR with node size 10}
+        \label{FDR_node10}
+    \end{subfigure}
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_dependence_20_DAG_threshold0.3.pdf}
+        \caption{SHD with node size 20}
+        \label{SHD_node20}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_dependence_20_DAG_threshold0.3.pdf}
+        \caption{FDR with node size 20}
+        \label{FDR_node20}
+    \end{subfigure}
+    \caption{SHD and FDR with dependence proportion 0.1, 0.3, 0.5, and 0.8}
+    \label{SHD_FDR_node10_20}
+\end{figure}
 

### 2023-12-02 Tex/paper.tex
run IAF 30 nodes

@@
 \usepackage{hyperref}
+\usepackage[inkscapeformat=png]{svg}
 \modulolinenumbers[5]
@@
 For the size of the hidden layer, we use 64.
+Note that the $\mathbf{L_X}$ is an additional output from the encoder network.
+Since the initial output $\mathbf{L_X}$ is not lower triangular matrix, we apply the lower triangular matrix transformation to $\mathbf{L_X}$ to make it lower triangular matrix.
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{fig/model.png}
+    \includegraphics[width=0.9\textwidth]{fig/model.pdf}
     \caption{Architecture of the proposed model.}
@@
 \begin{equation}
-    \mathcal{L}(A, W, \lambda) =-\mathcal{L}_{\mathrm{ELBO}} + \tau_A \Vert A \Vert_1 + \lambda h(A) + \frac{c}{2} |h(A)|^2.
+    \mathcal{L}(A, W, \lambda) =-\mathcal{L}_{\mathrm{ELBO}} + \tau \Vert A \Vert_1 + \lambda h(A) + \frac{c}{2} |h(A)|^2.
 \end{equation}
@@
 The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
+We found that $\tau=0.1$ works well for a node size of 50.
+For other node sizes, we use $\tau$ proportional to the inverse of the square of the node size.
 The third and fourth terms are the augmented Lagrangian terms.
-Since larger value of $c$ makes the acyclicity constraint to zero, the training process is done by gradually increasing the value of $c$.
-\cite{yu2019daggnn}\\
+We gradually increase the value of $c$ during the training process, as a larger value of $c$ reduces the acyclicity constraint to zero \cite{yu2019daggnn}.
 
@@
 
-First, when the proportion of bidirectional edges is fixed to 0.5, we compare the performance of our method with DAG-GNN.
-The results are shown in Figure \ref*{SHD_FDR_prop50}.
-Our method outperforms DAG-GNN in terms of both SHD and FDR.
-Also, at the node size of 20, our method has shown stable training, which results a huge difference in SHD and FDR.
-The example of estimated graphs of each method is shown in Figure \ref*{example_graphs}.
-The estimated graph from DAG-GNN(third column) has much more edges than the true graph(first column), while our method(second column) has much similar number of edges with the true graph.\\
+Firstly, we set the proportion of bidirectional edges to 0.5 and compared the performance of our method with that of DAG-GNN.
+The results, as depicted in Figure \ref*{SHD_FDR_prop50}, demonstrate that our method surpasses DAG-GNN in terms of both SHD and FDR.
+Notably, with a node size of 20, our method exhibits stable training, resulting in a significant difference in SHD and FDR.
+Figure \ref*{example_graphs} presents the estimated graphs produced by each method.
+The graph estimated by DAG-GNN (third column) contains considerably more edges than the true graph (first column), whereas the graph estimated by our method (second column) has a number of edges much closer to that of the true graph.
 
@@
 
-Next, we compare the performance of two methods with different proportions of bidirectional edges.
-The results are shown in Figure \ref*{SHD_FDR_node10_20}.
-When the node size is 10, DAG-GNN outperforms our method in terms of SHD.
-However, when the node size is 20, our method outperforms DAG-GNN in terms of both SHD and FDR.\\
+Next, we juxtapose the performance of the two methods, each with varying proportions of bidirectional edges.
+The results, as illustrated in Figure \ref*{SHD_FDR_node10_20}, reveal that when the node size is 10, DAG-GNN surpasses our method in terms of SHD.
+However, when the node size is increased to 20, our method takes the lead, outperforming DAG-GNN in both SHD and FDR metrics.
 
@@
     \end{subfigure}
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
+        \caption{SHD with node size 30}
+        \label{SHD_node30}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
+        \caption{FDR with node size 30}
+        \label{FDR_node30}
+    \end{subfigure}
     \caption{SHD and FDR with dependence proportion 0.1, 0.3, 0.5, and 0.8}
     \label{SHD_FDR_node10_20}
+\end{figure}
+
+For the number of predicted edges, we observe that our method produces a number of edges closer to the true number of edges than DAG-GNN, especially when the node size is 20.
+As shown in Figure \ref*{pred_edge}, the number of predicted edges from both methods are similar when the node size is 10.
+However, when the node size is 20, our method produces a number of predicted edges closer to the true number of edges than DAG-GNN.
+Considering that our method outperforms DAG-GNN in terms of SHD and FDR, we can conclude that our method produces a more accurate graph structure than DAG-GNN.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_dependence_10_DAG_threshold0.3.pdf}
+        \caption{Predicted edges with node size 10}
+        \label{pred_edge_node10}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_dependence_20_DAG_threshold0.3.pdf}
+        \caption{Predicted edges with node size 20}
+        \label{pred_edge_node20}
+    \end{subfigure}
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
+        \caption{Predicted edges with node size 30}
+        \label{pred_edge_node30}
+    \end{subfigure}
+    \caption{Number of predicted edges with dependence proportion 0.1, 0.3, 0.5, and 0.8}
+    \label{pred_edge}
 \end{figure}

### 2023-12-04 Tex/paper.tex
tex

@@
 
-Next, we juxtapose the performance of the two methods, each with varying proportions of bidirectional edges.
-The results, as illustrated in Figure \ref*{SHD_FDR_node10_20}, reveal that when the node size is 10, DAG-GNN surpasses our method in terms of SHD.
+Next, we compare the performance of the two methods, each with varying proportions of bidirectional edges.
+The results, as illustrated in Figure \ref*{SHD_FDR}, reveal that when the node size is 10, DAG-GNN surpasses our method in terms of SHD.
 However, when the node size is increased to 20, our method takes the lead, outperforming DAG-GNN in both SHD and FDR metrics.
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
     \end{subfigure}
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_dependence_50_DAG_threshold0.3.pdf}
+        \caption{SHD with node size 50}
+        \label{SHD_node50}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_dependence_50_DAG_threshold0.3.pdf}
+        \caption{FDR with node size 50}
+        \label{FDR_node50}
+    \end{subfigure}
     \caption{SHD and FDR with dependence proportion 0.1, 0.3, 0.5, and 0.8}
-    \label{SHD_FDR_node10_20}
+    \label{SHD_FDR}
 \end{figure}
 
-For the number of predicted edges, we observe that our method produces a number of edges closer to the true number of edges than DAG-GNN, especially when the node size is 20.
+In terms of the number of predicted edges, we observe that our method produces a number of edges closer to the true number of edges than DAG-GNN, especially when the node size is 20.
 As shown in Figure \ref*{pred_edge}, the number of predicted edges from both methods are similar when the node size is 10.
-However, when the node size is 20, our method produces a number of predicted edges closer to the true number of edges than DAG-GNN.
+However, when the node size is 20, 30 and 50, our method produces a number of predicted edges closer to the true number of edges than DAG-GNN.
 Considering that our method outperforms DAG-GNN in terms of SHD and FDR, we can conclude that our method produces a more accurate graph structure than DAG-GNN.
@@
     \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_dependence_50_DAG_threshold0.3.pdf}
+        \caption{Predicted edges with node size 50}
+        \label{pred_edge_node50}
+    \end{subfigure}
     \caption{Number of predicted edges with dependence proportion 0.1, 0.3, 0.5, and 0.8}

### 2023-12-04 Tex/paper.tex
tex

@@
 Firstly, we set the proportion of bidirectional edges to 0.5 and compared the performance of our method with that of DAG-GNN.
-The results, as depicted in Figure \ref*{SHD_FDR_prop50}, demonstrate that our method surpasses DAG-GNN in terms of both SHD and FDR.
+The results, as depicted in Figure \ref*{SHD_FDR_prop30}, demonstrate that our method surpasses DAG-GNN in terms of both SHD and FDR.
 Notably, with a node size of 20, our method exhibits stable training, resulting in a significant difference in SHD and FDR.
@@
 
-\begin{figure}[H]
+\begin{figure}
     \centering
     \begin{subfigure}{0.45\textwidth}
-        \includegraphics[width=\textwidth]{fig/SHD_dependence_50.pdf}
-        \caption{SHD with dependence proportion 0.5}
-        \label{SHD_prop50}
+        \includegraphics[width=\textwidth]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
+        \caption{SHD with dependence proportion 0.3}
+        \label{SHD_prop30}
     \end{subfigure}
@@
     \begin{subfigure}{0.45\textwidth}
-        \includegraphics[width=\textwidth]{fig/FDR_dependence_50.pdf}
-        \caption{FDR with dependence proportion 0.5}
-        \label{FDR_prop50}
+        \includegraphics[width=\textwidth]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
+        \caption{FDR with dependence proportion 0.3}
+        \label{FDR_prop30}
     \end{subfigure}
-    \caption{SHD and FDR with dependence proportion 0.5}
-    \label{SHD_FDR_prop50}
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
+        \caption{Number of predicted edges}
+        \label{pred_edge_prop30}
+    \end{subfigure}
+    \caption{SHD, FDR, and number of predicted edges with dependence proportion 0.3}
+    \label{SHD_FDR_prop30}
 \end{figure}

### 2023-12-04 NOTEARS_train.py

@@ -76,7 +76,7 @@ np.random.seed(args.seed)
 
 now = datetime.datetime.now().strftime('%m%d_%H%M')
 if args.dependence_type == 1:
-    folder = f'results/dependence/{now}_NOTEARS_node{args.node_size}_prop{int(args.dependence_prop*100)}'
+    folder = f'results/dependence/{now}_NOTEARS_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
 else:
     folder = f'results/independence/{args.graph_dist}/{now}_NOTEARS_node{args.node_size}'
 
@@ -139,6 +139,7 @@ graph[np.abs(graph) < 0.3] = 0
 # print(graph)
 fdr, tpr, fpr, shd, nnz = count_accuracy(G, nx.DiGraph(graph))
 print('threshold 0.3, Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz, file=log)
+print('threshold 0.3, Accuracy: fdr', fdr, ' tpr ', tpr, ' fpr ', fpr, 'shd', shd, 'nnz', nnz)
 
 f = open(folder + '/trueG.txt', 'w')
 matG = np.matrix(nx.to_numpy_array(G))

### 2023-12-04 model.py

@@ -176,7 +176,7 @@ class VAE_IAF(nn.Module):
             self.encoder_L = nn.Linear(args.z_dims, args.z_size)
         else:
             self.encoder_L = nn.Linear(args.encoder_hidden, args.z_size)
-        
+
         self.softmax = nn.Softmax()
 
         for m in self.modules():

### 2023-12-04 train.py

@@ -91,7 +91,7 @@ parser.add_argument('--x_dims', type=int, default=1, #changed here
                     help='The number of input dimensions: default 1.')
 parser.add_argument('--z_dims', type=int, default=1,
                     help='The number of latent variable dimensions: default the same as variable size.')
-parser.add_argument('--number_of_flows', type=int, default=5,
+parser.add_argument('--number_of_flows', type=int, default=1,
                     help='The number of HF flows: default 5.')
 parser.add_argument('--flow_type', type=str, default='IAF',
                     help='The type of flows: "DAGGNN", "IAF", "HF"(Householder), "ccIAF"')

### 2023-12-04 train.py
run IAF vs DAGGNN 10 nodes independent laplace and exponential

@@ -107,6 +107,10 @@ parser.add_argument('--logits', type=int, default=1,
 args = parser.parse_args()
 args.z_size = args.node_size # the number of latent variables
 
+# set up tau_A
+if args.tau_A == 0 and args.flow_type == 'IAF':
+    args.tau_A = 0.1 * (args.node_size / 50) ** 2
+
 # Device configuration (GPU or MPS or CPU)
 
 if torch.cuda.is_available() and args.cuda:

### 2023-12-05 Tex/paper.tex
tex

@@
 
-Here is the introduction of our experiment on independent noise case
+We compare first the performance of our method with DAG-GNN in the case where the noise variables are independent.
+To be specific, we consider the case where the noise variables follow an independent Gaussian, Laplace, and exponential distribution.
+The results of the experiments are shown in Figure \ref*{SHD_FDR_gaussian}, \ref*{SHD_FDR_laplace}, and \ref*{SHD_FDR_exponential}.
+In all cases, our method outperforms DAG-GNN in terms of both SHD and FDR.\\
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_independence_laplace.pdf}
+        \caption{SHD with Laplace noise}
+        \label{SHD_laplace}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_independence_laplace.pdf}
+        \caption{FDR with Laplace noise}
+        \label{FDR_laplace}
+    \end{subfigure}
+    \caption{SHD and FDR with Laplace noise}
+    \label{SHD_FDR_laplace}
+\end{figure}
 
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{fig/comparison_dep_20_50_seed41.pdf}
+    \includegraphics[width=0.8\textwidth]{fig/comparison_dep_20_50_seed31.pdf}
     \caption{Example of estimated graphs with node size 20 and dependence proportion 0.5}

### 2023-12-05 Tex/paper.tex
commit

@@
 To be specific, we consider the case where the noise variables follow an independent Gaussian, Laplace, and exponential distribution.
-The results of the experiments are shown in Figure \ref*{SHD_FDR_gaussian}, \ref*{SHD_FDR_laplace}, and \ref*{SHD_FDR_exponential}.
+The results of the experiments are shown in Figure \ref*{SHD_FDR_laplace}, and \ref*{SHD_FDR_exponential}.
 In all cases, our method outperforms DAG-GNN in terms of both SHD and FDR.\\
@@
     \label{SHD_FDR_laplace}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_independence_exponential.pdf}
+        \caption{SHD with exponential noise}
+        \label{SHD_exponential}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_independence_exponential.pdf}
+        \caption{FDR with exponential noise}
+        \label{FDR_exponential}
+    \end{subfigure}
+    \caption{SHD and FDR with exponential noise}
+    \label{SHD_FDR_exponential}
 \end{figure}

### 2023-12-05 Tex/paper.tex

@@
 The results of the experiments are shown in Figure \ref*{SHD_FDR_laplace}, and \ref*{SHD_FDR_exponential}.
-In all cases, our method outperforms DAG-GNN in terms of both SHD and FDR.\\
+In all cases, our method outperforms DAG-GNN in terms of both SHD and FDR as the number of nodes increases.\\
 
@@
 
+Also, we observe that the number of predicted edges from our method is closer to the true number of edges than DAG-GNN.
+As shown in Figure \ref*{pred_edge_laplace_exponential}, the number of predicted edges from both methods are similar when the node size is 10.
+However, as the node size increases, our method produces smaller number of predicted edges than DAG-GNN.
+Considering that our method outperforms DAG-GNN in terms of SHD and FDR, we can conclude that our method produces a more accurate graph structure than DAG-GNN in the cases where the noise variables are independent.
+Figure \ref*{example_graphs_indep} presents the estimated graphs produced by each method.
+The graph estimated by DAG-GNN (third column) contains considerably more edges than the true graph (first column), whereas the graph estimated by our method (second column) has a number of edges much closer to that of the true graph.
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_laplace.pdf}
+        \caption{Predicted edges with Laplace noise}
+        \label{pred_edge_laplace}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.45\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_exponential.pdf}
+        \caption{Predicted edges with exponential noise}
+        \label{pred_edge_exponential}
+    \end{subfigure}
+    \caption{Number of predicted edges with Laplace and exponential noise}
+    \label{pred_edge_laplace_exponential}
+\end{figure}
+
+\begin{figure}[H]
+    \centering
+    \includegraphics[width=0.8\textwidth]{fig/comparison_indep_30_exponential_seed301.pdf}
+    \caption{Example of estimated graphs with node size 30 and exponential noise}
+    \label{example_graphs_indep}
+\end{figure}
+
 \subsection{Dependent Noise Case}
@@
 
-\begin{figure}
+\begin{figure}[p]
     \centering

### 2023-12-06 model.py
conflict

@@ -1,378 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import torch.distributions as D
-from torch.autograd import Variable
-<<<<<<< Updated upstream
-<<<<<<< Updated upstream
-import torch.nn.utils.weight_norm as wn
-import math
-=======
-from utils import *
->>>>>>> Stashed changes
-=======
-from utils import *
->>>>>>> Stashed changes
-
-
-<<<<<<< Updated upstream
-# What to do?
-=======
-# z로 변수변환 시 volume을 보존한다고 하는 convex combination (refer to Tomczak & Welling (2017))
-
-class combination_L(nn.Module):
-    def __init__(self,args):
-        super(combination_L, self).__init__()
-        self.args = args
-
-    def forward(self, L, y):
-        '''
-        :param L: batch_size (B) x latent_size^2 * number_combination (L^2 * C)
-        :param y: batch_size (B) x number_combination (C)
-        :return: L_combination = y * L
-        '''
-        # calculate combination of Ls
-        L_tensor = L.view( -1, self.args.z1_size**2, self.args.number_combination ) # resize to get B x L^2 x C
-        y = y.unsqueeze(1).expand(y.size(0), self.args.z1_size**2, y.size(1)) # expand to get B x L^2 x C
-        L_combination = torch.sum( L_tensor * y, 2 ).squeeze()
-        return L_combination
-
-# self.args.number_combination 을 설정해야 될지도
-
-class FlowLayer(nn.Module):
-    """
-    Flow layer class
-    """
-    def __init__(self, args):
-        super(FlowLayer, self).__init__()
-        self.args = args
-    
-    def forward(self, L, z):
-        """
-        :param L: batch_size (B) x latent_size^2 (L^2) from encoder output
-        :param z: batch_size (B) x latent_size (L) from encoder output z0
-        :return: z_new = L * z
-        """
-        # transform L to lower triangular matrix
-        L_matrix = L.view(-1, self.args.z_size, self.args.z_size) # resize to get B x L x L
-        LTmask = torch.tril(torch.ones(self.args.z_size, self.args.z_size), diagonal=-1) # lower-triangular mask matrix
[…]

### 2023-12-06 train.py

@@ -1,386 +1,0 @@
-<<<<<<< Updated upstream
-# What to do?
-=======
-import time
-import datetime
-import pickle
-import os
-import argparse
-
-import torch
-from torch.optim import lr_scheduler
-from torch.utils.data import Dataset, DataLoader
-from torch.autograd import Variable
-from torch.utils.data.dataset import TensorDataset
-import math
-import numpy as np
-from utils import *
-from model import *
-from data import *
->>>>>>> Stashed changes
-
-from __future__ import print_function
-
-from utils.distributions import log_Normal_diag, log_Normal_standard, log_Bernoulli
-
-# 1. Arguments(argparse)
-
-
-
-<<<<<<< Updated upstream
-=======
-args = parser.parse_args()
-print(args.dependence_type)
-
-# Device configuration (GPU or MPS or CPU)
-
-if torch.cuda.is_available():
-    device = torch.device('cuda')
-elif torch.backends.mps.is_available():
-    device = torch.device('mps')
-else:
-    device = torch.device('cpu')
-
-# Seed
-torch.manual_seed(args.seed)
-
-# Folder to save the results, models, dataset
-
-now = datetime.datetime.now().strftime('%Y%m%d_%H%M%S')
-if args.dependence_type == 1:
-    folder = f'results/dependence/{now}'
-else:
-    folder = f'results/independence/{args.graph_dist}/{now}'
-
-if not os.path.exists(folder):
-    os.makedirs(folder)
-
-# Save the arguments
-meta_file = os.path.join(folder, 'meta.pkl')
-vae_indep_file = os.path.join(folder, 'vae_indep.pt')
-vae_dep_file = os.path.join(folder, 'vae_dep.pt')
-
-log_file = os.path.join(folder, 'log.txt')
-log = open(log_file, 'w')
-
-pickle.dump(args, open(meta_file, 'wb'))
-
-# 2. Load Data
-
-# 2.1. Generate DAG
-G = generate_random_dag(d = args.node_size, degree=args.graph_degree, seed=args.seed)
-
-# 2.2. Generate Data
-if args.dependence_type == 1:
-    X = generate_linear_sem_correlated(G, args.data_sample_size, args.dependence_prop, args.seed)
-else:
[…]

### 2023-12-06 Tex/paper.tex
Tex experiment 수정완료

@@
 
-\subsection{Independent Noise Case}
+\subsection{Independent Noise Cases}
 
-We compare first the performance of our method with DAG-GNN in the case where the noise variables are independent.
-To be specific, we consider the case where the noise variables follow an independent Gaussian, Laplace, and exponential distribution.
-The results of the experiments are shown in Figure \ref*{SHD_FDR_laplace}, and \ref*{SHD_FDR_exponential}.
-In all cases, our method outperforms DAG-GNN in terms of both SHD and FDR as the number of nodes increases.\\
+To compare the performance of our method with DAG-GNN, we conduct experiments on random graphs with 10, 20, and 30 nodes, assuming independent noise variables following Gaussian, Laplace, and exponential distributions.
+The results for the Gaussian case are shown in Figure \ref*{SHD_FDR_gaussian}.
+It can be observed that DAG-GNN outperforms our method in terms of SHD and FDR, considering the number of predicted edges.
+This is because DAG-GNN assumes that the noise variables follow a standard Gaussian distribution, making it more suitable for cases with independent Gaussian noise variables.
+In contrast, our method may be considered overparameterized in such cases.\\
+
+\begin{figure}[H]
+    \centering
+    \begin{subfigure}{0.3\textwidth}
+        \includegraphics[width=\textwidth]{fig/SHD_independence_normal.pdf}
+        \caption{SHD with Gaussian noise}
+        \label{SHD_gaussian}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.3\textwidth}
+        \includegraphics[width=\textwidth]{fig/FDR_independence_normal.pdf}
+        \caption{FDR with Gaussian noise}
+        \label{FDR_gaussian}
+    \end{subfigure}
+    \hfill
+    \begin{subfigure}{0.3\textwidth}
+        \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_normal.pdf}
+        \caption{Number of predicted edges}
+        \label{pred_edge_gaussian}
+    \end{subfigure}
+    \caption{SHD, FDR, and number of predicted edges with Gaussian noise}
+    \label{SHD_FDR_gaussian}
+\end{figure}
+
+However, in cases where the noise variables follow Laplace or exponential distributions, our method consistently outperformed DAG-GNN in terms of both SHD and FDR.
+The results of the experiments are shown in Figure \ref*{SHD_FDR_laplace} and \ref*{SHD_FDR_exponential}.
+As the number of nodes increases, our method consistently demonstrates better performance compared to DAG-GNN in terms of both SHD and FDR.\\
 
@@
 
-Also, we observe that the number of predicted edges from our method is closer to the true number of edges than DAG-GNN.
-As shown in Figure \ref*{pred_edge_laplace_exponential}, the number of predicted edges from both methods are similar when the node size is 10.
-However, as the node size increases, our method produces smaller number of predicted edges than DAG-GNN.
-Considering that our method outperforms DAG-GNN in terms of SHD and FDR, we can conclude that our method produces a more accurate graph structure than DAG-GNN in the cases where the noise variables are independent.
+Furthermore, we observed that our method produces a more accurate graph structure than DAG-GNN in cases where the noise variables are independent.
+As shown in Figure \ref*{pred_edge_laplace_exponential}, the number of predicted edges from both methods is similar when the node size is 10.
+However, as the node size increases, our method consistently produces a smaller number of predicted edges compared to DAG-GNN.
+This is in line with our method's superior performance in terms of SHD and FDR.
 Figure \ref*{example_graphs_indep} presents the estimated graphs produced by each method.
 The graph estimated by DAG-GNN (third column) contains considerably more edges than the true graph (first column), whereas the graph estimated by our method (second column) has a number of edges much closer to that of the true graph.
+Therefore, our method not only achieves better performance in terms of accuracy but also provides a more reliable estimation of the true graph structure.\\
 
@@
 
-\subsection{Dependent Noise Case}
+\subsection{Dependent Noise Cases}
 
-Firstly, we set the proportion of bidirectional edges to 0.5 and compared the performance of our method with that of DAG-GNN.
-The results, as depicted in Figure \ref*{SHD_FDR_prop30}, demonstrate that our method surpasses DAG-GNN in terms of both SHD and FDR.
-Notably, with a node size of 20, our method exhibits stable training, resulting in a significant difference in SHD and FDR.
+To begin with, we set the proportion of bidirectional edges to 0.5 and compared the performance of our method with that of DAG-GNN.
+The results, as depicted in Figure \ref*{SHD_FDR_prop30}, demonstrate that our method outperforms DAG-GNN in terms of both SHD and FDR.
+Notably, with a node size of 20, our method exhibited stable training, resulting in a significant difference in SHD and FDR.
 Figure \ref*{example_graphs} presents the estimated graphs produced by each method.
-The graph estimated by DAG-GNN (third column) contains considerably more edges than the true graph (first column), whereas the graph estimated by our method (second column) has a number of edges much closer to that of the true graph.
+The graph estimated by DAG-GNN (third column) contains considerably more edges than the true graph (first column), whereas the graph estimated by our method (second column) has a number of edges much closer to that of the true graph.\\
 
@@
 The results, as illustrated in Figure \ref*{SHD_FDR}, reveal that when the node size is 10, DAG-GNN surpasses our method in terms of SHD.
-However, when the node size is increased to 20, our method takes the lead, outperforming DAG-GNN in both SHD and FDR metrics.
+However, when the node size is increased to 20, our method takes the lead, outperforming DAG-GNN in both SHD and FDR metrics.\\
 
@@
 
-In terms of the number of predicted edges, we observe that our method produces a number of edges closer to the true number of edges than DAG-GNN, especially when the node size is 20.
-As shown in Figure \ref*{pred_edge}, the number of predicted edges from both methods are similar when the node size is 10.
-However, when the node size is 20, 30 and 50, our method produces a number of predicted edges closer to the true number of edges than DAG-GNN.
-Considering that our method outperforms DAG-GNN in terms of SHD and FDR, we can conclude that our method produces a more accurate graph structure than DAG-GNN.
+In terms of the number of predicted edges, our method consistently produces a number of edges closer to the true number compared to DAG-GNN, especially when the node size is greater than 10.
+As depicted in Figure \ref*{pred_edge}, the number of predicted edges from both methods are similar when the node size is 10.
+However, for node sizes 20, 30, and 50, our method consistently produces a number of predicted edges that aligns more closely with the true number of edges compared to DAG-GNN.
+Considering the better performance of our method in terms of SHD and FDR metrics, we can conclude that our method yields a more accurate graph structure than DAG-GNN.
 
-\begin{figure}[H]
+\begin{figure}
     \centering
@@
 
-Here is the conclusion
+In this paper, we proposed a method to learn the structure of a semi-Markovian DAG using a flow-based VAE.
+We conducted experiments on simulated data and compared the performance of our method with that of DAG-GNN.
+The results demonstrate that our method outperforms DAG-GNN in terms of both SHD and FDR metrics, especially when the noise variables are dependent and the size of graph is large.
+
+However, contrary to our initial expectations, the lower triangular matrix $\mathbf L$ learned in the IAF layer did not capture the covariance matrix of the actual noise variables.
+This seems to be due to the entanglement problem in the latent space, and resolving this issue remains a future research task.
+If the full covariance matrix of the noise variables could be found, we would be able to directly identify bidirectional edges in the graph, enabling more accurate learning of the semi-Markovian graph.
 

### 2023-12-07 model_newIAF.py
tex

@@ -1,0 +1,198 @@
+import torch
+import torch.nn as nn
+import torch.nn.functional as F
+import torch.distributions as D
+from torch.autograd import Variable
+import torch.nn.utils.weight_norm as wn
+import math
+from utils import *
+import numpy as np
+
+
+from torch.autograd import Variable
+
+class FlowLayer(nn.Module):
+    """
+    Flow layer class
+    """
+    def __init__(self, args):
+        super(FlowLayer, self).__init__()
+        self.args = args
+    
+    def forward(self, L, z):
+        """
+        :param L: batch_size (B) x latent_size^2 (L^2) from encoder output
+        :param z: batch_size (B) x latent_size (L) from encoder output z0
+        :return: z_new = L * z
+        """
+        # transform L to lower triangular matrix
+        L_matrix = L.view(-1, self.args.z_size, self.args.z_size) # resize to get B x L x L
+        LTmask = torch.tril(torch.ones(self.args.z_size, self.args.z_size), diagonal=-1) # lower-triangular mask matrix
+        I = Variable(torch.eye(self.args.z_size, self.args.z_size).expand(L_matrix.size(0), self.args.z_size, self.args.z_size))
+        if self.args.cuda:
+            LTmask = LTmask.cuda()
+            I = I.cuda()
+        LTmask = Variable(LTmask)
+        LTmask = LTmask.unsqueeze(0).expand(L_matrix.size(0), self.args.z_size, self.args.z_size)
+        LT = torch.mul(L_matrix, LTmask) + I # Lower triangular batches
+        z_new = torch.bmm(LT, z) # B x L x L * B x L x 1 = B x L x 1
+
+        return z_new, LT
+
+class Encoder(nn.Module):
+    """
+    Encoder class for VAE
+    """
+    def __init__(self, args, adj_A, tol=0.1):
+        super(Encoder, self).__init__()
+
+        self.args = args
+        self.adj_A = nn.Parameter(torch.from_numpy(adj_A).double(), requires_grad=True)
+        self.Wa = nn.Parameter(torch.zeros(args.z_dims), requires_grad=True)
+        self.fc1 = nn.Linear(args.x_dims, args.encoder_hidden, bias=True)
+        self.fc2 = nn.Linear(args.encoder_hidden, args.z_dims, bias=True)
+        self.batch_size = args.batch_size
[…]

### 2023-12-07 train_new.py

@@ -11,11 +11,25 @@ from torch.utils.data.dataset import TensorDataset
 import math
 import numpy as np
 from utils import *
-from model import *
+from model_newIAF import *
 from data import *
 from loss import *
 from tqdm import tqdm
 import pandas as pd
+import matplotlib.pyplot as plt
+import seaborn as sns
+import scienceplots
+
+plt.style.use(['science', 'ieee'])
+from matplotlib import cm
+from matplotlib.colors import ListedColormap, LinearSegmentedColormap
+
+top = cm.get_cmap('Oranges_r', 128)
+bottom = cm.get_cmap('Blues', 128)
+
+newcolors = np.vstack((top(np.linspace(0, 1, 128)),
+                       bottom(np.linspace(0, 1, 128))))
+newcmp = ListedColormap(newcolors, name='OrangeBlue')
 
 os.environ["KMP_DUPLICATE_LIB_OK"]="TRUE"
 
@@ -108,8 +122,8 @@ args = parser.parse_args()
 args.z_size = args.node_size # the number of latent variables
 
 # set up tau_A
-if args.tau_A == 0 and args.flow_type == 'IAF':
-    args.tau_A = 0.1 * (args.node_size / 50) ** 2
+# if args.tau_A == 0 and args.flow_type == 'IAF':
+#     args.tau_A = 0.1 * (args.node_size / 50) ** 2
 
 # Device configuration (GPU or MPS or CPU)
 
@@ -135,9 +149,9 @@ else:
     raise ValueError(f'Invalid flow type: {args.flow_type}.')
 
 if args.dependence_type == 1:
-    folder = f'results/dependence/{now}_{flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
+    folder = f'results_new/dependence/{now}_{flow_type}_node{args.node_size}_prop{int(args.dependence_prop*100)}_seed{args.seed}'
 else:
-    folder = f'results/independence/{args.graph_dist}/{now}_{flow_type}_node{args.node_size}_seed{args.seed}'
+    folder = f'results_new/independence/{args.graph_dist}/{now}_{flow_type}_node{args.node_size}_seed{args.seed}'
 
 if not args.lagrange:
     folder += '_noL'
[…]

### 2023-12-10 Tex/paper.tex
tex 수정

@@
 
-Here is the introduction of VAEs
+VAE(Variational Auto-Encoder)\cite{kingma2013auto} is a neural network architecture that consists of a generative model and a variational inference.
+In VAE architecture, it is assumed that the continuous latent variable exists and its posterior probability is intractable.
+Therefore, variational inference is introduced to approximate the real posterior probability with ELBO(evidence lower bound).\\
+
+Let $x$ be an observed variable and $z$ be a latent variable.
+$x$ is generated by some random process involving $z$.
+For parameter $\theta$, random variable $z$ is generated from prior distribution $p_{\theta}(z)$, then $x$ is generated from conditional distribution $p_{\theta}(x|z)$.
+And $p_{\theta}(x,z)$ becomes their parametric joint distribution.
+Given a dataset $\mathsf{X} = \{ x^{(1)}, ...
+, x^{(N)}\}$, maximum marginal likelihood learning is to maximize log likelihoods $\log{p_{\theta}(\mathsf{X})} = \sum_{i=1}^{N}{\log{p_{\theta}(x^{(i)})}}.$ However, marginal likelihood $p_{\theta}(x) = \int{p_{\theta}(z)p_{\theta}(x|z)dz}$ is intractable in general.
+Instead, we introduce the model parameter $\phi$ and a parametric inference model $q_{\phi}(z|x).$ Then we can reformulate the original log likelihoods with $q_{\phi}(z|x).$ \\
+\begin{equation*}
+	\begin{aligned}
+		\log{p_{\theta}(x)} =  \mathbb{E}_{q_{\phi}(z|x)}\left[ \log{p_{\theta}(x)} \right] &= \mathbb{E}_{q_{\phi}(z|x)}\left[ \log{p_{\theta}(x)} + \log{p_{\theta}(z|x)} - \log{p_{\theta}(z|x)} \right] \\
+		&= \mathbb{E}_{q_{\phi}(z|x)}\left[ \log{p_{\theta}(x, z)} - \log{q_{\phi}(z|x)} \right] + \mathbb{E}_{q_{\phi}(z|x)}\left[ \log{q_{\phi}(z|x)} - \log{p_{\theta}(z|x)} \right] \\
+	\end{aligned}
+\end{equation*}
+\\
+
+The second RHS term of the last equation is the KL divergence of the approximate posterior from the true posterior.
+Since KL divergence is non-negative, the first RHS term of the last equation can be expressed as $\mathcal{L}(\theta, \phi; x)$ and it is called variational lower bound.
+\begin{equation*}
+	\begin{aligned}
+	\mathcal{L}(\theta, \phi; x) &= \mathbb{E}_{q_{\phi}(z|x)}\left[ \log{p_{\theta}(x, z)} - \log{q_{\phi}(z|x)} \right] \\
+	\log{p_{\theta}(x)} &= \mathcal{L}(\theta, \phi; x) + D_{KL}( q_{\phi}(z|x) || p_{\theta}(z|x)) \geq \mathcal{L}(\theta, \phi ; x) \\
+	\end{aligned}
+\end{equation*}
+\\
+Therefore $\mathcal{L}(\theta, \phi; x)$ is the lower bound of the evidence, marginal likelihood.
+Under some condition of the distribution family $\{ q_{\phi} \}, $ maximizing $\mathcal{L}(\theta, \phi; x)$ with respect to $\{ \theta, \phi \}$ will simultaneously maximize log-likelihood $\log{p_{\theta}(x)}$ and minimize $ D_{KL}( q_{\phi}(z|x) || p_{\theta}(z|x))$.
+In VAE architecture, we set $\mathcal{L}(\theta, \phi; x)$ as objective function to be maximized.\\
+
+The ELBO also can be written as 
+\begin{equation*}
+	\begin{aligned}
+		\mathcal{L}(\theta, \phi ; x) &= -\mathbb{E}_{q_{\phi}(z|x)}\left[ \log{q_{\phi}(z|x)} - \log{p_{\theta}(z)} - \log{p_{\theta}(x|z)}  \right] \\
+		&= -D_{KL}( q_{\phi}(z|x) || p_{\theta}(z)) + \mathbb{E}_{q_{\phi}(z|x)}\left[ \log{p_{\theta}(x|z)}  \right]. \\
+	\end{aligned}
+\end{equation*}
+
+To estimate the lower bound, sampling method is used.
+However, Monte Carlo estimate of the expectation over $q_{\phi}(z|x)$ should be differentiable with respect to $\phi$ in gradient method.
+Reparametrization trick is introduced to solve this problem.
+It is often possible to express random variable $z \sim q_{\phi}(z|x) $ as a deterministic variable $z = g_{\phi}(\epsilon, x)$ for auxiliary variable $\epsilon \sim p(\epsilon)$.
+This reparametrization moves the uncertainty induced by $\phi$ into $\epsilon$ and makes the whole estimation differentiable.
 
@@
 
-To compare the performance of our method with DAG-GNN, we conduct experiments on random graphs with 10, 20, and 30 nodes, assuming independent noise variables following Gaussian, Laplace, and exponential distributions.
-The results for the Gaussian case are shown in Figure \ref*{SHD_FDR_gaussian}.
-It can be observed that DAG-GNN outperforms our method in terms of SHD and FDR, considering the number of predicted edges.
-This is because DAG-GNN assumes that the noise variables follow a standard Gaussian distribution, making it more suitable for cases with independent Gaussian noise variables.
-In contrast, our method may be considered overparameterized in such cases.\\
+To compare the performance of our method with DAG-GNN, we conducted experiments on random graphs with independent noise variables.
+Specifically, we considered cases where the noise variables followed independent Gaussian, Laplace, and exponential distributions.
+The results of the experiments are shown in Figures \ref*{SHD_FDR_gaussian}, \ref*{SHD_FDR_laplace}, and \ref*{SHD_FDR_exponential}.
+In the case of Gaussian noise, despite using the original settings of DAG-GNN, the normalizing flow did not show superior performance compared to DAG-GNN.
+On average, for 10 nodes, our model and DAG-GNN achieved SHD values of 16.5 and 13, respectively.
+For 20 nodes, the SHD values were 24.7 and 22.1, and for 30 nodes, they were 21.8 and 18.4.
+However, in the case of non-Gaussian noise, our method consistently outperformed DAG-GNN in terms of both SHD and FDR as the number of nodes increased.
 
@@
     \end{subfigure}
-    \caption{SHD, FDR, and number of predicted edges with Gaussian noise}
+    \caption{SHD, FDR, and number of predicted edges with independent Gaussian noise}
     \label{SHD_FDR_gaussian}
@@
 
-However, in cases where the noise variables follow Laplace or exponential distributions, our method consistently outperformed DAG-GNN in terms of both SHD and FDR.
+In cases where the noise variables follow Laplace or exponential distributions, our method consistently outperformed DAG-GNN in terms of both SHD and FDR.
 The results of the experiments are shown in Figure \ref*{SHD_FDR_laplace} and \ref*{SHD_FDR_exponential}.
-As the number of nodes increases, our method consistently demonstrates better performance compared to DAG-GNN in terms of both SHD and FDR.\\
+As the number of nodes increases, our method consistently demonstrates better performance compared to DAG-GNN in terms of SHD.
+The FDR values of our method are also lower than those of DAG-GNN, except for the case of 20 with exponential noise.\\
 
@@
     \end{subfigure}
-    \caption{SHD and FDR with Laplace noise}
+    \caption{SHD and FDR with independent Laplace noise}
     \label{SHD_FDR_laplace}
@@
     \end{subfigure}
-    \caption{SHD and FDR with exponential noise}
+    \caption{SHD and FDR with independent exponential noise}
     \label{SHD_FDR_exponential}
@@
     \centering
-    \includegraphics[width=0.8\textwidth]{fig/comparison_indep_30_exponential_seed301.pdf}
+    \includegraphics[width=0.8\textwidth]{fig/comparison_indep_30_laplace_seed201.pdf}
     \caption{Example of estimated graphs with node size 30 and exponential noise}
@@
 
-To begin with, we set the proportion of bidirectional edges to 0.5 and compared the performance of our method with that of DAG-GNN.
+To begin with, we set the proportion of bidirectional edges to 0.3 and compared the performance of our method with that of DAG-GNN.
 The results, as depicted in Figure \ref*{SHD_FDR_prop30}, demonstrate that our method outperforms DAG-GNN in terms of both SHD and FDR.
@@
 We conducted experiments on simulated data and compared the performance of our method with that of DAG-GNN.
-The results demonstrate that our method outperforms DAG-GNN in terms of both SHD and FDR metrics, especially when the noise variables are dependent and the size of graph is large.
+The results demonstrate that our method outperforms DAG-GNN in terms of both SHD and FDR metrics, especially when the noise variables have dependent structure and when the size of graph is large.
+In addition, our method produced a number of predicted edges closer to the true number of edges compared to DAG-GNN.\\
 
@@
 This seems to be due to the entanglement problem in the latent space, and resolving this issue remains a future research task.
-If the full covariance matrix of the noise variables could be found, we would be able to directly identify bidirectional edges in the graph, enabling more accurate learning of the semi-Markovian graph.
+If the full covariance matrix of the noise variables could be found, it could be possible to directly identify bidirectional edges in the graph, enabling more accurate learning of the semi-Markovian graph.
 

### 2023-12-10 Tex/slide.tex
tex 수정

\documentclass{beamer}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{kotex}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{graphicx}
\usepackage{caption}
\usepackage[backend=bibtex, style=authortitle-comp]{biblatex}
\addbibresource{references.bib}

\usetheme{Boadilla}
\title[NFG-VAE]{Learning semi-Markovian DAGs with flow-based VAE}
\author[Seoul National University]{Dangchan Kim, Byungguk Kang, Jaeseok Kim, Minchan Kim}
\institute[]{Seoul National University}
\date{\today}

\renewcommand{\footnotesize}{\fontsize{7pt}{5pt}\selectfont}

\begin{document}

\frame{\titlepage}

\setbeamertemplate{section in toc}[ball unnumbered]
\begin{frame}
    \frametitle{Table of Contents}
    \tableofcontents
\end{frame}

% This beamer tex file is based on my paper.tex

\section*{Overview}

\begin{frame}{Overview}
    \begin{itemize}
        \item \textbf{Problem}: Learning semi-Markovian DAGs
        \item \textbf{Solution}: Normalizing flow based VAE
        \item \textbf{Contribution}
        \begin{itemize}
            \item A new method for learning semi-Markovian DAGs
            \item A new method for learning DAGs with non-Gaussian noise
        \end{itemize}
    \end{itemize}
\end{frame}

\section{Preliminaries}

\begin{frame}{DAG and SEM}
    \begin{itemize}
        \item Directed Acyclic Graph(DAG) $\mathcal G = (V, E)$ with $V = \{1, \dots, m\}$
        \item Linear Structural Equation Model(SEM) is defined as
        \begin{equation*}
            X_i = \sum_{j \in \text{Pa}(i)} \beta_{ij} X_j + \epsilon_i
        \end{equation*}
        where $\epsilon_i$ is a noise variable.
        \item Matrix form can be written as
        \begin{equation}
            \mathbf X = \mathbf A^T \mathbf X + \mathbf \epsilon
        \end{equation}
        where $\mathbf A$ is an adjacency matrix of $\mathcal G$ and $\mathbf \epsilon$ is a noise vector.
    \end{itemize}
\end{frame}

\begin{frame}{DAG and SEM(Cont'd)}
    \begin{itemize}
        \item If $I-\mathbf A^T$ is nonsingular, SEM can be written as
        \begin{equation}
            \mathbf X = (\mathbf I - \mathbf A^T)^{-1} \mathbf \epsilon
        \end{equation}
        \item Usually we assume $\mathbf \epsilon \sim \mathcal N(\mathbf 0, \mathbf \Sigma)$ and
        \begin{equation}
            \mathbf X \sim \mathcal N(\mathbf 0, (\mathbf I - \mathbf A^T)^{-1} \mathbf \Sigma (\mathbf I - \mathbf A)^{-1})
        \end{equation}
    \end{itemize}
\end{frame}

\begin{frame}{Semi-Markovian DAG}
    \begin{itemize}
        \item Semi-Markovian DAG is a DAG where the noise variables have dependencies \footcite{Shipster2006semimarkovian}
        \item The noise vector $\mathbf \epsilon$ is a semi-Markovian noise vector if
        \begin{equation}
            \Sigma \in \mathcal{W}(G) := \{ \Sigma \in \mathbb R^{m\times m} : \Sigma_{ij} = \Sigma_{ji} = 0 \text{ if } i \notin \text{sib}(j) \}
        \end{equation}
        where $\text{sib}(j) = \{i : i \sim j \text{ in } G\}$ \footcite{wang2017empirical}
    \end{itemize}
\end{frame}

\begin{frame}{DAG Learning}
    \begin{itemize}
        \item DAG learning is a problem of estimating $\mathcal G$ from $\mathbf X$
        \item NP-hard problem \footcite{ChickeringNPhard}
        \item Score-based methods and constraint-based methods are used
        \item NOTEARS : conversion of DAG learning to continuous optimization problem \footcite{zheng2018dags}
        \begin{equation}
            tr(A)
        \end{equation}
    \end{itemize}
\end{frame}

\subsection*{VAE and Normalizing Flow}

\begin{frame}{VAE}
    \begin{itemize}
        \item Variational Autoencoder(VAE) is a generative model that learns a latent variable model
        \item It is trained by maximizing the evidence lower bound(ELBO)
        \begin{equation}
            \mathcal L(\theta, \phi; \mathbf x) = \mathbb E_{q_\phi(\mathbf z | \mathbf x)}[\log p_\theta(\mathbf x | \mathbf z)] - \text{KL}(q_\phi(\mathbf z | \mathbf x) || p(\mathbf z))
        \end{equation}
        \item $q_\phi(\mathbf z | \mathbf x)$ is an approximate posterior and $p_\theta(\mathbf x | \mathbf z)$ is a generative model
    \end{itemize}
\end{frame}

\begin{frame}{Normalizing Flow}
    \begin{itemize}
        \item Normalizing flow transforms a simple distribution to a complex distribution \footcite{rezende15normalizingflows}
        \item It is a sequence of invertible transformations $f_t : \mathbb R^d \to \mathbb R^d$
        \begin{equation}
            z_T = f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)
        \end{equation}
        \item The probability density function of $z_T$ is
        \begin{equation}
            p(z_T) = p(z_0) \left| \det \left( \frac{\partial f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)}{\partial z_0} \right) \right|
        \end{equation}
    \end{itemize}
\end{frame}

\begin{frame}{Linear IAF}
    \begin{itemize}
        \item Inverse Autoregressive Flow(IAF) is a normalizing flow that uses autoregressive transformation \footcite{kingma2016iaf}
        \item Linear IAF is a special case of IAF where $f_t$ is a linear transformation
        \begin{equation}
            \bf z_t = \bf \mu_t + \bf \sigma_t \odot \bf z_{t-1}.
        \end{equation}
        \item The probability density function of $z_T$ is
        \begin{equation}
            \mathbf{z}_T = \mathbf{L(x)} \cdot \mathbf{z}_0.
        \end{equation}
        where $\mathbf{L(x)}$ is a lower triangular matrix from encoder network.
        \item We can use more flexible prior distribution by using normalizing flow.
    \end{itemize}
\end{frame}

\section{Method}

\begin{frame}{Model}
    Overall architecture of our model is shown in Figure below.
    \begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{fig/model.pdf}
        \caption{Overall architecture of our model}
        \label{fig:architecture}
    \end{figure}
\end{frame}

\begin{frame}{Learning}
    \begin{itemize}
        \item We use the same optimization procedure as DAG-GNN \footcite{yu2019daggnn}.
        \item The optimization procedure is minimizing the following loss function:
        \begin{equation}
            \mathcal{L}(A, W, \lambda) =-\mathcal{L}_{\mathrm{ELBO}} + \tau \Vert A \Vert_1 + \lambda h(A) + \frac{c}{2} |h(A)|^2.
        \end{equation}
        \item The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
        \item The third and fourth terms are the augmented Lagrangian terms.
        \item Gradually increasing the value of $c$ during the training process results the acyclicity constraint to zero.
    \end{itemize}
\end{frame}

\section{Experiment}

\begin{frame}{Experiment}
    \begin{itemize}
        \item We compare our method with the DAG-GNN \footcite{yu2019daggnn} with random graph datasets.
        \item thresholding value of extracting graph as 0.3
        \item Erdős–Rényi random graph with 5000 samples
        \item Node size : 10, 20, 30, and 50 nodes
        \item Noise type : (Independent) Gaussian, Laplace, Exponential and (Dependent) Gaussian
    \end{itemize}
\end{frame}

\begin{frame}{Experiment(Cont'd)}
    \begin{itemize}
        \item We evaluate the performance of our method using two metrics: Structural Hamming Distance (SHD), False Discovery Rate (FDR) with respect to the number of predicted edges.
        \item SHD measures the number of edge additions, deletions, and reversals required to transform the estimated graph into the true graph.
        \item FDR represents the ratio of false positives to the total number of predicted edges.
        \item These metrics are calculated by comparing the estimated graph with the true DAG.
        \item For each combination, we generate at least 5 random graphs and calculate the average metrics.
    \end{itemize}
\end{frame}

\section{Result}

\begin{frame}{Independent Noise}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_normal.pdf}
                \caption{SHD with normal}
                \label{fig:ind_gaussian_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_normal.pdf}
                \caption{FDR with normal}
                \label{fig:ind_gaussian_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    \begin{itemize}
        \item In the case of independent Gaussian noise, DAG-GNN(orange) performs better than our method(blue).
    \end{itemize}
\end{frame}

\begin{frame}{Independent Noise(Cont'd)}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_laplace.pdf}
                \caption{SHD with Laplace}
                \label{fig:ind_laplace_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_laplace.pdf}
                \caption{FDR with Laplace}
                \label{fig:ind_laplace_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    In the case of independent Laplace noise, our method(blue) performs better than DAG-GNN(orange).
\end{frame}

\begin{frame}{Independent Noise(Cont'd)}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_independence_exponential.pdf}
                \caption{SHD with Exponential}
                \label{fig:ind_exponential_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_independence_exponential.pdf}
                \caption{FDR with Exponential}
                \label{fig:ind_exponential_fdr}
            \end{figure}
        \end{column}
    \end{columns}
    In the case of independent Exponential noise, our method(blue) performs better than DAG-GNN(orange).
\end{frame}

\begin{frame}{Independent Noise(Cont'd)}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_laplace.pdf}
                \caption{Number of predicted edges with Laplace}
                \label{fig:ind_laplace_edges}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/Predicted Edges_independence_exponential.pdf}
                \caption{Number of predicted edges with Exponential}
                \label{fig:ind_exponential_edges}
            \end{figure}
        \end{column}
    \end{columns}
    Considering the number of predicted edges with the SHD and FDR, our method(blue) performs better than DAG-GNN(orange).
\end{frame}

\begin{frame}
    \begin{figure}
        \centering
        \includegraphics[height=3.5cm]{fig/comparison_indep_30_laplace_seed301.pdf}
        \label{fig:ind_exponential_comparison}
    \end{figure}
    \begin{figure}
        \centering
        \includegraphics[height=3.5cm]{fig/comparison_indep_30_exponential_seed301.pdf}
        \label{fig:ind_exponential_comparison}
        \captionsetup{justification=centering}
        \caption{Comparison of estimated graphs with Laplace(above) and Exponential(below) noise}
    \end{figure}
\end{frame}

\begin{frame}{Dependent Noise}
    Proportion of bidirectional edges fixed to 0.3.
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[width=\textwidth]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Proportion of bidirectional edges fixed to 0.3.
    \begin{figure}
        \centering
        \includegraphics[height=4cm]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
        \caption{Number of predicted edges with dependent Gaussian}
        \label{fig:dep_gaussian_edges}
    \end{figure}
    \begin{itemize}
        \item In the case of dependent Gaussian noise, our method(blue) performs better than DAG-GNN(orange).
        \item Considering the stable number of predicted edges with the SHD and FDR, our method(blue) outperforms.
    \end{itemize}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=10$ : similar performance between DAG-GNN(orange) and our method(blue).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_10_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_10}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_10_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_10}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=20$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_20_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_20}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_20_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_20}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=30$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_30}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_30}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Dependent Noise(Cont'd)}
    Node size $m=50$ : our method(blue) performs better than DAG-GNN(orange).
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/SHD_dependence_50_DAG_threshold0.3.pdf}
                \caption{SHD with dependent Gaussian}
                \label{fig:dep_gaussian_shd_50}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[height=4cm]{fig/FDR_dependence_50_DAG_threshold0.3.pdf}
                \caption{FDR with dependent Gaussian}
                \label{fig:dep_gaussian_fdr_50}
            \end{figure}
        \end{column}
    \end{columns}
\end{frame}

\begin{frame}{Number of predicted edges}
    \begin{columns}
        \begin{column}{0.5\textwidth}
            \begin{figure}
                \centering
                \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_10_DAG_threshold0.3.pdf}
                \caption{$m=10$}
                \label{fig:dep_gaussian_edges_10}
            \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
            \begin{figure}
            \centering
            \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_20_DAG_threshold0.3.pdf}
            \caption{$m=20$}
            \label{fig:dep_gaussian_edges_20}
            \end{figure}
        \end{column}
    \end{columns}
    \begin{columns}
        \begin{column}{0.5\textwidth}
        \begin{figure}
            \centering
            \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
            \caption{$m=30$}
            \label{fig:dep_gaussian_edges_30}
        \end{figure}
        \end{column}
        \begin{column}{0.5\textwidth}
        \begin{figure}
            \centering
            \includegraphics[scale=0.3]{fig/Predicted Edges_dependence_50_DAG_threshold0.3.pdf}
            \caption{$m=50$}
            \label{fig:dep_gaussian_edges_50}
        \end{figure}
        \end{column}
    \end{columns}    
\end{frame}

\section{Conclusion}

\begin{frame}{Conclusion}
    \begin{itemize}
        \item Proposed a method to learn the structure of a semi-Markovian DAG using a flow-based VAE.
        \item Conducted experiments on simulated data and compared the performance of our method with that of DAG-GNN.
        \item Our method outperforms DAG-GNN in terms of both SHD and FDR metrics, especially when the noise variables have a dependent structure and when the size of the graph is large.
        \item In terms of the number of predicted edges, our method results in fewer edges than DAG-GNN while maintaining the same or better level of SHD and FDR. 
    \end{itemize}
\end{frame}

\begin{frame}{Conclusion(Cont'd)}
    \begin{itemize}
        \item However, contrary to our initial expectations, the lower triangular matrix $\mathbf L$ learned in the IAF layer did not capture the covariance matrix of the actual noise variables.
        \item This seems to be due to the entanglement problem in the latent space, and resolving this issue remains a future research task.
        \item If the full covariance matrix of the noise variables could be found, it could be possible to directly identify bidirectional edges in the graph, enabling more accurate learning of the semi-Markovian graph.
    \end{itemize} 
\end{frame}

\begin{frame}[allowframebreaks]{References}
\printbibliography
\end{frame}
\end{document}

### 2023-12-10 model_newIAF.py

@@ -49,7 +49,7 @@ class Encoder(nn.Module):
         self.args = args
         self.adj_A = nn.Parameter(torch.from_numpy(adj_A).double(), requires_grad=True)
         self.Wa = nn.Parameter(torch.zeros(args.z_dims), requires_grad=True)
-        self.fc1 = nn.Linear(args.x_dims, args.encoder_hidden, bias=True)
+        self.fc1 = nn.Linear(args.z_dims, args.encoder_hidden, bias=True)
         self.fc2 = nn.Linear(args.encoder_hidden, args.z_dims, bias=True)
         self.batch_size = args.batch_size
         self.dropout = nn.Dropout(p=args.encoder_dropout)
@@ -84,10 +84,10 @@ class Encoder(nn.Module):
 
         adj_A = torch.eye(adj_A1.size()[0]).float()
 
-        logits = torch.matmul(adj_Aforz, inputs.float() + self.Wa) - self.Wa
+        logits = torch.matmul(adj_Aforz, inputs.float() + self.Wa) - self.Wa # B x L x 1
 
-        h0 = F.relu((self.fc1(logits))) # first hidden layer
-        h1 = (self.fc2(h0))
+        h0 = F.relu((self.fc1(logits))) # B x L x 1
+        h1 = (self.fc2(h0)) # B x L x 1
 
         # For Flow layer
         z_q_mean = self.encoder_mean(h1)
@@ -160,10 +160,31 @@ class VAE_IAF_SEM(nn.Module):
             self.encoder_L = nn.Linear(args.encoder_hidden, args.z_size)
 
         self.softmax = nn.Softmax()
+        self.bf1 = nn.Linear(args.z_size, args.z_size)
+        self.bf2 = nn.Linear(args.z_size, args.z_size)
 
         for m in self.modules():
             if isinstance(m, nn.Linear):
                 nn.init.xavier_normal_(m.weight.data)
+
+    def before_flow(self, logits):
+        """
+        :param logits: B x L x 1
+        :return: L : B x L x L
+        """
+        L0 = self.encoder_L(logits) # B x L x L
+        L1 = self.softmax(L0)
+        L2 = F.tanh(self.bf1(L1))
+
+        logits_squeeze = logits.squeeze(2) # B x L
+
+        ZTZ = torch.matmul(logits_squeeze.transpose(0, 1), logits_squeeze) # L x L
[…]

### 2023-12-10 train_new.py

@@ -603,5 +603,5 @@ for _, spine in axes[1].spines.items():
     spine.set_visible(True)
 axes[1].set_xticks([])
 axes[1].set_yticks([])
-fig.tight_layout(rect=[0, 0, .9, 1])
+# fig.tight_layout(rect=[0, 0, .9, 1])
 plt.savefig(folder + '/cov.png', dpi=300)

### 2023-12-10 References/DAGGNN/src/modules.py
레퍼런스코드 제거

@@ -1,350 +1,0 @@
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import math
-
-from torch.autograd import Variable
-from utils import my_softmax, get_offdiag_indices, gumbel_softmax, preprocess_adj, preprocess_adj_new, preprocess_adj_new1, gauss_sample_z, my_normalize
-
-_EPS = 1e-10
-
-
-class MLPEncoder(nn.Module):
-    """MLP encoder module."""
-    def __init__(self, n_in, n_xdims, n_hid, n_out, adj_A, batch_size, do_prob=0., factor=True, tol = 0.1):
-        super(MLPEncoder, self).__init__()
-
-        self.adj_A = nn.Parameter(Variable(torch.from_numpy(adj_A).double(), requires_grad=True))
-        self.factor = factor
-
-        self.Wa = nn.Parameter(torch.zeros(n_out), requires_grad=True)
-        self.fc1 = nn.Linear(n_xdims, n_hid, bias = True)
-        self.fc2 = nn.Linear(n_hid, n_out, bias = True)
-        self.dropout_prob = do_prob
-        self.batch_size = batch_size
-        self.z = nn.Parameter(torch.tensor(tol))
-        self.z_positive = nn.Parameter(torch.ones_like(torch.from_numpy(adj_A)).double())
-        self.init_weights()
-
-    def init_weights(self):
-        for m in self.modules():
-            if isinstance(m, nn.Linear):
-                nn.init.xavier_normal_(m.weight.data)
-            elif isinstance(m, nn.BatchNorm1d):
-                m.weight.data.fill_(1)
-                m.bias.data.zero_()
-
-
-    def forward(self, inputs, rel_rec, rel_send):
-
-        # inputs : batch_size * n_node * n_xdims
-
-        if torch.sum(self.adj_A != self.adj_A):
-            print('nan error \n')
-
-        # to amplify the value of A and accelerate convergence.
-        adj_A1 = torch.sinh(3.*self.adj_A)
-
-        # adj_Aforz = I-A^T
-        adj_Aforz = preprocess_adj_new(adj_A1)
-
-        adj_A = torch.eye(adj_A1.size()[0]).double()
-        H1 = F.relu((self.fc1(inputs)))
-        x = (self.fc2(H1))
[…]

### 2023-12-10 References/DAGGNN/src/train.py

@@ -1,565 +1,0 @@
-
-''''
-Main function for traininng DAG-GNN
-
-'''
-
-
-from __future__ import division
-from __future__ import print_function
-
-import time
-import argparse
-import pickle
-import os
-import datetime
-from torchviz import make_dot
-
-# import torch
-import torch.optim as optim
-from torch.optim import lr_scheduler
-import math
-
-# import numpy as np
-from utils import *
-from modules import *
-
-parser = argparse.ArgumentParser()
-
-# -----------data parameters ------
-# configurations
-parser.add_argument('--data_type', type=str, default= 'synthetic',
-                    choices=['synthetic', 'discrete', 'real'],
-                    help='choosing which experiment to do.')
-parser.add_argument('--data_filename', type=str, default= 'alarm',
-                    help='data file name containing the discrete files.')
-parser.add_argument('--data_dir', type=str, default= 'data/',
-                    help='data file name containing the discrete files.')
-parser.add_argument('--data_sample_size', type=int, default=5000,
-                    help='the number of samples of data')
-parser.add_argument('--data_variable_size', type=int, default=10,
-                    help='the number of variables in synthetic generated data')
-parser.add_argument('--graph_type', type=str, default='erdos-renyi',
-                    help='the type of DAG graph by generation method')
-parser.add_argument('--graph_degree', type=int, default=2,
-                    help='the number of degree in generated DAG graph')
-parser.add_argument('--graph_sem_type', type=str, default='linear-gauss',
-                    help='the structure equation model (SEM) parameter type')
-parser.add_argument('--graph_linear_type', type=str, default='nonlinear_2',
-                    help='the synthetic data type: linear -> linear SEM, nonlinear_1 -> x=Acos(x+1)+z, nonlinear_2 -> x=2sin(A(x+0.5))+A(x+0.5)+z')
-parser.add_argument('--edge-types', type=int, default=2,
-                    help='The number of edge types to infer.')
-parser.add_argument('--x_dims', type=int, default=1, #changed here
-                    help='The number of input dimensions: default 1.')
-parser.add_argument('--z_dims', type=int, default=1,
-                    help='The number of latent variable dimensions: default the same as variable size.')
-
-# -----------training hyperparameters
-parser.add_argument('--optimizer', type = str, default = 'Adam',
[…]

### 2023-12-10 References/DAGGNN/src/utils.py

@@ -1,757 +1,0 @@
-import numpy as np
-import torch
-from torch.utils.data.dataset import TensorDataset
-from torch.utils.data import DataLoader
-import torch.nn.functional as F
-import torch.nn as nn
-from torch.autograd import Variable
-import numpy as np
-import scipy.linalg as slin
-import scipy.sparse as sp
-import networkx as nx
-import pandas as pd
-from pandas import ExcelWriter
-from pandas import ExcelFile
-import os
-import glob
-import re
-import pickle
-import math
-from torch.optim.adam import Adam
-
-# data generating functions
-
-def simulate_random_dag(d: int,
-                        degree: float,
-                        graph_type: str,
-                        w_range: tuple = (0.5, 2.0)) -> nx.DiGraph:
-    """Simulate random DAG with some expected degree.
-
-    Args:
-        d: number of nodes
-        degree: expected node degree, in + out
-        graph_type: {erdos-renyi, barabasi-albert, full}
-        w_range: weight range +/- (low, high)
-
-    Returns:
-        G: weighted DAG
-    """
-    if graph_type == 'erdos-renyi':
-        prob = float(degree) / (d - 1)
-        B = np.tril((np.random.rand(d, d) < prob).astype(float), k=-1)
-    elif graph_type == 'barabasi-albert':
-        m = int(round(degree / 2))
-        B = np.zeros([d, d])
-        bag = [0]
-        for ii in range(1, d):
-            dest = np.random.choice(bag, size=m)
-            for jj in dest:
-                B[ii, jj] = 1
-            bag.append(ii)
-            bag.extend(dest)
-    elif graph_type == 'full':  # ignore degree, only for experimental use
-        B = np.tril(np.ones([d, d]), k=-1)
-    else:
-        raise ValueError('unknown graph type')
-    # random permutation
-    P = np.random.permutation(np.eye(d, d))  # permutes first axis only
-    B_perm = P.T.dot(B).dot(P)
-    U = np.random.uniform(low=w_range[0], high=w_range[1], size=[d, d])
-    U[np.random.rand(d, d) < 0.5] *= -1
-    W = (B_perm != 0).astype(float) * U
-    G = nx.DiGraph(W)
-    return G
-
-
-def simulate_sem(G: nx.DiGraph,
-                 n: int, x_dims: int,
[…]

### 2023-12-10 References/Flow-VAE/dynamicflow.py

@@ -1,620 +1,0 @@
-import argparse
-
-import torch
-import torch.nn as nn
-import torch.nn.functional as F
-import torch.distributions as distributions
-import torch.optim as optim
-
-import torchvision
-import torchvision.transforms as transforms
-import torchvision.utils as utils
-import numpy as np
-import matplotlib.pyplot as plt
-
-class PlanarFlow(nn.Module):
-    def __init__(self, dim):
-        """Instantiates one step of planar flow.
-
-        Args:
-            dim: input dimensionality.
-        """
-        super(PlanarFlow, self).__init__()
-
-        self.linear_u = nn.Linear(dim, dim)
-        self.linear_w = nn.Linear(dim, dim)
-        self.linear_b = nn.Linear(dim, 1)
-
-    def forward(self, x, v):
-        """Forward pass.
-
-        Args:
-            x: input tensor (B x D).
-            v: output from last layer of encoder (B x D).
-        Returns:
-            transformed x and log-determinant of Jacobian.
-        """
-        u, w, b = self.linear_u(v), self.linear_w(v), self.linear_b(v)
-
-        def m(x):
-            return F.softplus(x) - 1.
-        def h(x):
-            return torch.tanh(x)
-        def h_prime(x):
-            return 1. - h(x)**2
-
-        inner = (w * u).sum(dim=1, keepdim=True)
-        u = u + (m(inner) - inner) * w / (w * w).sum(dim=1, keepdim=True)
-        activation = (w * x).sum(dim=1, keepdim=True) + b
-        x = x + u * h(activation)
-        psi = h_prime(activation) * w
-        log_det = torch.log(torch.abs(1. + (u * psi).sum(dim=1, keepdim=True)))
-
-        return x, v, log_det
-
-class RadialFlow(nn.Module):
-    def __init__(self, dim):
-        """Instantiates one step of radial flow.
-
-        Args:
-            dim: input dimensionality.
-        """
-        super(RadialFlow, self).__init__()
-
-        self.linear_a = nn.Linear(dim, 1)
-        self.linear_b = nn.Linear(dim, 1)
-        self.linear_c = nn.Linear(dim, dim)
-        self.d = dim
-
-    def forward(self, x, v):
-        """Forward pass.
-
-        Args:
-            x: input tensor (B x D).
-            v: output from last layer of encoder (B x D).
-        Returns:
-            transformed x and log-determinant of Jacobian.
-        """
-        a, b, c = self.linear_a(v), self.linear_b(v), self.linear_c(v)
-
[…]

### 2023-12-10 References/Flow-VAE/staticflow.py

@@ -1,679 +1,0 @@
-import argparse
-
-import torch
-import torch.nn as nn
-import torch.nn.init as init
-import torch.nn.functional as F
-import torch.distributions as distributions
-import torch.optim as optim
-
-import torchvision
-import torchvision.transforms as transforms
-import torchvision.utils as utils
-import numpy as np
-import matplotlib.pyplot as plt
-
-class PlanarFlow(nn.Module):
-    def __init__(self, dim):
-        """Instantiates one step of planar flow.
-
-        Reference:
-        Variational Inference with Normalizing Flows
-        Danilo Jimenez Rezende, Shakir Mohamed
-        (https://arxiv.org/abs/1505.05770)
-
-        Args:
-            dim: input dimensionality.
-        """
-        super(PlanarFlow, self).__init__()
-
-        self.u = nn.Parameter(torch.randn(1, dim))
-        self.w = nn.Parameter(torch.randn(1, dim))
-        self.b = nn.Parameter(torch.randn(1))
-
-    def forward(self, x):
-        """Forward pass.
-
-        Args:
-            x: input tensor (B x D).
-        Returns:
-            transformed x and log-determinant of Jacobian.
-        """
-        def m(x):
-            return F.softplus(x) - 1.
-        def h(x):
-            return torch.tanh(x)
-        def h_prime(x):
-            return 1. - h(x)**2
-
-        inner = (self.w * self.u).sum()
-        u = self.u + (m(inner) - inner) * self.w / self.w.norm()**2
-        activation = (self.w * x).sum(dim=1, keepdim=True) + self.b
-        x = x + u * h(activation)
-        psi = h_prime(activation) * self.w
-        log_det = torch.log(torch.abs(1. + (u * psi).sum(dim=1, keepdim=True)))
-
-        return x, log_det
-
-class RadialFlow(nn.Module):
-    def __init__(self, dim):
-        """Instantiates one step of radial flow.
-
-        Reference:
-        Variational Inference with Normalizing Flows
-        Danilo Jimenez Rezende, Shakir Mohamed
-        (https://arxiv.org/abs/1505.05770)
-
-        Args:
-            dim: input dimensionality.
-        """
-        super(RadialFlow, self).__init__()
-
-        self.a = nn.Parameter(torch.randn(1))
-        self.b = nn.Parameter(torch.randn(1))
-        self.c = nn.Parameter(torch.randn(1, dim))
-        self.d = dim
-
[…]

### 2023-12-10 References/IAF-VAE/layers.py

@@ -1,146 +1,0 @@
-import torch
-import numpy as np
-import torch.nn as nn
-import torch.nn.functional as F
-import torch.nn.utils.weight_norm as wn
-import torch.distributions as D
-
-# Basic Layers 
-# -------------------------------------------------------------------------------------------------------
-
-# taken from https://github.com/jzbontar/pixelcnn-pytorch
-class MaskedConv2d(nn.Conv2d):
-    def __init__(self, mask_type, *args, **kwargs):
-        super(MaskedConv2d, self).__init__(*args, **kwargs)
-        assert mask_type in {'A', 'B'}
-        self.register_buffer('mask', self.weight.data.clone())
-        _, _, kH, kW = self.weight.size()
-        self.mask.fill_(1)
-        self.mask[:, :, kH // 2, kW // 2 + (mask_type == 'B'):] = 0
-        self.mask[:, :, kH // 2 + 1:] = 0
-
-    def forward(self, x):
-        self.weight.data *= self.mask
-        return super(MaskedConv2d, self).forward(x)
-
-class ARMultiConv2d(nn.Module):
-    def __init__(self, n_h, n_out, args, nl=F.elu):
-        super(ARMultiConv2d, self).__init__()
-        self.nl = nl
-
-        convs, out_convs = [], []
-
-        for i, size in enumerate(n_h):
-            convs     += [MaskedConv2d('A' if i == 0 else 'B', args.z_size if i == 0 else args.h_size, args.h_size, 3, 1, 1)]
-        for i, size in enumerate(n_out):
-            out_convs += [MaskedConv2d('B', args.h_size, args.z_size, 3, 1, 1)]
-
-        self.convs = nn.ModuleList(convs)
-        self.out_convs = nn.ModuleList(out_convs)
-
-
-    def forward(self, x, context):
-        for i, conv_layer in enumerate(self.convs):
-            x = conv_layer(x)
-            if i == 0: 
-                x += context
-            x = self.nl(x)
-
-        return [conv_layer(x) for conv_layer in self.out_convs]
-
-
-# IAF building block
-# -------------------------------------------------------------------------------------------------------
-
-class IAFLayer(nn.Module):
-    def __init__(self, args, downsample):
-        super(IAFLayer, self).__init__()
-        n_in  = args.h_size
-        n_out = args.h_size * 2 + args.z_size * 2
-        
[…]

### 2023-12-10 References/IAF-VAE/main.py

@@ -1,249 +1,0 @@
-import torch
-import argparse
-import numpy as np
-import torch.nn as nn
-import torch.nn.functional as F
-
-from os.path import join
-from tensorboardX import SummaryWriter
-from torchvision.utils import save_image
-from collections import OrderedDict as OD 
-from torchvision import datasets, transforms, utils
-
-from layers import IAFLayer
-from utils  import * 
-
-# Model definition
-# ----------------------------------------------------------------------------------------------
-class VAE(nn.Module):
-    def __init__(self, args):
-        super(VAE, self).__init__()
-        self.register_parameter('h', torch.nn.Parameter(torch.zeros(args.h_size)))
-        self.register_parameter('dec_log_stdv', torch.nn.Parameter(torch.Tensor([0.])))
-
-        layers = []
-        # build network
-        for i in range(args.depth):
-            layer = []
-
-            for j in range(args.n_blocks):
-                downsample = (i > 0) and (j == 0)
-                layer += [IAFLayer(args, downsample)]
-
-            layers += [nn.ModuleList(layer)]
-
-        self.layers = nn.ModuleList(layers) 
-        
-        self.first_conv = nn.Conv2d(3, args.h_size, 4, 2, 1)
-        self.last_conv = nn.ConvTranspose2d(args.h_size, 3, 4, 2, 1)
-
-    def forward(self, input):
-        # assumes input is \in [-0.5, 0.5] 
-        x = self.first_conv(input)
-        kl, kl_obj = 0., 0.
-
-        h = self.h.view(1, -1, 1, 1)
-
-        for layer in self.layers:
-            for sub_layer in layer:
-                x = sub_layer.up(x)
-
-        h = h.expand_as(x)
-        self.hid_shape = x[0].size()
-
-        for layer in reversed(self.layers):
-            for sub_layer in reversed(layer):
-                h, curr_kl, curr_kl_obj = sub_layer.down(h)
-                kl     += curr_kl
-                kl_obj += curr_kl_obj
-
-        x = F.elu(h)
-        x = self.last_conv(x)
-        
-        x = x.clamp(min=-0.5 + 1. / 512., max=0.5 - 1. / 512.)
-
-        return x, kl, kl_obj
-
-
-    def sample(self, n_samples=64):
-        h = self.h.view(1, -1, 1, 1)
-        h = h.expand((n_samples, *self.hid_shape))
-        
-        for layer in reversed(self.layers):
-            for sub_layer in reversed(layer):
[…]

### 2023-12-10 References/IAF-VAE/utils.py

@@ -1,85 +1,0 @@
-import os
-import json
-import torch
-import numpy as np
-import torch.nn as nn
-import torch.nn.functional as F
-import torch.distributions as D
-import torch.nn.utils.weight_norm as wn
-from collections import OrderedDict as OD
-from collections import defaultdict as DD
-
-class dotdict(dict):
-    __getattr__ = dict.get
-    __setattr__ = dict.__setitem__
-    __delattr__ = dict.__delitem__
-
-def print_and_save_args(args, path):
-    print(args)
-    # let's save the args as json to enable easy loading
-    with open(os.path.join(path, 'args.json'), 'w') as f: 
-        json.dump(vars(args), f)
-
-def load_model_from_file(path):
-    with open(os.path.join(path, 'args.json'), 'rb') as f:
-        args = dotdict(json.load(f))
-
-    # create model
-    from main import VAE
-    model = VAE(args)
-
-    # load weights
-    model.load_state_dict(torch.load(os.path.join(path, 'best_model.pth')))
-
-    return model
-
-def set_seed(seed):
-    np.random.seed(0)
-    torch.manual_seed(0)
-    torch.backends.cudnn.benchmark = False
-    torch.backends.cudnn.deterministic = True
-
-def print_and_log_scalar(writer, name, value, write_no, end_token=''):
-    if isinstance(value, list):
-        if len(value) == 0: return
-        
-        str_tp = str(type(value[0]))
-        if type(value[0]) == torch.Tensor:
-            value = torch.mean(torch.stack(value))
-        elif 'float' in str_tp or 'int' in str_tp:
-            value = sum(value) / len(value)
-    zeros = 40 - len(name) 
-    name += ' ' * zeros
-    print('{} @ write {} = {:.4f}{}'.format(name, write_no, value, end_token))
-    writer.add_scalar(name, value, write_no)
-
-def maybe_create_dir(path):
-    if not os.path.exists(path):
-        os.makedirs(path)
-
-
-# create logging containers
-def reset_log():
-    # TODO: make it such that if key is not in dict, it's init with incoming content
-    # this way, we won't have to know in advance what we want to monitor
-    return DD(list)
-    logs = OD()
-    for name in ['inner log p(x|z)', 'log p(x|z)', 'log p(x|z) nn', 'commit', 'vq', 'kl', 'bpd', 'elbo']:
-        logs[name] = []
-    return logs
-
-
-# loss functions 
[…]

### 2023-12-10 References/NormalizingFlow/flow.py

@@ -1,155 +1,0 @@
-import torch
-import torch.nn as nn
-from tqdm import tqdm
-import numpy as np
-import matplotlib.pyplot as plt
-import math
-
-
-torch.manual_seed(0)
-
-def gaussian_log_pdf(z):
-    """
-    Arguments:
-    ----------
-    - z: a batch of m data points (size: m x data_dim)
-    """
-    return -.5 * (torch.log(torch.tensor([math.pi * 2], device=z.device)) + z ** 2).sum(1)
-
-class PlanarFlow(nn.Module):
-
-    def __init__(self, data_dim):
-        super().__init__()
-
-        self.u = nn.Parameter(torch.rand(data_dim))
-        self.w = nn.Parameter(torch.rand(data_dim))
-        self.b = nn.Parameter(torch.rand(1))
-        self.h = nn.Tanh()
-        self.h_prime = lambda z: (1 - self.h(z) ** 2)
-
-    def constrained_u(self):
-        """
-        Constrain the parameters u to ensure invertibility
-        """
-        wu = torch.matmul(self.w.T, self.u)
-        m = lambda x: -1 + torch.log(1 + torch.exp(x))
-        return self.u + (m(wu) - wu) * (self.w / (torch.norm(self.w) ** 2 + 1e-15))
-
-    def forward(self, z):
-        u = self.constrained_u()
-        hidden_units = torch.matmul(self.w.T, z.T) + self.b
-        x = z + u.unsqueeze(0) * self.h(hidden_units).unsqueeze(-1)
-        psi = self.h_prime(hidden_units).unsqueeze(0) * self.w.unsqueeze(-1)
-        log_det = torch.log((1 + torch.matmul(u.T, psi)).abs() + 1e-15)
-        return x, log_det
-
-class NormalizingFlow(nn.Module):
-
-    def __init__(self, flow_length, data_dim):
-        super().__init__()
-
-        self.layers = nn.Sequential(
-            *(PlanarFlow(data_dim) for _ in range(flow_length)))
-
-    def forward(self, z):
-        log_jacobians = 0
-        for layer in self.layers:
-            z, log_jacobian = layer(z)
-            log_jacobians += log_jacobian
-        return z, log_jacobians
-
-def train(flow, optimizer, nb_epochs, log_density, batch_size, data_dim):
-    training_loss = []
[…]

### 2023-12-10 References/VAE_VPFLOWS/experiment.py

@@ -1,156 +1,0 @@
-from __future__ import print_function
-import argparse
-
-import torch
-import torch.optim as optim
-
-import os
-
-from utils.load_data import load_dataset
-
-# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-# # # # # # # # # # #
-# START EXPERIMENTS # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # # #
-# # # # # # # # # # #
-
-# Training settings
-parser = argparse.ArgumentParser(description='VAE+Volume-Preserving Flows')
-# arguments for optimization
-parser.add_argument('--batch_size', type=int, default=100, metavar='N',
-                    help='input batch size for training (default: 100)')
-parser.add_argument('--test_batch_size', type=int, default=100, metavar='N',
-                    help='input batch size for testing (default: 100)')
-parser.add_argument('--epochs', type=int, default=1, metavar='N',
-                    help='number of epochs to train (default: 2000)')
-parser.add_argument('--lr', type=float, default=0.0005, metavar='LR',
-                    help='learning rate (default: 0.0005)')
-parser.add_argument('--early_stopping_epochs', type=int, default=50, metavar='N',
-                    help='number of epochs for early stopping')
-
-parser.add_argument('--warmup', type=int, default=100, metavar='N',
-                    help='number of epochs for warm-up')
-
-# cuda
-parser.add_argument('--no-cuda', action='store_true', default=False,
-                    help='enables CUDA training')
-# random seed
-parser.add_argument('--seed', type=int, default=1, metavar='S',
-                    help='random seed (default: 1)')
-# model: latent size, input_size
-parser.add_argument('--z1_size', type=int, default=2, metavar='N',
-                    help='latent size')
-parser.add_argument('--z2_size', type=int, default=0, metavar='N',
-                    help='latent size')
-parser.add_argument('--input_size', type=int, default=[1, 28, 28], metavar='N',
-                    help='input size')
-
-parser.add_argument('--activation', type=str, default=None, metavar='N',
-                    help='activation function')
-
[…]

### 2023-12-10 References/VAE_VPFLOWS/models/VAE.py

@@ -1,200 +1,0 @@
-from __future__ import print_function
-
-import torch
-import torch.utils.data
-import torch.nn as nn
-import torch.nn.functional as F
-from torch.autograd import Variable
-
-from utils.distributions import log_Bernoulli, log_Normal_diag, log_Normal_standard
-from utils.visual_evaluation import plot_histogram
-
-import numpy as np
-
-from scipy.misc import logsumexp
-
-def xavier_init(m):
-    s =  np.sqrt( 2. / (m.in_features + m.out_features) )
-    m.weight.data.normal_(0, s)
-
-class Gate(nn.Module):
-    def __init__(self):
-        super(Gate, self).__init__()
-
-    def forward(self, h, g):
-        return h * g
-
-#=======================================================================================================================
-class VAE(nn.Module):
-    def __init__(self, args):
-        super(VAE, self).__init__()
-
-        self.args = args
-
-        # encoder: q(z | x)
-        self.q_z_layers_pre = nn.ModuleList()
-        self.q_z_layers_gate = nn.ModuleList()
-
-        self.q_z_layers_pre.append( nn.Linear(np.prod(self.args.input_size), 300) )
-        self.q_z_layers_gate.append( nn.Linear(np.prod(self.args.input_size), 300) )
-
-        self.q_z_layers_pre.append( nn.Linear(300, 300) )
-        self.q_z_layers_gate.append( nn.Linear(300, 300) )
-
-        self.q_z_mean = nn.Linear(300, self.args.z1_size)
-        self.q_z_logvar = nn.Linear(300, self.args.z1_size)
-
-        # decoder: p(x | z)
-        self.p_x_layers_pre = nn.ModuleList()
-        self.p_x_layers_gate = nn.ModuleList()
-
-        self.p_x_layers_pre.append( nn.Linear(self.args.z1_size, 300) )
-        self.p_x_layers_gate.append( nn.Linear(self.args.z1_size, 300) )
-
-        self.p_x_layers_pre.append( nn.Linear(300, 300) )
-        self.p_x_layers_gate.append( nn.Linear(300, 300) )
-
-        self.p_x_mean = nn.Linear(300, np.prod(self.args.input_size))
-
-        self.sigmoid = nn.Sigmoid()
-
-        self.Gate = Gate()
-
-        # Xavier initialization (normal)
[…]

### 2023-12-10 References/VAE_VPFLOWS/models/VAE_HF.py

@@ -1,246 +1,0 @@
-from __future__ import print_function
-
-import torch
-import torch.utils.data
-import torch.nn as nn
-import torch.nn.functional as F
-from torch.autograd import Variable
-
-from utils.distributions import log_Bernoulli, log_Normal_diag, log_Normal_standard
-from utils.visual_evaluation import plot_histogram
-
-import numpy as np
-
-from scipy.misc import logsumexp
-
-def xavier_init(m):
-    s =  np.sqrt( 2. / (m.in_features + m.out_features) )
-    m.weight.data.normal_(0, s)
-
-class HF(nn.Module):
-    def __init__(self):
-        super(HF, self).__init__()
-
-    def forward(self, v, z):
-        '''
-        :param v: batch_size (B) x latent_size (L)
-        :param z: batch_size (B) x latent_size (L)
-        :return: z_new = z - 2* v v_T / norm(v,2) * z
-        '''
-        # v * v_T
-        vvT = torch.bmm( v.unsqueeze(2), v.unsqueeze(1) )  # v * v_T : batch_dot( B x L x 1 * B x 1 x L ) = B x L x L
-        # v * v_T * z
-        vvTz = torch.bmm( vvT, z.unsqueeze(2) ).squeeze(2) # A * z : batchdot( B x L x L * B x L x 1 ).squeeze(2) = (B x L x 1).squeeze(2) = B x L
-        # calculate norm ||v||^2
-        norm_sq = torch.sum( v * v, 1 ) # calculate norm-2 for each row : B x 1
-        norm_sq = norm_sq.expand( norm_sq.size(0), v.size(1) ) # expand sizes : B x L
-        # calculate new z
-        z_new = z - 2 * vvTz / norm_sq # z - 2 * v * v_T  * z / norm2(v)
-        return z_new
-
-class Gate(nn.Module):
-    def __init__(self):
-        super(Gate, self).__init__()
-
-    def forward(self, h, g):
-        return h * g
-
-#=======================================================================================================================
-class VAE(nn.Module):
-    def __init__(self, args):
-        super(VAE, self).__init__()
-
-        self.args = args
-
-        # encoder: q(z | x)
-        self.q_z_layers_pre = nn.ModuleList()
-        self.q_z_layers_gate = nn.ModuleList()
-
[…]

### 2023-12-10 References/VAE_VPFLOWS/models/VAE_ccLinIAF.py

@@ -1,263 +1,0 @@
-from __future__ import print_function
-
-import torch
-import torch.utils.data
-import torch.nn as nn
-import torch.nn.functional as F
-from torch.autograd import Variable
-
-from utils.distributions import log_Bernoulli, log_Normal_diag, log_Normal_standard
-from utils.visual_evaluation import plot_histogram
-
-import numpy as np
-
-from scipy.misc import logsumexp
-
-def xavier_init(m):
-    s =  np.sqrt( 2. / (m.in_features + m.out_features) )
-    m.weight.data.normal_(0, s)
-
-class linIAF(nn.Module):
-    def __init__(self,args):
-        super(linIAF, self).__init__()
-        self.args = args
-
-    def forward(self, L, z):
-        '''
-        :param L: batch_size (B) x latent_size^2 (L^2)
-        :param z: batch_size (B) x latent_size (L)
-        :return: z_new = L*z
-        '''
-        # L->tril(L)
-        L_matrix = L.view( -1, self.args.z1_size, self.args.z1_size ) # resize to get B x L x L
-        LTmask = torch.tril( torch.ones(self.args.z1_size, self.args.z1_size), k=-1 ) # lower-triangular mask matrix (1s in lower triangular part)
-        I = Variable( torch.eye(self.args.z1_size, self.args.z1_size).expand(L_matrix.size(0), self.args.z1_size, self.args.z1_size) )
-        if self.args.cuda:
-            LTmask = LTmask.cuda()
-            I = I.cuda()
-        LTmask = Variable(LTmask)
-        LTmask = LTmask.unsqueeze(0).expand( L_matrix.size(0), self.args.z1_size, self.args.z1_size ) # 1 x L x L -> B x L x L
-        LT = torch.mul( L_matrix, LTmask ) + I # here we get a batch of lower-triangular matrices with ones on diagonal
-
-        # z_new = L * z
-        z_new = torch.bmm( LT , z.unsqueeze(2) ).squeeze(2) # B x L x L * B x L x 1 -> B x L
-
-        return z_new
-
-class combination_L(nn.Module):
-    def __init__(self,args):
-        super(combination_L, self).__init__()
-        self.args = args
-
-    def forward(self, L, y):
-        '''
-        :param L: batch_size (B) x latent_size^2 * number_combination (L^2 * C)
-        :param y: batch_size (B) x number_combination (C)
[…]

### 2023-12-10 References/VAE_VPFLOWS/utils/distributions.py

@@ -1,26 +1,0 @@
-from __future__ import print_function
-import torch
-import torch.utils.data
-
-#=======================================================================================================================
-def log_Normal_diag(x, mean, log_var, average=False, dim=None):
-    log_normal = -0.5 * ( log_var + torch.pow( x - mean, 2 ) * torch.pow( torch.exp( log_var ), -1) )
-    if average:
-        return torch.mean( log_normal, dim )
-    else:
-        return torch.sum( log_normal, dim )
-
-def log_Normal_standard(x, average=False, dim=None):
-    log_normal = -0.5 * torch.pow( x , 2 )
-    if average:
-        return torch.mean( log_normal, dim )
-    else:
-        return torch.sum( log_normal, dim )
-
-def log_Bernoulli(x, mean, average=False, dim=None):
-    probs = torch.clamp( mean, min=1e-7, max=1.-1e-7 )
-    log_bernoulli = x * torch.log( probs ) + (1. - x ) * torch.log( 1. - probs )
-    if average:
-        return torch.mean( log_bernoulli, dim )
-    else:
-        return torch.sum( log_bernoulli, dim )

### 2023-12-10 References/VAE_VPFLOWS/utils/evaluation.py

@@ -1,232 +1,0 @@
-from __future__ import print_function
-
-import torch
-from torch.autograd import Variable
-
-import numpy as np
-
-import math
-
-from utils.distributions import log_Normal_diag, log_Normal_standard, log_Bernoulli
-from utils.visual_evaluation import plot_reconstruction, plot_scatter, plot_real, plot_generation, plot_manifold, plot_images, plot_scatter_VPflow
-
-import time
-# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-# ======================================================================================================================
-def evaluate_vae(args, model, train_loader, data_loader, epoch, dir, mode):
-    # set loss to 0
-    evaluate_loss = 0
-    evaluate_re = 0
-    evaluate_kl = 0
-    # set model to evaluation mode
-    model.eval()
-
-    # evaluate
-    for batch_idx, (data, target) in enumerate(data_loader):
-        if args.cuda:
-            data, target = data.cuda(), target.cuda()
-        data, target = Variable(data, volatile=True), Variable(target)
-
-        x = data
-        # forward pass
-        x_mean, x_logvar, z_q, z_q_mean, z_q_logvar = model.forward(x)
-        # loss function
-        # RE
-        RE = log_Bernoulli(x, x_mean)
-
-        # KL
-        log_p_z = log_Normal_standard(z_q, dim=1)
-        log_q_z = log_Normal_diag(z_q, z_q_mean, z_q_logvar, dim=1)
-        KL = - torch.sum(log_p_z - log_q_z)
-
-        evaluate_loss += ((-RE + KL) / data.size(0)).data[0]
-        evaluate_re += (-RE / data.size(0)).data[0]
-        evaluate_kl += (KL / data.size(0)).data[0]
-
-        # print N digits
-        if batch_idx == 1 and mode == 'validation':
-            plot_reconstruction(args, x_mean, epoch, dir, size_x=3, size_y=3)
-            if epoch == 1:
-                # VISUALIZATION: plot real images
-                plot_real(args, data[:26], dir + 'reconstruction/', size_x=3, size_y=3)
-
-    if mode == 'test':
-        # load all data
-        test_data = Variable(data_loader.dataset.data_tensor)
-        test_target = Variable(data_loader.dataset.target_tensor)
[…]

### 2023-12-10 References/VAE_VPFLOWS/utils/load_data.py

@@ -1,168 +1,0 @@
-from __future__ import print_function
-
-import torch
-import torch.utils.data as data_utils
-
-import numpy as np
-
-from scipy.io import loadmat
-import os
-
-import pickle
-# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-# ======================================================================================================================
-def load_static_mnist(args):
-    def lines_to_np_array(lines):
-        return np.array([[int(i) for i in line.split()] for line in lines])
-    with open(os.path.join('datasets', 'MNIST_static', 'binarized_mnist_train.amat')) as f:
-        lines = f.readlines()
-    x_train = lines_to_np_array(lines).astype('float32')
-    with open(os.path.join('datasets', 'MNIST_static', 'binarized_mnist_valid.amat')) as f:
-        lines = f.readlines()
-    x_val = lines_to_np_array(lines).astype('float32')
-    with open(os.path.join('datasets', 'MNIST_static', 'binarized_mnist_test.amat')) as f:
-        lines = f.readlines()
-    x_test = lines_to_np_array(lines).astype('float32')
-
-    # shuffle train data
-    np.random.shuffle(x_train)
-
-    # idle y's
-    y_train = np.zeros( (x_train.shape[0], 1) )
-    y_val = np.zeros( (x_val.shape[0], 1) )
-    y_test = np.zeros( (x_test.shape[0], 1) )
-
-    # pytorch data loader
-    train = data_utils.TensorDataset(torch.from_numpy(x_train), torch.from_numpy(y_train))
-    train_loader = data_utils.DataLoader(train, batch_size=args.batch_size, shuffle=True)
-
-    validation = data_utils.TensorDataset(torch.from_numpy(x_val).float(), torch.from_numpy(y_val))
-    val_loader = data_utils.DataLoader(validation, batch_size=args.test_batch_size, shuffle=False)
-
-    test = data_utils.TensorDataset(torch.from_numpy(x_test).float(), torch.from_numpy(y_test))
-    test_loader = data_utils.DataLoader(test, batch_size=args.test_batch_size, shuffle=True)
-
-    return train_loader, val_loader, test_loader
-
-
-# ======================================================================================================================
-def load_dynamic_mnist(args):
-    from keras.datasets import mnist
[…]

### 2023-12-10 References/VAE_VPFLOWS/utils/perform_experiment.py

@@ -1,127 +1,0 @@
-from __future__ import print_function
-
-import torch
-
-import copy
-
-import time
-
-# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-# ======================================================================================================================
-def experiment_vae(args, train_loader, val_loader, test_loader, model, optimizer, dir, model_name='vae'):
-    if model_name == 'vae':
-        from utils.training import train_vae as train
-        from utils.evaluation import evaluate_vae as evaluate
-    elif model_name == 'vae_HF' or model_name == 'vae_ccLinIAF':
-        from utils.training import train_vae_VPflow as train
-        from utils.evaluation import evaluate_vae_VPflow as evaluate
-    else:
-        raise Exception('Wrong name of the model!')
-
-    best_model = model
-    best_loss = 1000.
-    e = 0
-    train_loss_history = []
-    train_re_history = []
-    train_kl_history = []
-
-    val_loss_history = []
-    val_re_history = []
-    val_kl_history = []
-
-    time_history = []
-
-    for epoch in range(1, args.epochs + 1):
-        time_start = time.time()
-        model, train_loss_epoch, train_re_epoch, train_kl_epoch = train(epoch, args, train_loader, model, optimizer)
-
-        val_loss_epoch, val_re_epoch, val_kl_epoch = evaluate(args, model, train_loader, val_loader, epoch, dir, mode='validation')
-        time_end = time.time()
-
-        time_elapsed = time_end - time_start
-
-        # appending history
-        train_loss_history.append(train_loss_epoch), train_re_history.append(train_re_epoch), train_kl_history.append(
-            train_kl_epoch)
-        val_loss_history.append(val_loss_epoch), val_re_history.append(val_re_epoch), val_kl_history.append(
-            val_kl_epoch)
-        time_history.append(time_elapsed)
-
-        # printing results
-        print('Epoch: {}/{}, Time elapsed: {:.2f}s\n'
-              '* Train loss: {:.2f}   (RE: {:.2f}, KL: {:.2f})\n'
-              'o Val.  loss: {:.2f}   (RE: {:.2f}, KL: {:.2f})\n'
[…]

### 2023-12-10 References/VAE_VPFLOWS/utils/training.py

@@ -1,121 +1,0 @@
-from __future__ import print_function
-
-import torch
-from torch.autograd import Variable
-
-import numpy as np
-
-import math
-
-from utils.distributions import log_Normal_diag, log_Normal_standard, log_Bernoulli
-# -=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=-=
-
-# ======================================================================================================================
-def train_vae(epoch, args, train_loader, model, optimizer):
-    # set loss to 0
-    train_loss = 0
-    train_re = 0
-    train_kl = 0
-    # set model in training mode
-    model.train()
-
-    # start training
-    if args.warmup == 0:
-        beta = 1.
-    else:
-        beta = 1.* (epoch-1) / args.warmup
-        if beta > 1.:
-            beta = 1.
-    print('beta: {}'.format(beta))
-
-    for batch_idx, (data, target) in enumerate(train_loader):
-        if args.cuda:
-            data, target = data.cuda(), target.cuda()
-        data, target = Variable(data), Variable(target)
-        # dynamic binarization
-        if args.dynamic_binarization:
-                x = torch.bernoulli(data)
-        else:
-            x = data
-        # reset gradients
-        optimizer.zero_grad()
-        # forward pass
-        x_mean, x_logvar, z_q, z_q_mean, z_q_logvar = model.forward(x)
-        # loss function
-        RE = log_Bernoulli(data, x_mean, average=False)
-        # KL
-        log_p_z = log_Normal_standard(z_q, dim=1)
-        log_q_z = log_Normal_diag(z_q, z_q_mean, z_q_logvar, dim=1)
-        KL = beta * (- torch.sum(log_p_z - log_q_z) )
-
-        loss = (-RE + KL) / data.size(0)
-        # backward pass
-        loss.backward()
-        # optimization
-        optimizer.step()
-
-        train_loss += loss.data[0]
-        train_re += (-RE / data.size(0)).data[0]
-        train_kl += (KL / data.size(0)).data[0]
-
-    # calculate final loss
-    train_loss /= len(train_loader)  # loss function already averages over batch size
-    train_re /= len(train_loader)  # re already averages over batch size
-    train_kl /= len(train_loader)  # kl already averages over batch size
-
-    return model, train_loss, train_re, train_kl
-
[…]

### 2023-12-10 References/VAE_VPFLOWS/utils/visual_evaluation.py

@@ -1,232 +1,0 @@
-import os
-
-import numpy as np
-
-import torch
-from torch.autograd import Variable
-
-import matplotlib.pyplot as plt
-import matplotlib.gridspec as gridspec
-
-#=======================================================================================================================
-def plot_histogram( x, dir, mode ):
-
-    fig = plt.figure()
-
-    # the histogram of the data
-    n, bins, patches = plt.hist(x, 100, normed=True, facecolor='blue', alpha=0.5)
-
-    plt.xlabel('Log-likelihood value')
-    plt.ylabel('Probability')
-    plt.grid(True)
-
-    plt.savefig(dir + 'histogram_' + mode + '.png', bbox_inches='tight')
-    plt.close(fig)
-
-#=======================================================================================================================
-def plot_images( args, x_sample, dir, file_name, size_x=3, size_y=3):
-
-    fig = plt.figure(figsize=(size_x, size_y))
-    # fig = plt.figure(1)
-    gs = gridspec.GridSpec(size_x, size_y)
-    gs.update(wspace=0.05, hspace=0.05)
-
-    for i, sample in enumerate(x_sample):
-        ax = plt.subplot(gs[i])
-        plt.axis('off')
-        ax.set_xticklabels([])
-        ax.set_yticklabels([])
-        ax.set_aspect('equal')
-        plt.imshow(sample.reshape(args.input_size[1], args.input_size[2]), cmap='Greys_r')
-
-    plt.savefig(dir + file_name + '.png', bbox_inches='tight')
-    plt.close(fig)
-
-#=======================================================================================================================
-def plot_real( args, x_sample, dir, size_x=3, size_y=3):
-    x_sample = x_sample.data.cpu().numpy()[:size_x*size_y]
-
-    fig = plt.figure(figsize=(size_x, size_y))
-    gs = gridspec.GridSpec(size_x, size_y)
-    gs.update(wspace=0.05, hspace=0.05)
-
-    for i, sample in enumerate(x_sample):
-        ax = plt.subplot(gs[i])
-        plt.axis('off')
-        ax.set_xticklabels([])
-        ax.set_yticklabels([])
-        ax.set_aspect('equal')
-        plt.imshow(sample.reshape(args.input_size[1], args.input_size[2]), cmap='Greys_r')
-
[…]

### 2023-12-11 Tex/slide.tex
slide 최종

@@
     \begin{itemize}
-        \item Directed Acyclic Graph(DAG) $\mathcal G = (V, E)$ with $V = \{1, \dots, m\}$
+        \item Directed Acyclic Graph(DAG) $\mathcal G = (V, E)$ with $V = \{1, \dots, m\}$ is used to represent a causal structure \footcite{pearl_2009}
         \item Linear Structural Equation Model(SEM) is defined as
@@
     \begin{itemize}
-        \item If $I-\mathbf A^T$ is nonsingular, SEM can be written as
+        \item If $(\mathbf I-\mathbf A^T)$ is nonsingular, SEM can be written as
         \begin{equation}
@@
         \item Usually we assume $\mathbf \epsilon \sim \mathcal N(\mathbf 0, \mathbf \Sigma)$ and
-        \begin{equation}
+        \begin{equation*}
             \mathbf X \sim \mathcal N(\mathbf 0, (\mathbf I - \mathbf A^T)^{-1} \mathbf \Sigma (\mathbf I - \mathbf A)^{-1})
-        \end{equation}
+        \end{equation*}
     \end{itemize}
@@
         where $\text{sib}(j) = \{i : i \sim j \text{ in } G\}$ \footcite{wang2017empirical}
+        \item Learn cyclic adjacency matrix or acyclic adjacency matrix
     \end{itemize}
@@
         \item NP-hard problem \footcite{ChickeringNPhard}
-        \item Score-based methods and constraint-based methods are used
-        \item NOTEARS : conversion of DAG learning to continuous optimization problem \footcite{zheng2018dags}
-        \begin{equation}
-            tr(A)
-        \end{equation}
+        \item Score-based methods \footcite{Spirtes2000} and constraint-based methods \footcite{ChickeringGES} are used
+        \item NOTEARS\footcite{zheng2018dags}: Combinatorial optimization to Continuous optimization problem
+        \begin{equation*}
+            \min_{\mathbf{A}} F(A) \text{ subject to } G(A) \in \mathbb{D} \Rightarrow 
+            \min_{\mathbf{A}} F(A) \text{ subject to } h(A)=0
+        \end{equation*}
+        where $\mathbb D$ denotes the discrete space of DAGs on $m$ nodes and $h(A)=0$ iff $G(A)$ is acyclic.
     \end{itemize}
+\end{frame}
+
+\begin{frame}{DAG Learning(Cont'd)}
+    \begin{block}{Theorem}
+        A matrix $\mathbf A$ is the adjacency matrix of a DAG if and only if
+        \begin{equation*}
+            h(A) = \rm{tr}(\exp(\mathbf A \circ \mathbf A)) = m
+        \end{equation*}
+        where $\circ$ denotes the Hadamard product.
+    \end{block}
+    \begin{block}{Theorem}
+        A matrix $\mathbf A$ is the adjacency matrix of a DAG if and only if
+        \begin{equation*}
+            \rm{tr}[(I + \alpha \mathbf A \circ \mathbf A)^m] = m
+        \end{equation*}
+        for any $\alpha>0$.
+    \end{block}
 \end{frame}
@@
         \end{equation}
-        \item The probability density function of $z_T$ is
+        \item Transformation of the linear IAF can be written as
         \begin{equation}
@@
         \item We can use more flexible prior distribution by using normalizing flow.
+        \item When $\mathbf z_0$ has diagonal covariance, $\mathbf z_T$ then has full covariance matrix.
     \end{itemize}
@@
     \begin{itemize}
-        \item We use the same optimization procedure as DAG-GNN \footcite{yu2019daggnn}.
         \item The optimization procedure is minimizing the following loss function:
@@
         \item The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
+        \item $\tau=0.1 \times ({m\over 50})^2$ works well in our experiment setting.
         \item The third and fourth terms are the augmented Lagrangian terms.
@@
     \end{itemize}
+\end{frame}
+
+\begin{frame}{Experiment(Cont'd)}
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.8\textwidth]{fig/mixed_graph_example.png}
+        \caption{Example of mixed graph}
+        \label{fig:mixed_graph_example}
+    \end{figure}
+\end{frame}
+
+\begin{frame}{Experiment(Cont'd)}
+    \begin{figure}
+        \centering
+        \includegraphics[width=0.8\textwidth]{fig/mixed_graph_adjacency.png}
+        \caption{Adjacency matrix of previous example graph}
+        \label{fig:mixed_graph_adjacency}
+    \end{figure}
 \end{frame}

### 2023-12-11 Tex/paper.tex
paper, slide DAG learning 추가

@@
 \modulolinenumbers[5]
-% \linenumbers
 
@@
 
-\section{Directed Acyclic Graphs and Mixed Graphs}
-\subsection{Directed Acyclic Graphs}
+\section{Preliminaries}
 
-Here is the introduction of DAGs
+\subsection{Directed Acyclic Graphs and linear SEMs}
+
+Directed acyclic graphs (DAGs) are graphs that have directed edges and no cycles, which are often used for describing causality among variables.
+To identify causal relationships between variables, structural equation models are commonly used.\\
+
+Let $A \in \mathbb{R}^{m \times m}$ be the weighted adjacency matrix of the DAG with $m$ nodes and $X \in \mathbb{R}^{m \times d}$ be a sample of a joint distribution of $m$ variables, where each row corresponds to one variable.
+In the literature, a variable is typically a scalar, but it can be trivially generalized to a $d$-dimensional vector under the current setting.
+The linear SEM model reads
+\begin{equation}
+X = A^TX + Z,
+\end{equation}
+where $Z \in \mathbb{R}^{m \times d}$ is the noise matrix.
+When the graph nodes are sorted in the topological order, the matrix $A$ is strictly upper triangular.
+Hence, ancestral sampling from the DAG is equivalent to generating a random noise $Z$ followed by a triangular solve
+\begin{equation}
+X = (I - A^T)^{-1}Z,
+\end{equation}
+
+Especially, DAG-GNN generalizes linear SEM to learn the weighted adjacency matrix $A$ of a DAG by using a deep generative model\cite{yu2019daggnn}. The generalized linear SEM follows
+\begin{equation}
+    f_2^{-1}(X) = A^Tf_2^{-1}(X) + f_1(Z),
+\end{equation}
+where $f_1$ and $f_2$ are the parameterized functions that effectively perform (possibly nonlinear) transforms on $Z$ and $X$, respectively. And $f_2$ is invertible.
 
@@
 
-Here is the introduction of mixed graphs
+There are several graphical representations of causal models, with DAGs being the most popular due to their simplicity.
+DAGs are suitable under the assumptions of causal sufficiency (i.e., no latent common causes of the observed variables), acyclicity (absence of feedback loops), and no selection bias (i.e., no implicit conditioning on a common effect of the observed variables).
+DAGs have many convenient properties, including the Markov property (with different equivalent formulations, the most prominent being d-separation) and a simple causal interpretation.
+A more general class of graphs is acyclic directed mixed graphs (ADMGs), which use bidirected edges to represent latent confounding and also have a convenient Markov property (sometimes referred to as m-separation) and causal interpretation.
+When the assumption of acyclicity is dropped, allowing for feedback, the more general class of directed mixed graphs (DMGs) can be used, which are naturally associated with (possibly cyclic) SEMs and can represent feedback loops.
+The corresponding Markov properties and causal interpretation in the cyclic case are more subtle\cite{Bongers_2021}.\\
+
+In SEMs, if error terms are independent, we can describe the relationships between variables using a DAG.
+But if error terms are dependent, a SEM can be represented by a DMG which has both directed edges and bidirected edges.
+The former is referred to as the Markovian model.
+And the latter is the semi-Markovian model\cite{Shipster2006semimarkovian}.
 
@@
 
-Here is the introduction of graph learning
+\subsubsection{Constraint-Based and Score-Based Approaches}
+
+Most constraint-based approaches test for conditional independence in the empirical joint distribution in order to construct a graph that reflects this conditional independence.
+There are often multiple graphs that fulfill a given set of conditional independence, and so it is common for constraint-based approaches to output a graph representing some Markov equivalence class (MEC).\\
+
+Score-based approaches test the validity of a candidate graph $\mathcal{G}$ according to some scoring function $S$. The goal is therefore stated as\cite{peters}:
+\begin{equation}
+    \hat{\mathcal{G}} = \underset{\mathcal{G} \ \text {over} \ X}{\arg\max} S(\mathcal{D}, \mathcal{G})
+\end{equation}
+
+where $\mathcal{D}$ represents the empirical data for variables X.
+
+\subsubsection{Continuous optimization-Based Approaches}
+
+Continuous optimization methods are pervasive in the field of deep learning, whereby highly parameterized networks are optimized using variations on gradient descent\cite{Goodfellow}.
+Recently, there have been an increasing number of methods which seek to learn structure from data, while leveraging the advantages of continuous optimization.
+These continuous optimization approaches recast the combinatoric graph-search problem into a continuous optimization problem\cite{zheng2018dags}.
+In Equation \ref{CO}, the left-hand side represents the traditional approach, which seeks the adjacency matrix A that minimizes some score function $S(A)$, subject to the implied graph $\mathcal{G}(A)$ being in the set of valid DAGs.
+The right-hand side represents a characterization of the continuous optimization problem which, again, seeks the adjacency matrix A that minimizes some score function $S(A)$, but this time subject to the constraint $h(A) = 0$.
+Here, $h$ is the function used to enforce acyclicity in the inferred graph.
+\begin{equation}
+\label{CO}
+\begin{aligned}  
+    \min_{A \in \mathbb{R}^{d \times d}}S(A)  \qquad && \qquad \min_{A \in \mathbb{R}^{d \times d}}S(A) \\
+    \text{s.t.} \quad \mathcal{G}(A) \in DAGs \qquad && \qquad \text{s.t.} \quad h(A) = 0
+\end{aligned}
+\end{equation}
+
+Recently, continuous optimization-based approaches have been explored with the attention of deep learning methods.
+NOTEARS\cite{zheng2018dags} suggests a continuous optimization algorithm to learn DAGs for the first time.
+DAG-GNN\cite{yu2019daggnn} enhances NOTEARS by incorporating neural network functions and variational inference such that the score function is the Evidence Lower BOund (ELBO).
+Our model extends DAG-GNN by applying normalizing flows into Variational Autoencoders.
 

### 2023-12-11 Tex/slide.tex
paper, slide DAG learning 추가

@@
 \usepackage{caption}
-\usepackage[backend=bibtex, style=authortitle-comp]{biblatex}
+\usepackage[backend=bibtex]{biblatex}
 \addbibresource{references.bib}
+\setbeamertemplate{bibliography item}{\insertbiblabel}
+\setbeamertemplate{frametitle continuation}[from second][(cont'd)]
 
@@
 
-\begin{frame}{DAG and SEM}
+\subsection{Graphical Model}
+
+\begin{frame}[allowframebreaks]{DAG and SEM}
     \begin{itemize}
-        \item Directed Acyclic Graph(DAG) $\mathcal G = (V, E)$ with $V = \{1, \dots, m\}$ is used to represent a causal structure \footcite{pearl_2009}
+        \item Directed Acyclic Graph(DAG) $\mathcal G = (V, E)$ with $V = \{1, \dots, m\}$ is used to represent a causal structure \cite{pearl_2009}
         \item Linear Structural Equation Model(SEM) is defined as
@@
         \item Matrix form can be written as
-        \begin{equation}
+        \begin{equation*}
             \mathbf X = \mathbf A^T \mathbf X + \mathbf \epsilon
-        \end{equation}
+        \end{equation*}
         where $\mathbf A$ is an adjacency matrix of $\mathcal G$ and $\mathbf \epsilon$ is a noise vector.
     \end{itemize}
-\end{frame}
-
-\begin{frame}{DAG and SEM(Cont'd)}
+    \framebreak
     \begin{itemize}
         \item If $(\mathbf I-\mathbf A^T)$ is nonsingular, SEM can be written as
-        \begin{equation}
+        \begin{equation*}
             \mathbf X = (\mathbf I - \mathbf A^T)^{-1} \mathbf \epsilon
-        \end{equation}
+        \end{equation*}
         \item Usually we assume $\mathbf \epsilon \sim \mathcal N(\mathbf 0, \mathbf \Sigma)$ and
@@
         \end{equation*}
+        \item In most cases, we assume $\mathbf \Sigma = \mathbf I$ or $\mathbf \Sigma = \mathbf D$ where $\mathbf D$ is a diagonal matrix.
+        \item DAG-GNN \cite{yu2019daggnn} generalizes linear SEM using a deep neural network.
+        \begin{equation*}
+            f_2^{-1}(X) = A^Tf_2^{-1}(X) + f_1(Z),
+        \end{equation*}
     \end{itemize}
@@
 
-\begin{frame}{Semi-Markovian DAG}
+\begin{frame}[allowframebreaks]{Semi-Markovian DAG}
     \begin{itemize}
-        \item Semi-Markovian DAG is a DAG where the noise variables have dependencies \footcite{Shipster2006semimarkovian}
+        \item DAGs have many convenient properties (e.g. Markov property, d-separation, etc.)
+        \item More general class of DAGs is acyclic directed mixed graphs (ADMGs) i.e. semi-Markovian DAGs
+        \item Semi-Markovian DAG is a DAG where the noise variables have dependencies \cite{Shipster2006semimarkovian}
         \item The noise vector $\mathbf \epsilon$ is a semi-Markovian noise vector if
-        \begin{equation}
+        \begin{equation*}
             \Sigma \in \mathcal{W}(G) := \{ \Sigma \in \mathbb R^{m\times m} : \Sigma_{ij} = \Sigma_{ji} = 0 \text{ if } i \notin \text{sib}(j) \}
-        \end{equation}
-        where $\text{sib}(j) = \{i : i \sim j \text{ in } G\}$ \footcite{wang2017empirical}
-        \item Learn cyclic adjacency matrix or acyclic adjacency matrix
+        \end{equation*}
+        where $\text{sib}(j) = \{i : i \sim j \text{ in } G\}$ \cite{wang2017empirical}
+        \item Learn cyclic adjacency matrix \cite{wang2017empirical} or acyclic adjacency matrix
     \end{itemize}
@@
 
-\begin{frame}{DAG Learning}
+\subsection{DAG Learning}
+
+\begin{frame}[allowframebreaks]{DAG Learning}
     \begin{itemize}
         \item DAG learning is a problem of estimating $\mathcal G$ from $\mathbf X$
-        \item NP-hard problem \footcite{ChickeringNPhard}
-        \item Score-based methods \footcite{Spirtes2000} and constraint-based methods \footcite{ChickeringGES} are used
-        \item NOTEARS\footcite{zheng2018dags}: Combinatorial optimization to Continuous optimization problem
+        \item NP-hard problem \cite{ChickeringNPhard}
+        \item Score-based methods \cite{Spirtes2000} and constraint-based methods \cite{ChickeringGES} are used
+        \item NOTEARS\cite{zheng2018dags}: Combinatorial optimization to Continuous optimization problem
         \begin{equation*}
@@
     \end{itemize}
-\end{frame}
 
-\begin{frame}{DAG Learning(Cont'd)}
     \begin{block}{Theorem}
@@
 
-\subsection*{VAE and Normalizing Flow}
+\subsection{VAE and Normalizing Flow}
 
@@
         \item It is trained by maximizing the evidence lower bound(ELBO)
-        \begin{equation}
+        \begin{equation*}
             \mathcal L(\theta, \phi; \mathbf x) = \mathbb E_{q_\phi(\mathbf z | \mathbf x)}[\log p_\theta(\mathbf x | \mathbf z)] - \text{KL}(q_\phi(\mathbf z | \mathbf x) || p(\mathbf z))
-        \end{equation}
+        \end{equation*}
         \item $q_\phi(\mathbf z | \mathbf x)$ is an approximate posterior and $p_\theta(\mathbf x | \mathbf z)$ is a generative model
@@
     \begin{itemize}
-        \item Normalizing flow transforms a simple distribution to a complex distribution \footcite{rezende15normalizingflows}
+        \item Normalizing flow transforms a simple distribution to a complex distribution \cite{rezende15normalizingflows}
         \item It is a sequence of invertible transformations $f_t : \mathbb R^d \to \mathbb R^d$
-        \begin{equation}
+        \begin{equation*}
             z_T = f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)
-        \end{equation}
+        \end{equation*}
         \item The probability density function of $z_T$ is
-        \begin{equation}
+        \begin{equation*}
             p(z_T) = p(z_0) \left| \det \left( \frac{\partial f_T \circ f_{T-1} \circ \cdots \circ f_1(z_0)}{\partial z_0} \right) \right|
-        \end{equation}
+        \end{equation*}
     \end{itemize}
@@
 
-\begin{frame}{Linear IAF}
+\begin{frame}[allowframebreaks]{Linear IAF}
     \begin{itemize}
-        \item Inverse Autoregressive Flow(IAF) is a normalizing flow that uses autoregressive transformation \footcite{kingma2016iaf}
+        \item Inverse Autoregressive Flow(IAF) is a normalizing flow that uses autoregressive transformation \cite{kingma2016iaf}
         \item Linear IAF is a special case of IAF where $f_t$ is a linear transformation
-        \begin{equation}
+        \begin{equation*}
             \bf z_t = \bf \mu_t + \bf \sigma_t \odot \bf z_{t-1}.
-        \end{equation}
+        \end{equation*}
         \item Transformation of the linear IAF can be written as
-        \begin{equation}
+        \begin{equation*}
             \mathbf{z}_T = \mathbf{L(x)} \cdot \mathbf{z}_0.
-        \end{equation}
+        \end{equation*}
         where $\mathbf{L(x)}$ is a lower triangular matrix from encoder network.
@@
         \item When $\mathbf z_0$ has diagonal covariance, $\mathbf z_T$ then has full covariance matrix.
+        \item The KL divergence term of ELBO can be written as
+        \begin{equation*}
+            \begin{aligned}
+                D_{KL}(q_\phi(\mathbf{z}_0 | \mathbf{x}) \Vert p(\mathbf{z}_T)) &=
+                \log q_\phi(\mathbf{z}_0 | \mathbf{x}) - \log p(\mathbf{z}_T) \\
+                &= -\frac{1}{2}(\mathbf{z}_0 -\mu_\phi)^T\Sigma_\phi^{-1}(\mathbf{z}_0 -\mu_\phi) + \frac{1}{2}\mathbf{z}_T^T\mathbf{z}_T \\
+            \end{aligned}
+        \end{equation*}
+        where $\mu_\phi$ and $\Sigma_\phi$ are the mean and covariance matrix from the encoder network.
     \end{itemize}
@@
         \item The optimization procedure is minimizing the following loss function:
-        \begin{equation}
+        \begin{equation*}
             \mathcal{L}(A, W, \lambda) =-\mathcal{L}_{\mathrm{ELBO}} + \tau \Vert A \Vert_1 + \lambda h(A) + \frac{c}{2} |h(A)|^2.
-        \end{equation}
+        \end{equation*}
         \item The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
@@
 
-\begin{frame}{Experiment}
+\begin{frame}[allowframebreaks]{Experiment}
     \begin{itemize}
-        \item We compare our method with the DAG-GNN \footcite{yu2019daggnn} with random graph datasets.
+        \item We compare our method with the DAG-GNN \cite{yu2019daggnn} with random graph datasets.
         \item thresholding value of extracting graph as 0.3
@@
         \item Node size : 10, 20, 30, and 50 nodes
-        \item Noise type : (Independent) Gaussian, Laplace, Exponential and (Dependent) Gaussian
+        \item Noise type
+        \begin{itemize}
+            \item Independent : $\epsilon_i \sim \mathcal N(0, 1)$, $\epsilon_i \sim \text{Laplace}(0, 1)$, $\epsilon_i \sim \text{Exp}(1)$
+            \item Dependent : $\epsilon_i \sim \mathcal N(0, \mathbf \Sigma)$ where $\mathbf \Sigma$ is a covariance matrix of semi-Markovian DAG
         \end{itemize}
-\end{frame}
-
-\begin{frame}{Experiment(Cont'd)}
+    \end{itemize}
     \begin{figure}
@@
     \end{figure}
-\end{frame}
-
-\begin{frame}{Experiment(Cont'd)}
     \begin{figure}
@@
 
-\begin{frame}{Independent Noise}
+\begin{frame}[allowframebreaks]{Independent Noise}
     \begin{columns}
@@
     \end{itemize}
-\end{frame}
-
-\begin{frame}{Independent Noise(Cont'd)}
     \begin{columns}
@@
     In the case of independent Laplace noise, our method(blue) performs better than DAG-GNN(orange).
-\end{frame}
-
-\begin{frame}{Independent Noise(Cont'd)}
     \begin{columns}
@@
     In the case of independent Exponential noise, our method(blue) performs better than DAG-GNN(orange).
-\end{frame}
 
-\begin{frame}{Independent Noise(Cont'd)}
     \begin{columns}
@@
     Considering the number of predicted edges with the SHD and FDR, our method(blue) performs better than DAG-GNN(orange).
-\end{frame}
 
-\begin{frame}
     \begin{figure}
@@
         \label{fig:ind_exponential_comparison}
+        \captionsetup{justification=centering}
+        \caption{Comparison of estimated graphs with Laplace noise}
     \end{figure}
@@
         \captionsetup{justification=centering}
-        \caption{Comparison of estimated graphs with Laplace(above) and Exponential(below) noise}
+        \caption{Comparison of estimated graphs with Exponential noise}
     \end{figure}
@@
 
-\begin{frame}{Dependent Noise}
+\begin{frame}[allowframebreaks]{Dependent Noise}
     Proportion of bidirectional edges fixed to 0.3.
@@
     \end{columns}
-\end{frame}
 
-\begin{frame}{Dependent Noise(Cont'd)}
     Proportion of bidirectional edges fixed to 0.3.
@@
     \end{itemize}
-\end{frame}
 
-\begin{frame}{Dependent Noise(Cont'd)}
     Node size $m=10$ : similar performance between DAG-GNN(orange) and our method(blue).
@@
     \end{columns}
-\end{frame}
 
-\begin{frame}{Dependent Noise(Cont'd)}
     Node size $m=20$ : our method(blue) performs better than DAG-GNN(orange).
@@
     \end{columns}
-\end{frame}
 
-\begin{frame}{Dependent Noise(Cont'd)}
     Node size $m=30$ : our method(blue) performs better than DAG-GNN(orange).
@@
     \end{columns}
-\end{frame}
 
-\begin{frame}{Dependent Noise(Cont'd)}
     Node size $m=50$ : our method(blue) performs better than DAG-GNN(orange).
@@
     \end{columns}
-\end{frame}
 
-\begin{frame}{Number of predicted edges}
     \begin{columns}
@@
     \end{columns}
+
+    \begin{itemize}
+        \item FPR : our method performs better than DAG-GNN.\\
+        \item TPR : similar performance
+    \end{itemize}
+
+    \begin{columns}
+        \begin{column}{0.5\textwidth}
+            \begin{figure}
+            \centering
+            \includegraphics[scale=0.4]{fig/TPR_dependence_30_DAG_threshold0.3.pdf}
+            \caption{TPR at $m=30$}
+            \label{fig:dep_gaussian_tpr_30}
+            \end{figure}
+        \end{column}
+        \begin{column}{0.5\textwidth}
+            \begin{figure}
+            \centering
+            \includegraphics[scale=0.4]{fig/FPR_dependence_30_DAG_threshold0.3.pdf}
+            \caption{FPR at $m=30$}
+            \label{fig:dep_gaussian_fpr_30}
+            \end{figure}
+        \end{column}
+    \end{columns}
 \end{frame}
@@
 
-\begin{frame}{Conclusion(Cont'd)}
+\begin{frame}{Conclusion}
     \begin{itemize}

### 2023-12-11 Tex/slide.tex
slide 오류수정

@@
                 \centering
-                \includegraphics[width=\textwidth]{fig/SHD_dependence_30_DAG_threshold0.3.pdf}
+                \includegraphics[width=\textwidth]{fig/SHD_dependence_30.pdf}
                 \caption{SHD with dependent Gaussian}
@@
                 \centering
-                \includegraphics[width=\textwidth]{fig/FDR_dependence_30_DAG_threshold0.3.pdf}
+                \includegraphics[width=\textwidth]{fig/FDR_dependence_30.pdf}
                 \caption{FDR with dependent Gaussian}
@@
         \centering
-        \includegraphics[height=4cm]{fig/Predicted Edges_dependence_30_DAG_threshold0.3.pdf}
+        \includegraphics[height=4cm]{fig/Predicted Edges_dependence_30.pdf}
         \caption{Number of predicted edges with dependent Gaussian}

### 2023-12-12 Tex/slide.tex
commit
수정

@@
         \begin{equation*}
-            h(A) = \rm{tr}(\exp(\mathbf A \circ \mathbf A)) = m
+            h(A) = \mathrm{tr}(\exp(\mathbf A \circ \mathbf A)) = m
         \end{equation*}
@@
         \begin{equation*}
-            \rm{tr}[(I + \alpha \mathbf A \circ \mathbf A)^m] = m
+            \mathrm{tr}[(I + \alpha \mathbf A \circ \mathbf A)^m] = m
         \end{equation*}
@@
         \item Inverse Autoregressive Flow(IAF) is a normalizing flow that uses autoregressive transformation \cite{kingma2016iaf}
-        \item Linear IAF is a special case of IAF where $f_t$ is a linear transformation
-        \begin{equation*}
-            \bf z_t = \bf \mu_t + \bf \sigma_t \odot \bf z_{t-1}.
-        \end{equation*}
+        \begin{align*}
+            \mathbf z_0 &= \mathbf \mu_0 + \mathbf \sigma_0 \odot \mathbf \epsilon, \quad \mathbf \epsilon \sim \mathcal N(\mathbf 0, \mathbf I) \\
+            \mathbf z_t &= \mathbf \mu_t + \mathbf \sigma_t \odot \mathbf z_{t-1}.
+        \end{align*}
+        \item Linear IAF is the simplest case of IAF where $f_t$ is a linear transformation
         \item Transformation of the linear IAF can be written as
@@
         where $\mathbf{L(x)}$ is a lower triangular matrix from encoder network.
+        \framebreak
+        \item $\mathbf L$ can be regarded as a cholesky decomposition of covariance matrix.
         \item We can use more flexible prior distribution by using normalizing flow.
@@
         \item The second term is the L1 regularization term, which encourages sparsity of the adjacency matrix.
-        \item $\tau=0.1 \times ({m\over 50})^2$ works well in our experiment setting.
+        \item $\tau=0.1 \times (\frac{m}{50})^2$ works well in our experiment setting.
         \item The third and fourth terms are the augmented Lagrangian terms.
@@
         \item SHD measures the number of edge additions, deletions, and reversals required to transform the estimated graph into the true graph.
-        \item FDR represents the ratio of false positives to the total number of predicted edges.
+        \item FDR measures the proportion of incorrectly predicted edges among the predicted edges.
         \item These metrics are calculated by comparing the estimated graph with the true DAG.
@@
     \end{columns}
+    \framebreak
 
